---
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
    toc_float: true
    theme: readable
    highlight: zenburn
---

Report last updated `r date()`.


```{r knitrsetup, echo=FALSE}
library(myRfunctions)
library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(dev="png",
                      highlight=TRUE, autodep=TRUE,
                      comment='', fig.cap='',
                      tidy.opts=list(keep.blank.line=FALSE, width.cutoff=200),
                      cache=TRUE,warning=FALSE,message=FALSE,
                      error=FALSE, fig.width = 12, fig.height = 9)
options(bitmapType = 'cairo')
```


```{r libraries}
library(ggplot2)
library(reshape)
library(limma)
library(CHBUtils)
library(biomaRt)
library(dplyr)
library(cluster)
library(gridExtra)
library(logging)
library(org.Mm.eg.db)
library(DEGreport)
library(dplyr)
order_group=c("control", "day1", "day2", "day7", "day14")
# root_path = "~/orch/scratch/vishal_mirna_kidney/publish/FA-model"
result_files = file.path(root_path, "omics/files_publish")
dir.create(result_files, showWarnings = F, recursive = T)
basicConfig(level='INFO')
```

# Protein analysis {.tabset}

FC > 4 and FDR < 1%

```{r protein-de, results='asis'}
protein_file = file.path(root_path, "protein", "files_publish")
dir.create(protein_file, recursive = T, showWarnings = F)
counts = read.csv(file.path(root_path, "protein", "FA_Proteomics_RawData.csv"),skip = 3) %>% tidyr::separate(Protein.Id, c("sp", "Uniprot.ID", "type"),sep = "[::|::]", extra = "merge")
row.names(counts) = counts$Uniprot.ID

mapping = data.frame(id=rownames(counts))
mapping = annotate_df(mapping, "id", 'mmusculus_gene_ensembl', "uniprot_genename", "ensembl_gene_id")  %>% distinct(ensembl_gene_id)

mapping2 = data.frame(id=unique(counts$Gene.Symbol)[!is.na(unique(counts$Gene.Symbol))])
mapping2 = annotate_df(mapping2, "id", 'mmusculus_gene_ensembl', "wikigene_name", "ensembl_gene_id")  %>% distinct(ensembl_gene_id)

counts[as.character(mapping$id), "ensembl"] = mapping$ensembl_gene_id 
idx = match(as.character(mapping2$id),as.character(counts$Gene.Symbol))
counts[idx, "ensembl"] = mapping2$ensembl_gene_id
counts$symbol = convertIDs(as.character(counts$ensembl), "ENSEMBL", "SYMBOL", org.Mm.eg.db)
save_file(counts, "counts_annotated.csv", protein_file)

counts = counts %>% distinct(ensembl) %>% filter(!is.na(ensembl))
clean_counts = counts[,7:16]
rownames(clean_counts) = counts$ensembl
names(clean_counts) = sapply(names(clean_counts), tolower)
names(clean_counts)[1:10] = c("control-1", "control-2", "day1-1", "day1-2",
                       "day2-1", "day2-2", "day7-1", "day7-2", "day14-1", "day14-2")

cat("## Expression density of 'raw' data\n\n")
ggplot(melt(clean_counts), aes(x=log2(value), colour=variable)) + geom_density()

dge = edgeR::DGEList(clean_counts)
dge = edgeR::calcNormFactors(dge, method="TMM")

coldata = data.frame(row.names=colnames(clean_counts), samples=colnames(clean_counts)) %>% tidyr::separate(samples,into = c("group"), extra = "drop")
coldata$group = sapply(coldata$group, tolower)
norm_counts = edgeR::cpm(dge, log=T)

MA <- normalizeBetweenArrays(as.matrix(norm_counts), method="none")

cat("## Expression density of normalize data\n\n")
ggplot(melt(as.data.frame(MA)), aes(x=value, colour=variable)) + geom_density()

cat("## MDS plot\n\n")
mds(MA, condition = coldata$group)
save_file(MA, "fa_model_log2_counts.tsv", protein_file)
design <- model.matrix(~ 0 + coldata$group)
colnames(design) = c("day1", "day14", "day2", "day7", "normal")
fit <- lmFit(MA, design)
contrast.matrix <- makeContrasts(day1-normal, day2-day1, day7-day2, day14-day7, 
                                 day2-normal, day7-normal, day14-normal,
                                 day7-day1, day7-day2, 
                                 day14-day1, day14-day2,
                                 levels=design)
fit2 <- contrasts.fit(fit, contrast.matrix)
fit2 <- eBayes(fit2)
res = topTable(fit2, adjust="BH", n="Inf")
save_file(res, "fa_model.tsv", protein_file)
absMaxFC = rowMax(as.matrix(abs(res[,1:11])))
sign = row.names(res[absMaxFC>2 & res$adj.P.Val<0.01,])

cat("## MDS plot of DE proteins\n\n")
mds(MA[sign,], condition = coldata$group)

coldata$group = factor(coldata$group, levels=order_group)
cat("## Clustering analysis\n\n")
clusters = degPatterns(MA[sign,], coldata, minc = 15, summarize = "group", time="group", col = NULL)
.df=clusters$df
save_file(.df, "clusters_genes_cluster.tsv", protein_file)

```

# 3D interaction analysis

```{r load, results='asis'}

mrna_path = "rnaseq/final/2016-02-05_mrna_bio/files_publish"
mrna_matrix = read.csv(file.path(root_path, mrna_path, "fa_model_log2_counts.tsv"), row.names = 1)
mrna_clusters = read.csv(file.path(root_path, mrna_path, "clusters_genes.tsv"), row.names = 1)
mrna_col = data.frame(row.names=colnames(mrna_matrix), samples=colnames(mrna_matrix)) %>% tidyr::separate(samples,into = c("group"), extra = "drop") %>% mutate(group=factor(group, levels=order_group))
rownames(mrna_col) = colnames(mrna_matrix)
keep = rownames(mrna_col)[!is.na(mrna_col$group)]
mrna_col = mrna_col[keep,,drop=F]
mrna_matrix = mrna_matrix[,keep]

prot_path = "protein/files_publish"
prot_matrix = read.csv(file.path(root_path, prot_path, "fa_model_log2_counts.tsv"), row.names = 1)
prot_clusters = read.csv(file.path(root_path, prot_path, "clusters_genes_cluster.tsv"), row.names = 1)
prot_col = data.frame(row.names=colnames(prot_matrix), samples=colnames(prot_matrix)) %>% tidyr::separate(samples,into = c("group"), extra = "drop") %>% mutate(group=factor(group, levels=order_group))
rownames(prot_col) = colnames(prot_matrix)

mirna_path = "srnaseq/final/files_publish"
mirna_matrix = read.csv(file.path(root_path, mirna_path, "fa_model_mirna_log2_counts.tsv"), row.names = 1)
load(file.path(root_path, mirna_path, "ma.rda"))
mirna_col = data.frame(row.names=colnames(mirna_matrix), samples=colnames(mirna_matrix)) %>% tidyr::separate(samples,into = c("id", "group"), extra = "drop") %>% mutate(group=factor(gsub("day0", "control", tolower(group)), levels=order_group))
rownames(mirna_col) = colnames(mirna_matrix)
keep = rownames(mirna_col)[!is.na(mirna_col$group)]
mirna_col = mirna_col[keep,,drop=F]
mirna_matrix = mirna_matrix[,keep]

```

```{r plot, results='asis'}
library(SummarizedExperiment)
library(isomiRs)
mrna_keep = as.character(mrna_clusters$genes)
prot_keep = unique(c(intersect(mrna_keep, rownames(prot_matrix)), as.character(prot_clusters$genes)))
keep = intersect(mrna_keep, rownames(prot_matrix))

df = degMerge(list(mrna=mrna_matrix[keep,], prot=prot_matrix[keep,]), 
                   list(mrna=mrna_clusters), 
                   list(mrna=mrna_col, prot=prot_col), 
                   summarize="group", 
                   time="group", col=NULL, 
                   scale=TRUE)
prot_merged = do.call(rbind, df)
prot_merged$name = convertIDs(as.character(prot_merged$gene), "ENSEMBL", "SYMBOL", org.Mm.eg.db)
names(prot_merged) = paste0("prot_", names(prot_merged))

mi_rse = SummarizedExperiment(assays=SimpleList(norm=as.matrix(mirna_matrix)), colData= mirna_col)
gene_rse = SummarizedExperiment(assays=SimpleList(norm=as.matrix(mrna_matrix)), colData=mrna_col)
cor_tar = find_targets(mi_rse, gene_rse, ma)
mapping = melt(cor_tar) %>% filter(value<0)
keep = intersect(mrna_keep, as.character(unique(mapping[,1])))
mirna_keep = as.character(unique(mapping[,2]))

df = degMerge(list(mrna=mrna_matrix[keep,],
                   mirna=mirna_matrix[mirna_keep,]), 
              list(mrna=mrna_clusters), 
              list(mrna=mrna_col, mirna=mirna_col), 
              summarize="group", 
              time="group", col=NULL, 
              scale=TRUE, mapping=list(mirna=mapping))
mirna_merged = do.call(rbind, df)
mirna_merged$name = convertIDs(as.character(mirna_merged$pair), "ENSEMBL", "SYMBOL", org.Mm.eg.db)
names(mirna_merged) = paste0("mirna_", names(mirna_merged))

merged = merge(mirna_merged[mirna_merged$mirna_cor < -.6,c(6,7,1:5)], prot_merged[prot_merged$prot_cor > .6,c(5,7,1:4)], by=1, all=TRUE)
merged_all = merge(mirna_merged[c(6,7,1:5)], prot_merged[c(5,7,1:4)], by=1, all=TRUE)

save_file(merged, "omics_filtered.csv", result_files)
save_file(merged_all, "omics.csv", result_files)
save_file(prot_merged, "prot_omics.csv", result_files)
save_file(mirna_merged, "mirna_omics.csv", result_files)
```

# R Session Info

(useful if replicating these results)

```{r sessioninfo}
sessionInfo()
```


