---
output:
    knitrBootstrap::bootstrap_document:
        theme: readable
        highlight: zenburn
        theme.chooser: TRUE
        highlight.chooser: TRUE
    html_document:
        toc: true
        highlight: zenburn

---
```{r custom-code}
source("~/repos/myRfunctions/transactions.R")
dn <- "slack_mirna-mrna/"
path_files = "~/repos/pipelines/slack_cel_network/network"
```


```{r setup, echo=FALSE, eval=FALSE}
knitr::opts_chunk$set(tidy=TRUE, highlight=TRUE, dev="png",
                      cache=FALSE, highlight=TRUE, autodep=TRUE, warning=FALSE, error=FALSE,
                      eval=TRUE,
                      message=FALSE, prompt=TRUE, comment='', fig.cap='', bootstrap.show.code=FALSE)
render_2_drop("~/repos/pipelines/slack_cel_network/network/interaction.rmd",dn)

```
# Introduction

The main idea is to find miRNA that correlated to genes that have been found deregulated in RNAseq data. 
Currently using targetscanworm for the miRNA-gene pairs matrix.
miRNA-mRNA interactions with (negative) correlation < -.7 and FDR<0.1 are considered.

# Load data


```{r load}
suppressMessages(library(DESeq2,quietly=TRUE))
library(ggplot2)
library(dplyr)
library(tidyr)

mirna<-read.table("~/repos/pipelines/slack_cel_network/reports/new_smallrna/log_matrix_perfect.txt",header=T,row.names=1)
row.names(mirna)<-sub(".ref.0.0.0.0","",row.names(mirna))


mrna<-read.table("~/repos/pipelines/slack_cel_network/reports/rlog_mat.txt",header=T,row.names=1)
names(mrna) = c("N2_1_Day0","N2_1_Day5","N2_2_Day0","N2_2_Day5","N2_3_Day0","N2_3_Day5",     "mir71_1_Day0", "mir71_1_Day5","mir71_2_Day0","mir71_2_Day5","mir71_3_Day0","mir71_3_Day5")
#tf<-read.table("tf.list2",sep=" ")
#chea<-read.table("chea-database",header=F,sep=",")

n2_labels = c("N2_1_Day0","N2_1_Day5","N2_2_Day0","N2_2_Day5","N2_3_Day0","N2_3_Day5")
ko_labels = c("mir71_1_Day0","mir71_1_Day5","mir71_2_Day0","mir71_2_Day5","mir71_3_Day0","mir71_3_Day5")

```


Load target~miRNA (targetscanWorm) table and deregulated genes (from RNAseq data)
```{r load-targets}
gene_map = read.table("~/orch/scratch/celegans_network/celegans_network_raw/data/gene_map.dat",as.is=c(TRUE,TRUE))
ago = gene_map %>%filter(grepl("ago",V3))
target<-read.table("~/orch/scratch/celegans_network/celegans_network_raw/data/targets.csv",as.is=c(TRUE,TRUE))
target_tf<-read.table("~/orch/scratch/celegans_network/celegans_network_raw/data/aging_master_table.txt",as.is=c(TRUE,TRUE))
```


# Methods
Create function to detect negative correlation and get pvalue.
Very basic method: pearson correlation.
```{r cor-functions}
getcor_gene<-function(g,logmrna="",logmirna=""){
  #print(g)
  mir<-target[target[,2]==g,1]
  mir<-intersect(mir,row.names(logmirna))
  if (length(mir)>0){
  gexp=as.numeric(logmrna[g,])
  #print(mir)
  corlist<-do.call("rbind",(lapply(mir,function(m){
    mexp<-unlist(logmirna[m,])
    #print(mexp)
    #print(gexp)
    c<-cor.test(gexp,mexp,method="pearson")
    return(data.frame(rho=c$estimate,pvalue=c$p.value))
    })))
  #print(corlist)
  return(cbind(data.frame(g=g,mir=mir), corlist))
  }else{
    return(data.frame(g=g,mir="mir",rho=NA,pvalue=NA))
  }
}


getcor_mirna<-function(mir,logmrna="",logmirna=""){
  #print(g)
  genes<-target[target[,1]==mir,2]
  genes<-intersect(genes,row.names(logmrna))
  if (length(genes)>0){
  mexp=as.numeric(logmirna[mir,])
  #print(mir)
  corlist<-do.call("rbind",(lapply(genes,function(g){
    gexp<-unlist(logmrna[g,])
    #print(mexp)
    #print(gexp)
    c<-cor.test(gexp,mexp,method="pearson")
    return(data.frame(rho=c$estimate,pvalue=c$p.value))
    })))
  #print(corlist)
  return(cbind(data.frame(g=genes,mir=mir), corlist))
  }else{
    return(data.frame(g="genes",mir=mir,rho=NA,pvalue=NA))
  }
}
#testing 
#getcor("WBGene00000012",mrna[, n2_labels],mirna[, n2_labels])
```


```{r cor-wrapper}
getTargets = function(de,genes,micros,fn_cor){
    res<-do.call("rbind",lapply(de,fn_cor,logmrna=genes,logmirna=micros))
    res = res %>% filter(!is.na(rho) & rho < 0) %>% mutate(adj=p.adjust(pvalue,method="BH"))
    #hist(res$pvalue)
    res[order(res$rho,res$pvalue),]
}
```

apply function to data


```{r n2-cor, cache=FALSE}
n2_de<-read.table("~/repos/pipelines/slack_cel_network/reports/n2_de_genes.txt",as.is=TRUE)
n2_deg_mirna<-read.table("~/repos/pipelines/slack_cel_network/reports/new_smallrna/mirna_n2_de_genes_perfect.txt",header=T,row.names=1)
row.names(n2_deg_mirna)<-sub(".ref.0.0.0.0","",row.names(n2_deg_mirna))
n2_res = getTargets(as.character(n2_de$id), mrna[, n2_labels], mirna[, n2_labels],getcor_gene)
n2_mir_res = getTargets(as.character(row.names(n2_deg_mirna)), mrna[, n2_labels], mirna[, n2_labels],getcor_mirna)
n2_both_res = getTargets(as.character(row.names(n2_deg_mirna)), mrna[n2_de$id, n2_labels], mirna[, n2_labels],getcor_mirna)
n2_res_sign = n2_res %>% filter(adj<0.1 & rho< -.7)
n2_mir_res_sign = n2_mir_res %>% filter(adj<0.1 & rho< -.7)
n2_both_res_sign = n2_both_res %>% filter(adj<0.1 & rho< -.7)

ko_de<-read.table("~/repos/pipelines/slack_cel_network/reports/ko_de_genes.txt",as.is=TRUE)
ko_deg_mirna<-read.table("~/repos/pipelines/slack_cel_network/reports/new_smallrna/mirna_ko_de_genes_perfect.txt",header=T,row.names=1)
row.names(ko_deg_mirna)<-sub(".ref.0.0.0.0","",row.names(ko_deg_mirna))
ko_res = getTargets(as.character(ko_de$id), mrna[, ko_labels], mirna[, ko_labels],getcor_gene)
ko_mir_res = getTargets(as.character(row.names(ko_deg_mirna)), mrna[, ko_labels], mirna[, ko_labels],getcor_mirna)
ko_both_res = getTargets(as.character(row.names(ko_deg_mirna)), mrna[ko_de$id, ko_labels], mirna[, ko_labels],getcor_mirna)
ko_res_sign = ko_res %>% filter(adj<0.1 & rho< -.7)
ko_mir_res_sign = ko_mir_res %>% filter(adj<0.1 & rho< -.7)
ko_both_res_sign = ko_both_res %>% filter(adj<0.1 & rho< -.7)


```


# Results 

```{r interaction-tables}
fn_n2 = paste0(path_files,"/n2_interaction.txt")
write.table(n2_both_res_sign, fn_n2)
fn_ko = paste0(path_files,"/ko_interaction.txt")
write.table(ko_both_res_sign, fn_ko)
```

[n2 interactions](`r get_report_links(fn_n2)`)

[ko interactions](`r get_report_links(fn_ko)`)


We are using DEG miRNA and DEG genes. We calculate the correlation for the pairs
annotated in targetscanWorm and apply a p.adjust correction to keep only
pairs with correlation < -.7 and p-adjusted < 0.1 (10%)

## N2


### Results for miRNA-mRNA

Just relations with FDR<0.1 and rho< -0.7. Total of `r nrow(n2_both_res_sign)` interactions


by `r length(unique(n2_both_res_sign$mir))` miRNAs

```{r n2_genes_mirna}

mir.sum<-ddply(n2_both_res_sign,"mir",summarise,N=length(mir))
ggplot(mir.sum,aes(N))+geom_bar()+
  theme_bw() + labs(list(title="Distribution of genes/miRNA",x="targets per miRNA"))
```

and `r length(unique(n2_both_res_sign$g))` genes

```{r n2_mirna_genes}

mir.sum<-ddply(n2_both_res_sign,"g",summarise,N=length(mir))
ggplot(mir.sum,aes(N))+geom_bar()+
  theme_bw() + labs(list(title="Distribution of sites/gene ",x="miRNA per gene"))
```

### Interaction in custom pairs

```{r n2-custom-pairs}
knitr::kable( n2_both_res_sign %>% mutate(id=paste0(mir,g)) %>% filter( id %in% paste0(target_tf$V1,target_tf$V2)  ) )
```


### Network

Some miRNA target more than one gene (top 200 interactions)

```{r n2_network}
plotNetwork = function(dd){
    suppressMessages(library(igraph))
    gr<-graph.data.frame(dd[1:200,c(2,1)],directed = F)
    V(gr)$color<-"steelblue"
    V(gr)$color[grep("cel-",V(gr)$name)]<-"yellow2"
    V(gr)$name[grepl("WBG",V(gr)$name)]=""
    V(gr)$name<-sub("cel-","",V(gr)$name)
    plot.igraph(gr, layout=layout.auto,vertex.color=V(gr)$color,vertex.label.cex=1,vertex.size=15,edge.width=2,vertex.label.color="black")
}

plotNetwork(n2_both_res_sign)

cytoscape = function(dd, gdeg, mdeg){
    mdeg$id = row.names(mdeg)
    add_gdeg = dplyr::inner_join(dd[,c(2,1,3:5)], gdeg %>% dplyr::select(g=id,genelog2FC=log2FoldChange), by = "g") 
    add_mdeg = dplyr::inner_join(add_gdeg, mdeg %>% dplyr::select(mir=id,mirlog2FC = log2FoldChange), by = "mir") 
    add_mdeg

}

n2_cyto = cytoscape(n2_both_res_sign, n2_de, n2_deg_mirna) 
fn_n2 = paste0(path_files,"/n2_cytoscape.txt")
write.table(n2_cyto, fn_n2,row.names=F, quote = F)
```

[n2 cytoscape file](`r get_report_links(fn_n2)`)

### Some examples

```{r n2_examples}

plotTop = function(dd, genes, micros, labels, groups){
    for (i in 1:5){
        mie<-unlist(micros[as.character(dd[i,2]),labels])
        me<-unlist(genes[as.character(dd[i,1]),labels])
        d<-data.frame(mir=mie,mrna=me,groups=factor(groups,levels=unique(groups)))
        p<-ggplot(d)+
            geom_point(aes(mir,mrna,colour=groups))+
            scale_color_discrete("miRNA exp")+
            theme_bw()+
            labs(list(x=as.character(dd[i,2]),
                      y=as.character(dd[i,1])))
        plot(p)
        }
}
con = c("Day0","Day5","Day0","Day5","Day0","Day5")
plotTop(n2_both_res_sign, mrna, mirna, n2_labels, con)

```

### miRNA vs Gene

```{r n2-mirna-genes,fig.width=11,fig.height=11}
scale_exp = function(v){
    mx = max(v)
    mn = min(v)
    (v-mn)/(mx-mn)
}

fix_name = function(s){
    strsplit(s,"_")[[1]][3]
}

dat = data.frame()
for (m in unique(n2_both_res_sign$mir)){
    genes = n2_both_res_sign %>% dplyr::filter(mir==m) %>% dplyr::select(g)
    ma = as.data.frame(t(apply(mrna[as.character(genes$g),n2_labels],1,scale_exp))) %>% mutate(type="mRNA",gene=genes$g)
    mir = scale_exp(mirna[m,n2_labels])
    ma = tidyr::gather(ma,"sample","value",1:6)
    dd = rbind(ma,data.frame(type="mir",variable=names(mir),value=unlist(mir[1,]),gene="mir"))
    dd$time = sapply(as.character(dd$variable),fix_name)
    dd = dd %>% group_by(gene,type,time) %>% dplyr::summarise(expression=mean(value)) %>% ungroup()
    dd$gene = as.character(dd$gene)
    dd$type = as.character(dd$type)
    dd$time = as.character(dd$time)
    dd$mir = as.character(m)
    dat = rbind(dat,dd)
}

ggplot(dat,aes( x=time, y=expression, color=type, group=gene )) +
    geom_line() + labs(list(y="% expression")) +
    scale_color_brewer(guide=FALSE,palette = "Set1")+
    facet_wrap(~mir)



```


### Enrichment in DEG miRNAs/Genes

We assume that should be an enrichment of de-regulated miRNA and genes that have been detected here.
To calculate this, we create a background distribution doing 400 permutations. Each time, we randomly select the same number of miRNAs (we detected in the analysis) from the list of possible miRNA-Gene pairs. That list is containing only expressed miRNA in this dataset. 
After that we can calculate the overlap between the DEG list and the real miRNAs or the random miRNAs. 
To get a pvalue of enrichment, we apply: sum(real_overlap > random_overlap)/401. That is show in the figure as a number bext to the red line which indicates the observed value from the data.

If the overlap between the miRNAs targetting genes and DEG miRNAs is bigger than the values comming from 
the permutations, we can say that the overlap is above a random noise.

For miRNA:

```{r n2_mir_enrichment}
getEnrichment = function(dd, sign, interaction, exp, col){
    require(gplots)
    interaction = intersect(interaction,exp)
    common = length(intersect(as.character(unique(dd[,col])),as.character(sign)))
    #venn( list( interaction=as.character(unique(dd[,col])),DEG=as.character(sign) ) )
    random<-unlist(lapply(1:400,function(x){
        r<-sample( as.character(exp),length( unique(dd[,col]) ) ) # random interaction miRNAs
        o<-length(intersect(as.character(sign),as.character(r))) # intersect random interaction w dergulated mirna
        return(o)
        }))
    plot(density(unlist(random)),xlim=c(min(c(common,random)),max(c(common,random))),col="steelblue",xlab="",main="")
    abline(v=common,col="red")
    text(common,0.02,format(sum(random>=common)/401, digit=3),pos = 2)
    
}


getEnrichment(n2_both_res_sign, row.names(n2_deg_mirna), unique(target[,1]), row.names(mirna), 2)
title("miRNA with interaction enriched in DEG miRNAs along age")
```

For genes: In this case I randomly select the same number of genes we detec as targeted by miRNAs.

```{r n2_gene_enrichment}
getEnrichment(n2_both_res_sign, n2_de$id ,unique(target[,2]), row.names(mrna), 1)
title("Genes with interaction enriched in DEG genes along age")
```


## mir71 KO

### Results for miRNA-mRNA

Just relations with FDR<0.1 and rho< -0.7. Total of `r nrow(ko_both_res_sign)` interactions


by `r length(unique(ko_both_res_sign$mir))` miRNAs

```{r ko_genes_mirna}

mir.sum<-ddply(ko_both_res_sign,"mir",summarise,N=length(mir))
ggplot(mir.sum,aes(N))+geom_bar()+
  theme_bw() + labs(list(title="Distribution of genes/miRNA",x="targets per miRNA"))
```

and `r length(unique(ko_both_res_sign$g))` genes

```{r ko_mirna_genes}

mir.sum<-ddply(ko_both_res_sign,"g",summarise,N=length(mir))
ggplot(mir.sum,aes(N))+geom_bar()+
  theme_bw() + labs(list(title="Distribution of sites/gene ",x="miRNA per gene"))
```

### Interaction in custom pairs

```{r ko-custom-pairs}
knitr::kable( ko_both_res_sign %>% mutate(id=paste0(mir,g)) %>% filter( id %in% paste0(target_tf$V1,target_tf$V2)  ) )
```


### Network

Some miRNA target more than one gene (top 200 interactions)

```{r ko_network}
plotNetwork(ko_both_res_sign)
ko_cyto = cytoscape(ko_both_res_sign, ko_de, ko_deg_mirna) 
fn_ko = paste0(path_files,"/ko_cytoscape.txt")
write.table(ko_cyto, fn_ko,row.names=F, quote = F)
```

[ko cytoscape file](`r get_report_links(fn_ko)`)

### Some examples

```{r ko_examples}

con = c("Day0","Day5","Day0","Day5","Day0","Day5")
plotTop(ko_both_res_sign, mrna, mirna, ko_labels, con)

```

### miRNA vs Gene

```{r ko-mirna-genes,fig.width=11,fig.height=11}

dat = data.frame()
for (m in unique(ko_both_res_sign$mir)){
    genes = ko_both_res_sign %>% dplyr::filter(mir==m) %>% dplyr::select(g)
    ma = as.data.frame(t(apply(mrna[as.character(genes$g),ko_labels],1,scale_exp))) %>% mutate(type="mRNA",gene=genes$g)
    mir = scale_exp(mirna[m,ko_labels])
    ma = tidyr::gather(ma,"sample","value",1:6)
    dd = rbind(ma,data.frame(type="mir",variable=names(mir),value=unlist(mir[1,]),gene="mir"))
    dd$time = sapply(as.character(dd$variable),fix_name)
    dd = dd %>% group_by(gene,type,time) %>% dplyr::summarise(expression=mean(value)) %>% ungroup()
    dd$gene = as.character(dd$gene)
    dd$type = as.character(dd$type)
    dd$time = as.character(dd$time)
    dd$mir = as.character(m)
    dat = rbind(dat,dd)
}

ggplot(dat,aes( x=time, y=expression, color=type, group=gene )) +
    geom_line() + labs(list(y="% expression")) +
    scale_color_brewer(guide=FALSE,palette = "Set1")+
    facet_wrap(~mir)

```

### Enrichment in DEG miRNAs/Genes

For miRNA

```{r ko_mir_enrichment}

getEnrichment(ko_both_res_sign, row.names(ko_deg_mirna), unique(target[,1]), row.names(mirna), 2)
title("miRNA with interaction enriched in DEG miRNAs along age")
```

For Genes

```{r ko_gene_enrichment}
getEnrichment(ko_both_res_sign, ko_de$id ,unique(target[,2]), row.names(mrna), 1)
title("Genes with interaction enriched in DEG genes along age")
```


# Comparison of N2 and mir71-KO

We calculate whether the common miRNAs/Genes are more or less than expected if
the two list of interactions, in both conditions, were taken randomly.
We work only with all interactions where the miRNA and the mRNA are expressed in our
conditions.

## Common miRNAs

```{r common-mirna}
get_commons = function(dd1, dd2, cols=1){
    x1_vec = unique(n2_both_res_sign[,cols])
    x2_vec = unique(ko_both_res_sign[,cols])

    length(  intersect(x1_vec, x2_vec) )/min( length(x1_vec), length(x2_vec) )
}

venn(list( n2=unique(n2_both_res_sign$mir), ko=unique(ko_both_res_sign$mir) ) )
n2_ratio_mir = get_commons(n2_both_res_sign, ko_both_res_sign, 2)
n2_ratio_gene = get_commons(n2_both_res_sign, ko_both_res_sign, 1)

```

Expected overlap

```{r expected-mirna}
expected_overlap = function(perm, dd1="",dd2="",u="", cols=1){
    x1 = nrow(dd1)
    x2 = nrow(dd2)
    rdd1 = (dplyr::sample_n(u, x1))[,cols]
    rdd2 = (dplyr::sample_n(u, x2))[,cols]
    n = length( intersect( unique(rdd1), unique(rdd2) ) )
    n/(min( length(unique(rdd1)), length(unique(rdd2)) ))
}

plot_random = function(random,common){
    plot(density(unlist(random)),xlim=c(min(c(common,random)),max(c(common,random))),col="steelblue",xlab="",main="")
    abline(v=common,col="red")
    text(common,0.02,format(sum(random>=common)/401, digit=3),pos = 2)
}

universe = target[,2:1] %>% filter(V1 %in% row.names(mirna) & V2 %in% row.names(mrna))

random_overlap = sapply(1:400, expected_overlap, dd1=n2_both_res_sign,dd2=ko_both_res_sign,u=universe,cols=2)

plot_random(random_overlap, n2_ratio_mir)
```


## Common genes

```{r common-genes}
venn(list( n2=unique(n2_both_res_sign$g), ko=unique(ko_both_res_sign$g) ) )
dataset = "celegans_gene_ensembl"
filter = "ensembl_gene_id"
symbol = "external_gene_name"
entres <- "entrezgene"
annotate_df = function(g, dataset, filter, symbol) {
  require(biomaRt)
  ensembl = useMart('ensembl', dataset = dataset)
  getBM(attributes=c(filter, symbol, "description"), filters=filter, values=g,
                    mart=ensembl)
  
}
```



```{r random-genes}
random_overlap = sapply(1:400, expected_overlap, dd1=n2_both_res_sign,dd2=ko_both_res_sign,u=universe,cols=1)

plot_random(random_overlap, n2_ratio_gene)

```

All this it could mean that the common targeted genes overlap more than expected by random,
probably because they represent changes during time. But the common miRNAs are less than
expected by random, maybe because the two conditions have a different set of miRNA
expressed and the regulatory network have changed to regulated the mainly genes related 
to the phenotype.

## Difference in numbers of targets for each miRNA

```{r mirna-diff-targets,results='asis'}
dd = merge(n2_both_res_sign %>% group_by(mir) %>% dplyr::summarise( n2 = n()) ,
ko_both_res_sign %>% group_by(mir) %>% dplyr::summarise( ko = n()) ,all=T)

dd = left_join(dd, target %>% filter(grepl("WBG",V2)) %>% group_by(V1) %>% dplyr::summarise( pairs = n()) %>% mutate(ratio=pairs/length(unique(target$V2)), mir=V1) , by='mir') %>% dplyr::select(mir,n2,ko,ratio)

knitr::kable(dd)

```


## tables

```{r genes-tables}
fn_common = paste0(path_files,"/genes_common.txt")
common_anno = annotate_df( intersect(unique(n2_both_res_sign$g), unique(ko_both_res_sign$g)) , dataset, filter, symbol)
write.table(common_anno,fn_common)

fn_n2 = paste0(path_files,"/genes_n2.txt")
n2_anno = annotate_df( setdiff(unique(n2_both_res_sign$g), unique(ko_both_res_sign$g)) , dataset, filter, symbol)
write.table(n2_anno,fn_n2)

fn_ko = paste0(path_files,"/genes_ko.txt")
ko_anno = annotate_df( setdiff(unique(ko_both_res_sign$g), unique(n2_both_res_sign$g)) , dataset, filter, symbol)
write.table(ko_anno,fn_ko)
```

[n2 genes](`r get_report_links(fn_n2)`)

[ko genes](`r get_report_links(fn_ko)`)

[common genes](`r get_report_links(fn_common)`)

```{r based-interaction-tables}
fn_common = paste0(path_files,"/interaction_common.txt")
n2 = n2_both_res_sign %>% mutate(id = paste0(g,mir))
ko = ko_both_res_sign %>% mutate(id = paste0(g,mir))
common_anno = n2 %>% filter(id %in% ko$id) %>% dplyr::select(g,mir)
write.table(common_anno,fn_common)

fn_dif_mir = paste0(path_files,"/interaction_mir_substitution.txt")
same_gene = inner_join(n2[,1:2],ko[,1:2],by="g") %>% filter(! mir.x %in% mir.y) %>% transmute(n2.mir=mir.x,ko.mir=mir.y)
write.table(same_gene,fn_dif_mir)
```

[common interaction](`r get_report_links(fn_common)`)

[common genes with different miRNA regulation](`r get_report_links(fn_dif_mir)`)
