---
title: "chip-seq overlap and annotation"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    theme: readable
    highlight: zenburn
---
   
```{r projsetup, echo=FALSE, warning=FALSE, message=FALSE}
project="chip-seq"
labPI="Kate Jeffrey’s"
lablocation=""
analystname="Lorena Pantano"
analystemail="lpantano@hsph.harvard.edu"

```


```{r knitrsetup, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
library(rmarkdown)
library(myRfunctions)
require(TxDb.Hsapiens.UCSC.hg19.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene

knitr::opts_chunk$set(tidy=TRUE, highlight=TRUE, dev="png",
                      cache=TRUE, highlight=TRUE, autodep=TRUE, 
                      warning=FALSE, error=FALSE,
                      eval=TRUE, fig.width= 11, fig.height = 11,
                      message=FALSE, prompt=TRUE, comment='', fig.cap='', bootstrap.show.code=FALSE)
options(bitmapType = 'cairo')
```


```{r render,echo=FALSE,eval=FALSE}
# render_2_drop("overlap.Rmd", "jeffrey_chipseq")
```

---

# Chip-seq analysis

Analysis for`r labPI` group.  

Contact `r analystname` (`r analystemail`) for additional details.

The most recent update of this html document occurred: `r date()`

The sections below provide code to reproduce the included results and plots. 

---

## Setup

### Libraries

```{r functions, warning=FALSE, message=FALSE}
## loading packages
library(knitr)
library(ChIPseeker)
library(clusterProfiler)
library(dplyr)
library(ggplot2)
library(UpSetR)
library(GenomicRanges)
library(readr)
```

```{r load}
baseDirSP = "/Users/lpantano/orch/scratch/chipseq_ma_shannan/re-analysis/sp140"
baseDir = "/Users/lpantano/orch/scratch/chipseq_ma_shannan/re-analysis/peaks_files"
path_results=file.path(baseDir, "filtered")
dir.create(path_results, showWarnings = FALSE)

files <- list(sp140_0h = c(file.path(baseDirSP, "SP140_0h_IP_a_peaks.broadPeak")), 
              sp140_4h = c(file.path(baseDirSP, "SP140_4h_IP_a_peaks.broadPeak")))
broad_header = c("chrom", "start", "end", "peak", "score", "fake", 
                 "foldchange", "log10pvalue", "log10qvalue")
narrow_header = c("chrom", "start", "end", "peak", "score", "fake", 
                 "foldchange", "log10pvalue", "log10qvalue", "summit")

peaks = list(
    sp140_0h=makeGRangesFromDataFrame(
        read_tsv(files[[1]], col_names=broad_header) %>%
            filter(foldchange > 0),
        keep.extra.columns = TRUE, 
        ignore.strand = TRUE, end.field = "end",
        start.field = "start", seqnames.field = "chrom"
    ),
    sp140_4h=makeGRangesFromDataFrame(
        read_tsv(files[[2]], col_names=broad_header) %>%
            filter(foldchange > 0),
        keep.extra.columns = TRUE, 
        ignore.strand = TRUE, end.field = "end",
        start.field = "start", seqnames.field = "chrom"
    )
)

k4me3 = read.table(file.path(baseDir, "H3K4me3_merged_peaks.scores.narrowPeak"), header=F,stringsAsFactors = F) %>% filter(!grepl("_",V1))
names(k4me3) = c("chrom", "start", "end", "score", "foldchange")
k4me3 = makeGRangesFromDataFrame(k4me3, keep.extra.columns = TRUE, 
                                    ignore.strand = TRUE, end.field = "end",
                                    start.field = "start", seqnames.field = "chrom")

k4me1 = read.table(file.path(baseDir, "H4K4me1_peaks.narrowPeak"), header=F,stringsAsFactors = F) %>% filter(!grepl("_",V1))
names(k4me1) = narrow_header
k4me1 = makeGRangesFromDataFrame(k4me1, keep.extra.columns = TRUE, 
                                    ignore.strand = TRUE, end.field = "end",
                                    start.field = "start", seqnames.field = "chrom")

k27ac = read.table(file.path(baseDir, "H3K27ac_peaks.narrowPeak"), header=F,stringsAsFactors = F) %>% filter(!grepl("_",V1))
names(k27ac) = narrow_header
k27ac = makeGRangesFromDataFrame(k27ac, keep.extra.columns = TRUE, 
                                    ignore.strand = TRUE, end.field = "end",
                                    start.field = "start", seqnames.field = "chrom")

k27me3 = read.table(file.path(baseDir, "H3K27me3_peaks.broadPeak"), header=F,stringsAsFactors = F)  %>% filter(!grepl("_",V1))
names(k27me3) = broad_header
k27me3 = makeGRangesFromDataFrame(k27me3, keep.extra.columns = TRUE, 
                                    ignore.strand = TRUE, end.field = "end",
                                    start.field = "start", seqnames.field = "chrom")

```

```{r raw-peaks}
basic_stat = data.frame(type=c("K27ac", "K27me3", "K4me3", "K4me1", "SP140_0h", "SP140_4h"),
           peaks=c(length(k27ac), length(k27me3), length(k4me3), length(k4me1), length(peaks$sp140_0h), length(peaks$sp140_4h)))
```


## Choosing the best cutoff

We used the same rule for the cutoff for each file. It is the value just after the
larger peak, that is considered the background peak value. Everything after that
is considered a strong peak.

```{r hist-peaks}
hist(mcols(k27ac)[,"foldchange"], breaks=seq(1,60,0.5), xlim = c(0,20))
abline(v=5, col="red")
hist(mcols(k27me3)[,"foldchange"], breaks=seq(1,15,0.5))
abline(v=3, col="red")
hist(mcols(k4me1)[,"foldchange"], breaks=seq(1,15,0.5))
abline(v=5, col="red")
hist(mcols(k4me3)[,"foldchange"], breaks=seq(1,51,0.5), xlim = c(0,20))
abline(v=6, col="red")
hist(mcols(peaks$sp140_0h)[,"foldchange"], breaks=seq(1,15,0.5))
abline(v=3.5, col="red")
hist(mcols(peaks$sp140_4h)[,"foldchange"], breaks=seq(1,15,0.5))
abline(v=3.5, col="red")
```

## Peaks after filtering

```{r negatives}
cutoff = list("k27me3"=3,"k4me1"=5,"k4me3"=6,"sp140_0h"=3.5,"sp140_4h"=3.5,"k27ac"=5)
save_file(as.data.frame(peaks$sp140_0h[mcols(peaks$sp140_0h)[,"foldchange"]<cutoff[["sp140_0h"]],]), "sp140_0h_removed.txt", path_results)
save_file(as.data.frame(peaks$sp140_4h[mcols(peaks$sp140_4h)[,"foldchange"]<cutoff[["sp140_4h"]],]), "sp140_4h_removed.txt", path_results)
save_file(as.data.frame(k27me3[mcols(k27me3)[,"foldchange"]<cutoff[["k27me3"]],]), "k27me3_removed.txt", path_results)
save_file(as.data.frame(k27ac[mcols(k27ac)[,"foldchange"]<cutoff[["k27ac"]],]), "k27ac_removed.txt", path_results)
save_file(as.data.frame(k4me3[mcols(k4me3)[,"foldchange"]<cutoff[["k4me3"]],]), "k4me3_removed.txt", path_results)
save_file(as.data.frame(k4me1[mcols(k4me1)[,"foldchange"]<cutoff[["k4me1"]],]), "k4me1_removed.txt", path_results)
```


```{r stats, results='asis'}
peaks$sp140_0h = peaks$sp140_0h[mcols(peaks$sp140_0h)[,"foldchange"]>cutoff[["sp140_0h"]],]
peaks$sp140_4h = peaks$sp140_4h[mcols(peaks$sp140_4h)[,"foldchange"]>cutoff[["sp140_4h"]],]
k27me3 = k27me3[mcols(k27me3)[,"foldchange"]>cutoff[["k27me3"]],]
k27ac = k27ac[mcols(k27ac)[,"foldchange"]>cutoff[["k27ac"]],]
k4me3 = k4me3[mcols(k4me3)[,"foldchange"]>cutoff[["k4me3"]],]
k4me1 = k4me1[mcols(k4me1)[,"foldchange"]>cutoff[["k4me1"]],]

sprintf("K27ac cutoff %s: %s peaks\n", cutoff[["k27ac"]], length(k27ac))
sprintf("K27me3 cutoff %s: %s peaks\n", cutoff[["k27me3"]], length(k27me3))
sprintf("K4me3 cutoff %s: %s peaks\n", cutoff[["k4me3"]], length(k4me3))
sprintf("K4me1 cutoff %s: %s peaks\n", cutoff[["k4me1"]], length(k4me1))
sprintf("SP140_0h cutoff %s: %s peaks\n", cutoff[["sp140_0h"]], length(peaks$sp140_0h))
sprintf("SP140_4h cutoff %s: %s peaks\n", cutoff[["sp140_4h"]], length(peaks$sp140_4h))
```

```{r stats-after, results='asis'}
basic_stat = cbind(basic_stat,
                   data.frame(
           cleaned_peaks=c(length(k27ac), length(k27me3), length(k4me3), length(k4me1), length(peaks$sp140_0h), length(peaks$sp140_4h))) )

knitr::kable(basic_stat)

```

## Standard annotation

```{r anno}
peak_rda <- file.path(baseDirSP, "peak.rda")
his_rda <- file.path(baseDirSP, "histones.rda")

if (!file.exists(peak_rda)){
    peakAnnoList <- lapply(peaks, annotatePeak, TxDb = txdb, tssRegion = c(-5000,5000), 
                           annoDb = "org.Hs.eg.db", verbose = FALSE)
    
    histAnnoList <- lapply(list(k27ac,k4me3,k27me3,k4me1), annotatePeak, 
                           TxDb = txdb, tssRegion = c(-5000,5000), 
                           annoDb = "org.Hs.eg.db", verbose = FALSE)
    names(histAnnoList) = c("k27ac", "k4me3", "k27me3", "k4me1")
    idx1 = !grepl("Promoter", mcols(histAnnoList[["k27ac"]]@anno)[,"annotation"])
    # idx2 = !grepl("Promoter", mcols(histAnnoList[["k4me1"]]@anno)[,"annotation"])
    # enhancer = reduce(sort(c(k27ac[idx1, ],k4me1[idx2,],ignore.mcols=TRUE)))
    enhancer = k27ac[idx1,]
    histAnnoList[["enhancer"]] =  annotatePeak(enhancer, TxDb = txdb, 
                                               tssRegion = c(-5000,5000), 
                                               annoDb = "org.Hs.eg.db", verbose = FALSE)
    save(peakAnnoList, file=peak_rda)
    save(histAnnoList, file=his_rda)
}else{
    load(peak_rda)
    load(his_rda)
}

idx = histAnnoList$k27me3@detailGenomicAnnotation$Promoter==TRUE
save_file(as.data.frame(histAnnoList$k27me3@anno[idx,]), "H3K27me3_tss.txt", path_results)
idx = histAnnoList$k4me3@detailGenomicAnnotation$Promoter==TRUE
save_file(as.data.frame(histAnnoList$k4me3@anno[idx,]), "H3K4me3_tss.txt", path_results)

save_file(as.data.frame(histAnnoList$k27ac@anno), "H3K27ac.txt", path_results)
save_file(as.data.frame(histAnnoList$k27me3@anno), "H3K27me3.txt", path_results)
save_file(as.data.frame(histAnnoList$k4me3@anno), "H3K4me3.txt", path_results)
save_file(as.data.frame(histAnnoList$k4me1@anno), "H3K4me1.txt", path_results)
save_file(as.data.frame(peakAnnoList$sp140_0h@anno), "SP140_0h.txt", path_results)
save_file(as.data.frame(peakAnnoList$sp140_4h@anno), "SP140_4h.txt", path_results)

```

```{r, eval=FALSE, echo=FALSE}
gene_k4me3 = mcols(histAnnoList$k4me3@anno)[,c("score","ENSEMBL","annotation")] %>% as.data.frame() %>% mutate(size=width(histAnnoList$k4me3@anno)) %>% dplyr::filter(grepl("Promoter",annotation)) %>% group_by(ENSEMBL) %>% summarise(averagek4=sum(score),k4=n(), sizek4=sum(size))

gene_k27me3 = mcols(histAnnoList$k27me3@anno)[,c("score","ENSEMBL","annotation")] %>% as.data.frame() %>% mutate(size=width(histAnnoList$k27me3@anno))%>% dplyr::filter(grepl("Promoter",annotation)) %>% group_by(ENSEMBL) %>% summarise(averagek27=sum(score),k27=n(), sizek27=sum(size))

gene_sp1400h = mcols(peakAnnoList$sp140_0h@anno)[,c("score","ENSEMBL","annotation")] %>% as.data.frame()  %>% mutate(size=width(peakAnnoList$sp140_0h@anno))%>% dplyr::filter(grepl("Promoter",annotation)) %>% group_by(ENSEMBL) %>% summarise(averagesp0h=sum(score),sp0h=n(), sizesp0h=sum(size))

gene_sp1404h = mcols(peakAnnoList$sp140_4h@anno)[,c("score","ENSEMBL","annotation")] %>% as.data.frame() %>% dplyr::filter(grepl("Promoter",annotation)) %>% group_by(ENSEMBL) %>% summarise(averagesp4h=sum(foldchange), size=sum(size))


df = left_join(gene_k4me3, gene_sp1400h)
df = df[order(df$averagek4, decreasing = T),]
df$rank = 1:nrow(df)
# df[is.na(df)] = 0
ggplot(df, aes(x=averagek4, y=averagesp0h)) + geom_point(alpha=.1) + scale_x_log10() + scale_y_log10()
ggplot(df, aes(x=averagek4/sizek4, y=averagesp0h/sizesp0h)) + geom_point(alpha=.1) + scale_x_log10() + scale_y_log10()
ggplot(df, aes(x=rank, y=averagesp0h/sp0h)) + geom_point(alpha=.3) + scale_x_log10() + scale_y_log10() + xlim(0,100)
ggplot(df, aes(x=sizek4*averagek4, y=sizesp0h*averagesp0h)) + geom_point(alpha=.3) + scale_x_log10() + scale_y_log10()
ggplot(df, aes(x=sizek4/rank, y=sizesp0h/averagesp4h)) + geom_point()


df = left_join(gene_k27me3, gene_sp1400h)
df = df[order(df$averagek27),]
df$rank = 1:nrow(df)
# df[is.na(df)] = 0
ggplot(df, aes(x=averagek27, y=averagesp0h)) + geom_point(alpha=.3) + scale_x_log10() + scale_y_log10()
ggplot(df, aes(x=rank, y=averagesp4h)) + geom_point(alpha=.3)


df = left_join(gene_k27me3, gene_k4me3)
df = df[order(df$averagek27),]
df$rank = 1:nrow(df)
ggplot(df, aes(x=averagek27, y=averagek4)) + geom_point(alpha=.3) + scale_x_log10() + scale_y_log10()


```


### Overlap with all histone marks

It is tricky to overlap 4 peaks of sets, because first you need to define an
unique set of peaks in the whole data set. For instance, it happens than two peaks overlap
with one peak in other data set. When comparing 4 sets, the problem is complex, and
my approach is to define an unique set of peaks, and then overlap this one with 
every other set of peaks.

To define an unique set of peaks, I put all peaks together and merge the ones that
overlap. This will cause that the total peaks in each data set, will differ slightly 
in the total peaks seen in the figure. For instance, it happens than one region has:

```
SP140_0h has 3 peaks
H3K27ac has 1 peak
H3K4me3 has 2 peaks
H3K4me1 has 0 peaks
```

This will end up as 1 unique peak, and will be counted only once for SP140_0h, H3K27ac
and H3K4me3.


```{r full-overlap, fig.width=10}

plot_common = function(all_peaks, sets){
    .void = lapply(names(sets), function(s){
        r = sets[[s]]
        idx = findOverlaps(all_peaks, r, maxgap = 501)
        mcols(all_peaks)[,s] <<- FALSE
        mcols(all_peaks)[queryHits(idx), s] <<- TRUE
    })
    ma = as.matrix(as.data.frame(mcols(all_peaks)[,names(sets)]))
    ma[ma] = 1
    UpSetR::upset(as.data.frame(ma),sets = names(sets))
}

.void = lapply(names(peakAnnoList), function(name){
    class_anno = peakAnnoList[[name]]
    range = slot(class_anno,"anno")
    
    all_peaks = reduce(sort(c(k4me1, k27ac, range, ignore.mcols=TRUE)))
    sets = list(sp140=range, H3K4me1=k4me1, H3K27ac=k27ac)
    names(sets)[1] = name
    plot_common(all_peaks, sets)
    
    all_peaks = reduce(sort(c(k4me3, k27ac, range, ignore.mcols=TRUE)))
    sets = list(sp140=range, H3K4me3=k4me3, H3K27ac=k27ac)
    names(sets)[1] = name
    plot_common(all_peaks, sets)
    

})

```

### Overlap only at TSS +/- 5Kb regions.

```{r tss-overlap, fig.width=10}
.void = lapply(names(peakAnnoList), function(name){
    class_anno = peakAnnoList[[name]]
    range = slot(class_anno,"anno")
    
    k4me3=slot(histAnnoList[["k4me3"]], "anno")
    k27me3=slot(histAnnoList[["k27me3"]], "anno")
    
    idx_sp140 = grepl("Promoter", mcols(range)[,"annotation"])
    idx_k4me3 = grepl("Promoter", mcols(k4me3)[,"annotation"])
    idx_k27me3 = grepl("Promoter", mcols(k27me3)[,"annotation"])
    
    all_peaks = reduce(sort(c(k27me3[idx_k27me3, ], k4me3[idx_k4me3, ], range[idx_sp140, ], ignore.mcols=TRUE)))
    sets = list(sp140= range[idx_sp140, ], H3K27me3=k27me3[idx_k27me3, ], H3K4me3=k4me3[idx_k4me3, ])
    names(sets)[1] = name
    plot_common(all_peaks, sets)
})
```


## New annotation with enhancer

```{r local_fn}

getGenomicAnnoStat <- function(peakAnno) {
    if ( class(peakAnno) == "GRanges" )
        peakAnno <- as.data.frame(peakAnno)
    anno <- peakAnno$annotation
    ## anno <- sub(" \\(.+", "", anno)

    anno[grep("exon 1 of", anno)] <- "1st Exon"
    anno[grep("intron 1 of", anno)] <- "1st Intron"
    anno[grep("Exon \\(", anno)] <- "Other Exon"
    anno[grep("Intron \\(", anno)] <- "Other Intron"
    anno[grep("Downstream", anno)] <- "Downstream (<=3kb)"
    
    ## count frequency
    anno.table <- table(anno)
    
    ## calculate ratio
    anno.ratio <- anno.table/ sum(anno.table) * 100
    anno.df <- as.data.frame(anno.ratio)
    colnames(anno.df) <- c("Feature", "Frequency")
    anno.df$Numbers <- anno.table
    lvs <- c(## "Promoter (<=1kb)",
             ## "Promoter (1-2kb)",
             ## "Promoter (2-3kb)",
             "Promoter (<= 3kb)",
             "Promoter",
             "Promoter-H3K27me3",
             "Enhancer",
             "H3K4me1",
             "Enhancer-H3K4me1",
             "5' UTR",
             "3' UTR",
             "1st Exon",
             "Other Exon",
             "1st Intron",
             "Other Intron",
             "Downstream (<=3kb)",
             "Distal Intergenic",
             "Others")

    anno.df$Feature <- factor(anno.df$Feature, levels=lvs[lvs %in% anno.df$Feature])
    anno.df <- anno.df[order(anno.df$Feature),]
    return(anno.df)
}

```

Promoter defined as the region at +/- 5Kb from TSS

Enhancer defined as peaks that overlap H3K27ac +  (+/- 5kb far aways from any TSS or gene)

```{r dist-top, eval=FALSE, echo=FALSE}
p = peakAnnoList[[1]]@anno
#p = p[width(p)>1000,]
#idx = order(p$foldchange, decreasing = T)
#p = p[1:round(length(idx)*0.1),]
pro = (grepl("Promoter",p$annotation))
plot(log10(width(p[pro,])),
     mcols(p[pro,])$foldchange)
hist(log10(width(p[pro,])))
hist(mcols(p[pro,])$foldchange)

dis = (grepl("Distal", p$annotation))
plot(log10(width(p[dis,])),
     mcols(p[dis,])$foldchange)
hist(log10(width(p[dis,])))
hist(mcols(p[dis,])$foldchange)
```


```{r top-annotation}
require(TxDb.Hsapiens.UCSC.hg19.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene

top10 = list()
idx = order(mcols(peaks[[1]])$foldchange*width(peaks[[1]]), decreasing = T)
top10[[1]] = peaks[[1]][idx[1:round(length(idx)/10)],]
idx = order(mcols(peaks[[2]])$foldchange*width(peaks[[2]]), decreasing = T)
top10[[2]] = peaks[[2]][idx[1:round(length(idx)/10)],]

topPeakAnnoList <- lapply(top10, annotatePeak, TxDb = txdb, tssRegion = c(-5000,5000),annoDb = "org.Hs.eg.db", verbose = FALSE)
names(topPeakAnnoList) = c("top_SP140_0h","top_SP140_4h")
```


```{r new-annotation, results='asis'}
all = c(peakAnnoList, topPeakAnnoList)
peakAnnoList_mod = lapply(names(all), function(name){
    class_anno = all[[name]]
    range = slot(class_anno,"anno")
    idx = findOverlaps(range[,"annotation"], k27ac, maxgap = 501)
    mcols(range)[,"enhancer"] = FALSE
    mcols(range)[queryHits(idx),"enhancer"] = TRUE
    
    idx = findOverlaps(range[,"annotation"], k4me1, maxgap = 501)
    mcols(range)[,"k4me1"] = FALSE
    mcols(range)[queryHits(idx),"k4me1"] = TRUE
    ann = mcols(range)
    ann_col = ann[,"annotation"]
    
    idx = findOverlaps(range[,"annotation"], k27me3, maxgap = 501)
    mcols(range)[,"k27me3"] = FALSE
    mcols(range)[queryHits(idx),"k27me3"] = TRUE
    ann = mcols(range)
    ann_col = ann[,"annotation"]

    
    new_col = ann_col
    # new_col[1:length(ann_col)] = "Others"
    new_col[grepl("Promoter", ann_col)] = "Promoter"
    new_col[ann[,"enhancer"] & new_col != "Promoter"] = "Enhancer"
    new_col[ann[,"k4me1"] & new_col != "Promoter"] = "H3K4me1"
    new_col[ann[,"enhancer"] & new_col == "H3K4me1"] = "Enhancer-H3K4me1"
    # new_col[ann[,"k27me3"] & new_col == "Promoter"] = "Promoter-H3K27me3"
    
    logic_df = cbind(slot(class_anno,
                          "detailGenomicAnnotation"),
                     Enhancer = new_col =="Enhancer")
    logic_df$Promoter = new_col == "Promoter"
    logic_df$Promoter_H3K27me3 = new_col == "Promoter-H3K27me3"
    logic_df$Enhancer = new_col == "Enhancer"
    logic_df$H3K4me1 = new_col == "H3K4me1"
    # logic_df$Enhancer_H3K4me1 = new_col == "Enhancer-H3K4me1"
    
    mcols(range)[,"annotation"] = new_col
    df = getGenomicAnnoStat(range)
    slot(class_anno, "anno") <- range
    slot(class_anno, "annoStat") <- df
    slot(class_anno, "detailGenomicAnnotation") <- logic_df
    cat("\n\n ###",name,"\n\n")
    print(kable(df))
    cat("\n\n")
    print(plotAnnoBar(class_anno))
    cat("\n\n")
    plotAnnoPie(class_anno)
    cat("\n\n")
    
    save_file(as.data.frame(range), paste0(name, "_ann.txt"), baseDirSP,csv = FALSE)
    pro_file_name = paste0(name,"_promoter.txt")
    save_file(as.data.frame(range[new_col=="Promoter",])[,1:12],
          pro_file_name, baseDirSP)
    enh_file_name = paste0(name,"_enhancer.txt")
    save_file(as.data.frame(range[new_col=="Enhancer",][,1:12]),
              enh_file_name, baseDirSP)
    class_anno
})
peakAnnoList_mod = peakAnnoList_mod[1:2]
names(peakAnnoList_mod) = names(peakAnnoList)
```

## Ranked plot

According to the paper, they plot the same protein/histone mark, for instance, 
it would be ranked SP140_4h against the signal of SP140_4h. 

Only for Promoters peaks for H3K27me3. And Promoters
for H3K4me3.

```{r ranked}

    # idx = findOverlaps(k27ac, k27ac)
    # df = data.frame(histone = mcols(k27ac)[,"foldchange"])
    # df$sp140 = 0
    # df$sp140[subjectHits(idx)] = mcols(k27ac)[queryHits(idx),"foldchange"]
    # plot((df[order(df$histone),"sp140"]), pch=20, 
    #      xlab = "ranked H3K27ac signal", ylab="score", main=name)

.void=lapply(names(peakAnnoList), function(name){
    class_anno = peakAnnoList_mod[[name]]
    range = slot(class_anno,"anno")
    # Only in enchancer
    idx_promoter=grep("Promoter", mcols(range)[,"annotation"])
    promoter_range = range[idx_promoter,]
    
    hist_range = slot(histAnnoList[['k27me3']], "anno")
    hist_keep = hist_range[grepl("Promoter", mcols(hist_range)[,"annotation"]),]
    idx = findOverlaps(promoter_range, hist_keep)
    df = data.frame(histone = mcols(hist_keep)[,"foldchange"])
    df$sp140 = 0
    df$sp140[subjectHits(idx)] = mcols(promoter_range)[queryHits(idx),4]
    plot((df[order(df$histone),"sp140"]), pch=20, 
         xlab = "ranked H3K27me3 signal", ylab="fold change enrichment", main=name)

    # Only in promoters
    idx_promoter=grep("Promoter", mcols(range)[,"annotation"])
    promoter_range = range[idx_promoter,]
    hist_range = slot(histAnnoList[['k4me3']], "anno")
    hist_keep = hist_range[grepl("Promoter", mcols(hist_range)[,"annotation"]),]
    idx = findOverlaps(promoter_range, hist_keep)
    df = data.frame(histone = mcols(hist_keep)[,"foldchange"])
    df$sp140 = 0
    df$sp140[subjectHits(idx)] = mcols(promoter_range)[queryHits(idx),4]
    df$rank = order(df$histone)
    #plot(density(df$rank[df$sp140>0]))
    plot((df[order(df$histone),"sp140"]), pch=20,
         xlab = "ranked H3K4me3 signal", ylab="fold change enrichment", main=name)

    
})
```


#R Session Info

(useful if replicating these results)

```{r sessioninfo}
sessionInfo()
```


