---
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
    theme: readable
    highlight: zenburn
---

The most recent update of this html document occurred: `r date()`


```{r setup}
library(knitr)
library(myRfunctions)
library(ggplot2)
library(reshape)
library(DESeq2)
library(genefilter)
library(CHBUtils)
library(gtools)
library(gridExtra)
library(devtools)
library(dplyr)
library(isomiRs)
library(pheatmap)

knitr::opts_chunk$set(tidy=TRUE, highlight=TRUE, dev="png", fig.width=6,fig.heigh=6,
               cache=FALSE, highlight=TRUE, autodep=TRUE, 
               warning=FALSE, error=FALSE,
               message=FALSE, prompt=TRUE, comment='', fig.cap='', bootstrap.show.code=FALSE)

root_path = "~/orch/scratch/kontardis_ipsc/srna-ipsc/work/upload"
root_file = file.path(root_path, "srna_out_files")
dir.create(root_file, showWarnings = FALSE)

metadata_fn =  list.files(file.path(root_path), pattern = "summary.csv$",recursive = T, full.names = T)
metadata = read.csv(metadata_fn, row.names="sample_id")
condition = names(metadata)[1]
design = metadata
metadata$group = c("mutant", "mutant", "control", "control", "mutant")
formula = ~ condition # modify this to get your own formula, it should be a column in your metadata
isde=FALSE # turn this true to make DE ananlysis
```

```{r create-report, echo=FALSE, eval=FALSE}
# render_2_drop("smrna-summary.Rmd", "Kontaridis/ipsc-smallrna")
```

# Exploratory analysis

In this section we will see descriptive figures about quality of the data, 
reads with adapter, reads mapped to miRNAs, reads mapped to other small RNAs. 

## size distribution

After adapter removal, we can plot the size distribution of the small RNAs.

```{r adapter-check,fig.width=10}
files = list.files(file.path(root_path),pattern = "trimming_stats",recursive = T)
isadapter = length(files)>0
```

```{r adapter,fig.width=10,eval=isadapter}
names(files) = sapply(files, function(x){
  gsub("-ready.trimming_stats", "", basename(x))
})


tab = data.frame()
for (sample in rownames(metadata)) {
    d = read.table(file.path(root_path, files[sample]), sep=" ")
    tab = rbind(tab, d %>% mutate(sample=sample, group=metadata[sample, condition]))
}


reads_adapter = tab %>% group_by(sample, group) %>% summarise(total=sum(V2))
ggplot(reads_adapter, aes(x=sample,y=total,fill=group)) +
    geom_bar(stat="identity", position = "dodge") +
    ggtitle("total number of reads with adapter") +
    ylab("# reads") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 


ggplot(tab, aes(x=V1,y=V2, group=sample)) +
    geom_bar(stat="identity", position = "dodge") +
    facet_wrap(~group, ncol=2)+
    ggtitle("size distribution") +
    ylab("# reads") + xlab("size") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    facet_wrap(~sample)
```

```{r mirna-load}
files = list.files(file.path(root_path),pattern = "mirbase-ready",recursive = T,full.names = T)
ismirbase = length(files) > 0
mirdeep2_files = list.files(file.path(root_path),pattern = "novel-ready",recursive = T,full.names = T)
ismirdeep2 = length(mirdeep2_files) > 0
```

## miRNA

### total miRNA expression annotated with mirbase

```{r mirna,results='hide', eval=ismirbase}
names(files) = sapply(files, function(x){
  gsub("-mirbase-ready.counts", "", basename(x))
})

obj <- IsomirDataSeqFromFiles(files = files[rownames(design)], design = design , header = T, skip=0)
```


```{r mirna-mirbase, eval=ismirbase}
ggplot( data.frame(sample=colnames(counts(obj)), total=colSums(counts(obj)))) +
    geom_bar(aes(x=sample,y=total), stat='identity')+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
mirna_step <- as.data.frame(colSums(counts(obj)))
```

### Distribution of mirna expression

```{r depth, eval=ismirbase}
ggplot(melt(counts(obj))) +
    geom_boxplot(aes(x=X2,y=value))+
    scale_y_log10()+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### cumulative distribution of miRNAs

```{r cum, eval=ismirbase}
cs <- as.data.frame(apply(counts(obj),2,function(x){cumsum(sort(x, decreasing = T))}))
cs$pos <- 1:nrow(cs)

ggplot((melt(cs,id.vars = "pos")))+
    geom_line(aes(x=pos,y=value,color=variable))+
    scale_y_log10()
```

### Clustering
```{r mirna-clustering, eval=ismirbase}
counts = counts(obj)
dds = DESeqDataSetFromMatrix(counts[rowSums(counts>0)>3,], colData=design, design=~1)
vst = rlog(dds)
pheatmap(assay(vst), annotation_col = metadata, show_rownames = F, 
         clustering_distance_cols = "correlation",
         clustering_method = "ward.D")
```

### MDS plot
```{r pca, eval=ncol(counts) > 1, eval=ismirbase}
mds(assay(vst), condition = metadata[,"group"], k=3)
```

### complexity

Number of miRNAs with > 3 counts.

```{r complexity, eval=ismirbase}

kable(as.data.frame(colSums(counts>10)))

```



## Others small RNA

The data was analyzed with [seqcluster](http://seqcluster.readthedocs.org/)

This tools used all reads, uniquely mapped and multi-mapped reads. The first
step is to cluster sequences in all locations they overlap. The second step is to 
create meta-clusters: is the unit that merge all clusters that share the 
same sequences. This way the output are meta-clusters, common sequences that could
come from different region of the genome.


### genome covered

In this table 1 means % of the genome with at least 1 read, and 0 means %
of the genome without reads.

```{r genome-covered, results='asis'}
fn_json = (list.files(file.path(root_path),pattern = "seqcluster.json",recursive = T,full.names = T))[1]
seq_dir = dirname(fn_json)

isseqcluster = length(fn_json)>0
#cov_stats <- read.table(file.path(root_path, "..", "align", "seqs_rmlw.bam_cov.tsv"),sep="\t",check.names = F)

#kable(cov_stats[cov_stats$V1=="genome",] %>%  dplyr::select(coverage=V2,ratio_genome=V5), row.names = FALSE)

```

The normal value for human data with strong small RNA signal is: 0.0002.
This will change for smaller genomes.

### classification

Number of reads in the data after each step:

* raw: initial reads
* cluster: after cluster detection
* multimap: after meta-cluster detection

```{r reads-track,eval=isseqcluster}
reads_stats <- read.table(file.path(seq_dir, "read_stats.tsv"),sep="\t",check.names = F)
ggplot(reads_stats, aes(x=V2, y=V1, fill=V3)) + 
    geom_bar(stat = 'identity', position = 'dodge') +
    labs(list(x="samples", y="reads")) +
    scale_fill_brewer("steps", palette = 'Set1')+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


```{r cluster,eval=isseqcluster}

clus <- read.table(file.path(seq_dir, "counts.tsv"),header=T,sep="\t",row.names=1, check.names = FALSE)
ann <- clus[,2]
toomany <- clus[,1]
clus_ma <- clus[,3:ncol(clus)]
clus_ma = clus_ma[,row.names(design)]

```


Check complex meta-clusters: This kind of events happen when there are small RNA over the whole genome, and all
repetitive small rnas map to thousands of places and sharing many sequences in many positions.
If any meta-cluster is > 40% of the total data, maybe it is worth to add some filters
like: minimum number of counts `-e` or `--min--shared` in `seqcluster prepare`


```{r complex-meta-clusters,eval=isseqcluster}
library(edgeR)
clus_ma_norm = cpm(DGEList(clus_ma),normalized.lib.sizes = TRUE)
head(clus_ma_norm[toomany>0,])
```

### complexity

Number of miRNAs with > 10 counts.

```{r complexity-clus,eval=isseqcluster}

kable(as.data.frame(colSums(clus_ma>10)))

```

### Contribution by class

```{r cluster_type,eval=isseqcluster}
rRNA <- colSums(clus_ma[grepl("rRNA",ann) & grepl("miRNA",ann)==F,])
miRNA <- colSums(clus_ma[grepl("miRNA",ann),])
tRNA <- colSums(clus_ma[grepl("tRNA",ann) & grepl("rRNA",ann)==F & grepl("ncRNA",ann)==F & grepl("miRNA",ann)==F,])
rmsk <- colSums(clus_ma[grepl("ncRNA",ann) & grepl("rRNA",ann)==F & grepl("miRNA",ann)==F,])
total <- colSums(clus_ma)

dd <- data.frame(samples=names(rRNA),
                 rRNA=rRNA,
                 miRNA=miRNA,
                 tRNA=tRNA,
                 ncRNA=rmsk,
                total=total)
ggplot(melt(dd)) +
    geom_bar(aes(x=samples,y=value,fill=variable),
             stat='identity',
             position="dodge")+
    scale_fill_brewer(palette = "Set1")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

dd_norm = dd
dd_norm[,2:5] = sweep(dd[,2:5],1,dd[,6],"/")
ggplot(melt(dd_norm[,1:5])) +
    geom_bar(aes(x=samples,y=value,fill=variable),
             stat='identity',
             position="dodge")+
    scale_fill_brewer(palette = "Set1")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
    labs(list(title="relative proportion of small RNAs",y="% reads"))
```

```{r clus-size-by-class}
# size_clus <- read.table(file.path(root_path, "..", "seqcluster", "cluster", "size_counts.tsv"),sep="\t",check.names = F)
```


# R Session Info

(useful if replicating these results)

```{r sessioninfo}
sessionInfo()
```
