---
title: "Proteomic analysis"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    theme: readable
    highlight: zenburn
---
 
The most recent update of this html document occurred: `r date()`

```{r opts, echo=FALSE}
knitr::opts_chunk$set(tidy=TRUE, highlight=TRUE, dev="png", fig.width=6,fig.heigh=6,
               cache=TRUE, highlight=TRUE, autodep=TRUE, 
               warning=FALSE, error=FALSE,
               message=FALSE, prompt=TRUE, comment='', fig.cap='', bootstrap.show.code=FALSE)
```

```{r setup}
library(knitr)
library(ggplot2)
library(reshape)
library(DESeq2)
library(CHBUtils)
library(limma)
library(gtools)
library(gridExtra)
library(devtools)
library(dplyr)
library(tidyr)
library(pheatmap)
library(rio)
```

```{r load-files}
exp1 = import("Experiment1_Data.xlsx")
exp1$Gene_Name = gsub("\\'", "", exp1$Gene_Name)
colnames(exp1) = gsub("\\+", "p", colnames(exp1))
exp2 = import("Experiment2_Data.xlsx")
exp2$Gene_Name = gsub("\\'", "", exp2$Gene_Name)
colnames(exp2) = gsub("\\+", "p", colnames(exp2))

data = inner_join(exp1[,1:11] %>% filter(Gene_Name != ""), 
                  exp2[,1:11] %>% filter(Gene_Name != ""), by="Gene_Name") %>%
    select(Gene_Name, Entrez_Gene_ID=Entrez_Gene_ID.x, Description=Description.x,
           d1_nt=nt.x, d1_sh17=sh17.x, d1_sh16=sh16.x,
           d1_ntp=ntp.x, d1_sh17p=sh17p.x, d1_sh16p=sh16p.x,
           d2_nt=nt.y, d2_sh17=sh17.y, d2_sh16=sh16.y,
           d2_ntp=ntp.y, d2_sh17p=sh17p.y, d2_sh16p=sh16p.y)

counts = data[,4:15]
row.names(counts) = data$Gene_Name


metadata = rbind(data.frame(donor = "donor1", group = colnames(exp1)[3:8]) %>%
                     mutate(name=paste0("d1_", group)),
                 data.frame(donor = "donor2", group = colnames(exp2)[3:8])  %>%
                     mutate(name=paste0("d2_", group))) %>% 
                     mutate(treat = ifelse(grepl("p$", group), "plus", "minus")) %>%
                     mutate(type = gsub("p$", "", group))
rownames(metadata) = metadata$name
metadata = metadata[colnames(counts),]
metadata


```

## Distribution of expression values

```{r description}

counts = counts[rowSums(counts[,1:6]>0)>2 & rowSums(counts[,6:12]>0)>2,]

log2counts = log2(counts+1)

ggplot(melt(log2counts), aes(x=value, color=variable)) +
    geom_density() + theme_bw()
mds(log2counts, k=2, condition = metadata$donor) + theme_bw()
```
## Differential analysis with DESeq2

Just using DESeq2 without any further addition to the analysis

## Dispersion values
```{r dispersion}
dds = DESeqDataSetFromMatrix(counts, metadata, design = ~ donor + treat + type)
# rdds = rlog(dds)
dds = DESeq(dds)
DESeq2::plotDispEsts(dds)
```

```{r simple-de}
sf <- sizeFactors(dds)
disp <- dispersions(dds)
dds_group = DESeqDataSetFromMatrix(counts, metadata, design = ~ donor + group)
sizeFactors(dds_group) <- sf
dispersions(dds_group) <- disp
dds_group <- nbinomWaldTest(dds_group)
```

How the gene SET looks like:
```{r set-gene}
res_sh16_vs_control = results(dds_group, list("groupsh16", "groupnt"))
suppressWarnings(DEGreport::degPlot(dds_group, res_sh16_vs_control["SET",,drop=FALSE] , n=1, xs="treat", group = "type", batch = "donor" ))
```

### WT treatment effect

* nt vs nt+

```{r deseq2-nt-alone, results='asis'}
res_nt_vs_plus = results(dds_group, list("groupntp", "groupnt"))
DESeq2::plotMA(res_nt_vs_plus)
res =  res_nt_vs_plus %>% as.data.frame() %>% tibble::rownames_to_column("id") %>%
    arrange(padj) %>% tibble::column_to_rownames("id")

res %>% head(15) %>% knitr::kable()

volcano_density_plot(res[,c("log2FoldChange", "pvalue")], shade.colour = "orange")

suppressWarnings(DEGreport::degPlot(dds_group, res , n=9, xs = "treat", group = "type", batch = "donor" ))

deseq2_ntplus_effect = res

```

### sh16 treatment effect

* sh16 vs sh16+

```{r deseq2-sh16-alone, results='asis'}
res_sh16_vs_plus = results(dds_group, list("groupsh16p", "groupsh16"))
DESeq2::plotMA(res_sh16_vs_plus)
res =  res_sh16_vs_plus %>% as.data.frame() %>% tibble::rownames_to_column("id") %>%
    arrange(padj) %>% tibble::column_to_rownames("id")

res %>% head(15) %>% knitr::kable()

volcano_density_plot(res[,c("log2FoldChange", "pvalue")], shade.colour = "orange")

suppressWarnings(DEGreport::degPlot(dds_group, res , n=9, xs = "treat", group = "type", batch = "donor" ))

deseq2_sh16plus_effect = res

```

### sh17 treatment effect

* sh17 vs sh17+

```{r deseq2-sh117-alone, results='asis'}
res_sh17_vs_plus = results(dds_group, list("groupsh17p", "groupsh17"))
DESeq2::plotMA(res_sh17_vs_plus)
res =  res_sh17_vs_plus %>% as.data.frame() %>% tibble::rownames_to_column("id") %>%
    arrange(padj) %>% tibble::column_to_rownames("id")

res %>% head(15) %>% knitr::kable()

volcano_density_plot(res[,c("log2FoldChange", "pvalue")], shade.colour = "orange")

suppressWarnings(DEGreport::degPlot(dds_group, res , n=9, xs = "treat", group = "type", batch = "donor" ))

deseq2_sh17plus_effect = res

```


### Proteins induced by SET in WT but not in SET-KD

The idea is to use the DE genes from nt vs nt+ and remove the ones that changed
due to sh6 vs sh16+ or sh17 vs sh17+.

We used the FDR < 0.05 to get the DE and pvalue>0.2 to get the non-changed genes.

```{r deseq2-dependet-nt}
deseq2_sh_dependent =
    inner_join(deseq2_ntplus_effect 
           %>% tibble::rownames_to_column("id") %>% filter(padj<0.05) %>%
               select(id,nt_padj=padj),
           inner_join(
           deseq2_sh16plus_effect  %>% 
               tibble::rownames_to_column("id") %>% filter(pvalue>0.2) %>%
               select(id,sh16_padj=pvalue),
           deseq2_sh17plus_effect  %>% 
               tibble::rownames_to_column("id") %>% filter(pvalue>0.2) %>%
               select(id,sh17_padj=pvalue),
           by="id"),
           by ="id") %>% rowwise() %>% 
    mutate(sh_max = max(sh16_padj, sh17_padj)) %>%
    mutate(diff = nt_padj - sh_max) %>% arrange((nt_padj))

suppressWarnings(DEGreport::degPlot(dds_group, deseq2_ntplus_effect[deseq2_sh_dependent$id,], n=9, xs = "treat", group = "type", batch = "donor" ))


```

```{r dependet-nt, eval=FALSE, echo=FALSE}
### Proteins induced by SET in WT but not in KD, complex model

dds_depend = DESeqDataSetFromMatrix(counts, metadata, design = ~ donor + type + treat)
d = colData(dds_depend)
d$type = relevel(d$type, "sh17")
full = model.matrix(~ donor + type + treat + type:treat, d)
reduce = full[,-6]
dds_depend = DESeq(dds_depend, test = "LRT", reduced=reduce, full=full)

deseq2_res_nt_depend = results(dds_depend)

DESeq2::plotMA(deseq2_res_nt_depend)

res =  deseq2_res_nt_depend %>% as.data.frame() %>% 
    tibble::rownames_to_column("id") %>%
    arrange(pvalue) %>% tibble::column_to_rownames("id")

res %>% head(15) %>% knitr::kable()

suppressWarnings(DEGreport::degPlot(dds_group, res, n=9, xs = "treat", group = "type", batch = "donor" ))


```

```{r sh16-treatment, results='asis', eval=FALSE, echo=FALSE}
### Different in treatments with sh16
keep = !grepl("sh17", row.names(metadata))
dds_complex = DESeqDataSetFromMatrix(counts[,keep], metadata[keep,], design = ~ donor + type + treat +  type*treat)
dds_complex = DESeq(dds_complex)
# sizeFactors(dds_complex) <- sf[keep]
# dispersions(dds_complex) <- disp
# dds_complex <- nbinomWaldTest(dds_complex)
dds_complex <- DESeq2::nbinomLRT(dds_complex, reduced = ~ donor + type + treat)

res_sh16_treatment = results(dds_complex)

DESeq2::plotMA(res_sh16_treatment)

res =  res_sh16_treatment %>% as.data.frame() %>% tibble::rownames_to_column("id") %>%
    arrange(pvalue) %>% tibble::column_to_rownames("id")

res %>% head(15) %>% knitr::kable()

volcano_density_plot(res[,c("log2FoldChange", "padj")], shade.colour = "orange")

suppressWarnings(DEGreport::degPlot(dds_complex, res, n=6, xs = "treat", group = "type", batch = "donor" ))

```

## Differential analysis with lima-voom as microarray data

```{r limma-fn}
d = model.matrix(~ 0 + donor + treat,  metadata)
y =normalizeBetweenArrays(log2(counts), 
                            method = "quantile")

limma_ma = voomaByGroup(y, group = metadata$type,
                 d, plot=FALSE)

limmaPlot = function (counts, genes, metadata, xs = "time", group = "condition", 
    batch = NULL) 
{
    pp = lapply(genes, function(gene) {
        dd = data.frame(count = counts[gene,], time=metadata[,xs])
        if (is.null(group)) {
            dd$treatment = "one_group"
        }
        else {
            dd$treatment = metadata[row.names(dd), group]
        }
        if (!is.null(batch)) {
            dd$batch = metadata[row.names(dd), batch]
            p = ggplot(dd, aes(x = time, y = count, color = batch, 
                shape = treatment))
        }
        else {
            p = ggplot(dd, aes(x = time, y = count, color = treatment, 
                shape = treatment))
        }
        p = p + stat_smooth(aes(x = time, y = count, group = treatment, 
            color = treatment), fill = "grey80") + geom_jitter(size = 1, 
            alpha = 0.7, height = 0, width = 0.2) + theme_bw(base_size = 7) + 
            ggtitle(gene)
        if (length(unique(dd$treatment)) == 1) {
            p = p + scale_color_brewer(guide = FALSE, palette = "Set1") + 
                scale_fill_brewer(guide = FALSE, palette = "Set1")
        }
        p
    })
    n = ceiling(length(pp))
    do.call(grid.arrange, pp)
}
```


### WT treatment effect

```{r lima-voom-array-nt, results='asis'}
library(limma)
keep = grepl("nt", row.names(metadata))
d = model.matrix(~ 0 + donor + treat,  metadata[keep,])
y =normalizeBetweenArrays(log2(counts[,keep]), 
                            method = "quantile")

v = voomaByGroup(y, group = metadata$type[keep],
                 d, plot=TRUE)

f = lmFit(v, d)
fb = eBayes(f)


res = topTable(fb,coef = "treatplus", sort.by = "p", number = Inf)

knitr::kable(head(res, 15))

volcano_density_plot(res[,c("logFC", "P.Value")], shade.colour = "orange")

suppressWarnings(DEGreport::degPlot(dds_group, res, n=9, xs = "treat", group = "type", batch = "donor" ))

ntplus_effect = res
```

### sh16/sh17 treatment effect

```{r lima-voom-array-sh, results='asis'}
library(limma)
keep = !grepl("nt", row.names(metadata))
d = model.matrix(~ 0 + donor + treat + type,  metadata[keep,])
y =normalizeBetweenArrays(log2(counts[,keep]), 
                            method = "quantile")

v = voomaByGroup(y, group = metadata$type[keep],
                 d, plot=TRUE)

f = lmFit(v, d)
fb = eBayes(f)

res = topTable(fb,coef = "treatplus", sort.by = "p", number = Inf)

knitr::kable(head(res, 15))

volcano_density_plot(res[,c("logFC", "P.Value")], shade.colour = "orange")

suppressWarnings(DEGreport::degPlot(dds_group, res, n=9, xs = "treat", group = "type", batch = "donor" ))

shplus_effect = res
```

```{r lima-voom-array-sh16, eval=FALSE, echo=FALSE}
### sh16 vs control over treatment
library(limma)
keep = !grepl("sh17", row.names(metadata))
d = model.matrix(~ 0 + donor + type + treat + type*treat, metadata[keep,])
y =normalizeBetweenArrays(log2(counts[,keep]), 
                            method = "quantile")

v = voomaByGroup(y, group = metadata$type[keep],
                 d, plot=TRUE)

# cont.matrix <- makeContrasts("typesh16-treatplus",levels=d)
f = lmFit(v, d)
# fc = contrasts.fit(f, cont.matrix)
fb = eBayes(f)
# topTable(fb, coef="typesh16")
# topTable(fb, coef="typesh16:treatplus", sort.by = "p")
res = topTable(fb,coef = "typesh16:treatplus", sort.by = "p", number = Inf)

knitr::kable(head(res, 15))

volcano_density_plot(res[,c("logFC", "P.Value")], shade.colour = "orange")

suppressWarnings(DEGreport::degPlot(dds_group, res, n=9, xs = "treat", group = "type", batch = "donor" ))

sh16_vs_nt_over_plus_effect = res

```

```{r lima-voom-array-sh17, results='asis', eval=FALSE, echo=FALSE}
### sh17 vs control over treatment
library(limma)
keep = !grepl("sh16", row.names(metadata))
d = model.matrix(~ 0 + donor + type + treat + type*treat, metadata[keep,])
y =normalizeBetweenArrays(log2(counts[,keep]), 
                            method = "quantile")

v = voomaByGroup(y, group = metadata$type[keep],
                 d, plot=TRUE)

f = lmFit(v, d)
fb = eBayes(f)
res = topTable(fb,coef = "typesh17:treatplus", sort.by = "p", number = Inf)

knitr::kable(head(res, 15))

volcano_density_plot(res[,c("logFC", "P.Value")], shade.colour = "orange")

suppressWarnings(DEGreport::degPlot(dds_group, res, n=9, xs = "treat", group = "type", batch = "donor" ))

sh17_vs_nt_over_plus_effect = res

```


### Proteins induced by SET in WT but not in KD
```{r vooma-dependet-nt}
sh_dependent =
    inner_join(ntplus_effect 
           %>% tibble::rownames_to_column("id") %>% filter(adj.P.Val<0.1),
           shplus_effect  %>% 
               tibble::rownames_to_column("id") %>% filter(adj.P.Val>0.2),
           by ="id")

suppressWarnings(DEGreport::degPlot(dds_group, ntplus_effect[sh_dependent$id,], n=8, xs = "treat", group = "type", batch = "donor" ))

# suppressWarnings(limmaPlot(limma_ma$E, sh_dependent$id, metadata, xs = "treat", group = "type", batch = "donor" ))

```

## Normalize with RUVseq

It doesn't help a lot this time. I guess too few replicates to remove noise from any of them.

```{r ruvseq}
library(RUVSeq)
ruv = RUVs(counts(dds_group), cIdx=row.names(counts), k=1, scIdx = matrix(c(1:6, 7:12), nrow=6))
mds(ruv$normalizedCounts, k=2, metadata$group) + ggtitle("1 factor")
ruv = RUVs(as.matrix(counts), cIdx=row.names(counts), k=2, scIdx = matrix(c(1:6, 7:12), nrow=6))
mds(ruv$normalizedCounts, k=2, metadata$group) + ggtitle("2 factors")
ruv = RUVs(as.matrix(counts), cIdx=row.names(counts), k=3, scIdx = matrix(c(1:6, 7:12), nrow=6))
mds(ruv$normalizedCounts, k=2, metadata$group) + ggtitle("3 factors")

```


