---
output:
  knitrBootstrap::bootstrap_document:
    title: "Feinberg RNA-seq"
    theme: readable
    highlight: zenburn
    theme.chooser: FALSE
    highlight.chooser: FALSE


---

```{r projsetup, echo=FALSE, warning=FALSE, message=FALSE}
project="Feinberg - RNAseq"
labPI="Mark W. Feinberg"
lablocation="BWH"
analystname="Lorena Pantano"
analystemail="lpantano@hsph.harvard.edu"

```


```{r knitrsetup, echo=FALSE}
knitr::opts_chunk$set(tidy=TRUE, highlight=TRUE, dev="png", bootstrap.show.code=FALSE, bootstrap.show.output=TRUE, bootstrap.thumbnail.size="col-md-10", cache=TRUE, highlight=TRUE, autodep=TRUE, warning=FALSE, error=FALSE, message=FALSE, prompt=TRUE, comment='', fig.cap='', tidy.opts=list(keep.blank.line=FALSE, width.cutoff=200), fig.width = 8, fig.height = 8)
options(bitmapType = 'cairo')
```


```{r render,echo=FALSE,eval=FALSE}
library(knitrBootstrap)
library(rmarkdown)
source("~/repos/myRfunctions/transactions.R")
# render_2_drop("feinberg-report.Rmd", "feinberg/rnaseq_project_3")
```

---

# Feinberg RNA-seq

RNA-seq analysis for`r labPI` group at `r lablocation`.  

Contact `r analystname` (`r analystemail`) for additional details.

The most recent update of this html document occurred: `r date()`

The sections below provide code to reproduce the included results and plots. 

---

# Setup

## Libraries

```{r functions}
library(ggplot2)
library(reshape)
library(gplots)
library(edgeR)
devtools::load_all("~/repos/CHBUtils")
library(pheatmap)
library(pcaMethods)
library(gridExtra)
library(knitr)
library(biomaRt)
library(rmarkdown)
library(dplyr)
library(magrittr)
library(tidyr)
library(org.Hs.eg.db)
library(xlsx)
library(DESeq2)
library(knitr)
library(clusterProfiler)
source("~/repos/myRfunctions/transactions.R")
```

## Functions

```{r heatmap-function}

plot.pca <- function (pcadata, condition= NULL, condition2= "", type="score", xi=1, yi=2, text=TRUE, title=""){
    require(pcaMethods)
  k = dim( pcadata)[3]
    xnames = paste0("PC", 1:k, " ", round(pcadata@R2 * 100,  digits = 1), "%")
  if (type == "loadings"){
     df <- loadings(pcadata)
	} else {
	   df <- scores(pcadata)
	}
    df = as.data.frame(df[, c(xi, yi)])
    names(df) = c("one", "two")
	df$label = rownames(df)
	if (text){
	    if (!is.null(condition)) {
	        df$condition = condition
	        p = ggplot(df, aes(one, two, label = label, color = condition)) + 
	            geom_text(aes(one, two, label = label), size = 3) + 
	            labs(list(x = xnames[xi], y = xnames[yi])) + 
				      scale_x_continuous() +
            ggtitle(title) +
            theme(plot.title = element_text(size=20, face="bold", vjust=2)) 

	    }
	    else {
	        p = ggplot(df, aes(one, two)) + 
			    geom_text(aes(one, two, label = label), size = 3) + 
				labs(list(x = xnames[xi], y = xnames[yi])) + 
				scale_x_continuous() +
            ggtitle(title) +
            theme(plot.title = element_text(size=20, face="bold", vjust=2))           
	    }
	} else {
	    if (!is.null(condition)) {
	        df$condition = condition
	        df$condition2 = condition2	        
	        p = ggplot(df, aes(one, two, label = label, color = condition, shape = condition2)) + 
	            geom_point(aes(one, two, label = label), size = 4) + 
	            labs(list(x = xnames[xi], y = xnames[yi])) + 
            ggtitle(title) +
            theme(plot.title = element_text(size=20, face="bold", vjust=2)) +            
				scale_x_continuous()
	    }
	    else {
	        p = ggplot(df, aes(one, two)) + 
			    geom_point(aes(one, two, label = label), size = 4) + 
				labs(list(x = xnames[xi], y = xnames[yi])) + 
            ggtitle(title) +
            theme(plot.title = element_text(size=20, face="bold", vjust=2)) +              
				scale_x_continuous()
	    }	
	}
    return(p)
}


convertIDs <- function( ids, from, to, db, ifMultiple=c("putNA", "useFirst")) {
  stopifnot( inherits( db, "AnnotationDb" ) )
  ifMultiple <- match.arg( ifMultiple )
  suppressWarnings( selRes <- AnnotationDbi::select(
    db, keys=ids, keytype=from, columns=c(from,to) ) )
  if ( ifMultiple == "putNA" ) {
    duplicatedIds <- selRes[ duplicated( selRes[,1] ), 1 ]
    selRes <- selRes[ ! selRes[,1] %in% duplicatedIds, ]
  }
  return( selRes[ match( ids, selRes[,1] ), 2 ] )
}

```

## Setup

```{r qc-setup}

dataDir <- file.path("data")
# metaDir <- file.path(baseDir, "meta")
# resultsDir <- file.path(baseDir, "results")

#Used files
rimagefile <- "~/orch/scratch/feinberg_rnaseq/project_3/10033684.1_edgeR/10033684.1.edgeR.Rdata"
# edgerfile <- "data/10031812.1_edgeR/10027561.1group1%group2.toptags_edgeR.csv"
annogtf <- "~/orch/scratch/feinberg_rnaseq/project_3/10033684.1.GRCh38.combined.gtf"
# output files
res_files <- "~/orch/scratch/feinberg_rnaseq/project_3/files"
dir.create(res_files, showWarnings = FALSE)

# Load old image
load(rimagefile)

#Get count data
counts.org <- counts
metadata <- samples[,2:1]
rownames(metadata) = samples[,2]
#Set cut-offs
count.cutoff <- 10
p.cutoff <- 0.01
fc.cutoff <- 1

```


```{r idconversion}
#Getting conversion betwen xloc and ensembl
anno.conv <- read.table(annogtf, header=FALSE, sep="\t")
tmp <- strsplit(as.character(anno.conv[,9]), "; ")
tmp <- matrix(unlist(lapply(tmp, function(x) x[c(1,5)])), ncol=2, byrow = T )
tmp[,1] <- gsub("gene_id ", "", tmp[,1])
tmp[,2] <- gsub("oId ", "", tmp[,2])
anno.conv <- tmp
rm(tmp)

colnames(anno.conv) <- c("xloc", "ensembl")
anno.conv <- anno.conv[!duplicated(anno.conv[,1]),]
idx <- match(gsub("\"" , "", rownames(counts)), anno.conv[,1])
rownames(counts) <- anno.conv[idx,2]
```


This analysis is based on the R image file (`r rimagefile`), annotations from gtf-file (`r annogtf`).

# Quality control metrics

## Counts per gene

```{r boxplot-raw-new}
melted = melt(counts)
#colnames(melted) = c("sample", "count")
colnames(melted) = c("gene", "sample", "count")
melted$sample = factor(melted$sample, colnames(counts))
#melted$sample = reorder(melted$sample, colnames(counts))
melted$count = log(melted$count)
c1 <- ggplot(melted, aes(x=sample, y=count)) + geom_boxplot() +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90), plot.title = element_text(size=20, face="bold", vjust=2)) + 
    labs(y="log10 counts per gene", x="", title="log10 counts per gene")  
```

```{r boxplot-normalized-new}
y = DGEList(counts=counts)
y = calcNormFactors(y)
normalized_counts = cpm(y, normalized.lib.sizes=TRUE)
melted = melt(normalized_counts)
colnames(melted) = c("gene", "sample", "count")
melted$sample = factor(melted$sample, colnames(counts))
#melted$sample = reorder(melted$sample, colnames(counts))
melted$count = log(melted$count)
c2 <- ggplot(melted, aes(x=sample, y=count)) + geom_boxplot() +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90), plot.title = element_text(size=20, face="bold", vjust=2)) + 
    labs(y="log10 TMM-normalized counts per gene", x="", title="log10 TMM-normalized counts per gene")  
```
 
```{r density-normalized-new}
c3 <- ggplot(melted, aes(x=count, group=sample)) +
    geom_density() +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90), plot.title = element_text(size=20, face="bold", vjust=2)) + 
    labs(y="density", x="", title="Density of log10 TMM-normalized counts") 
```

```{r qc-plots2-new , fig.width = 16, fig.height = 16}
grid.arrange( c1, c2, c3, ncol=2)
```

## Correlation (Spearman) heatmap of TMM-normalized counts
```{r spearman-heatmap-normalized-new, fig.width = 8, fig.height = 8}
pheatmap(cor(normalized_counts, method="spearman"), annotation = metadata[,"condition",drop=F],main="Spearman correlation", clustering_method = "ward.D2")
```

## PCA plot of TMM-normalized log2 counts
```{r pca-normalized-log2-new, fig.width = 8, fig.height = 8}
pca.expr <- pca(t(log2(normalized_counts+1) ), method = "svd", nPcs = 5, center = T, scale="none", cv="q2")
p1 <- plot.pca(pca.expr, metadata$condition, metadata$condition, type="scores", text=F, title="Scores") + scale_color_brewer( palette = "Set1" )
p1
```

We are able to separate groups in the first PC component.

# Statistical analysis

Filtering away genes with less than `r count.cutoff` counts.

```{r dataprep}
colData <- metadata
countData <- counts
countData <- countData[rowSums(countData) > count.cutoff,]
colData$condition <- factor(colData$condition)
dds <- DESeqDataSetFromMatrix(countData = countData, colData = colData, design = ~  condition)
```

```{r fn-deseq}

handle_deseq2 = function(dds, summarydata, column) {
  all_combs = combn(levels(summarydata[,column]), 2, simplify=FALSE)
  all_results = list()
  contrast_strings = list()
  for(comb in all_combs) {
    contrast_string = paste(comb, collapse="_vs_")
    contrast = c(column, comb)
    res = results(dds, contrast=contrast)
    res = res[order(res$padj),]
    all_results = c(all_results, res)
    contrast_strings = c(contrast_strings, contrast_string)
  }
  names(all_results) = contrast_strings
  return(all_results)
}

plot_MA = function(res){
    for(i in seq(length(res))) {
        DESeq2::plotMA(res[[i]])
        title(paste("MA plot for contrast", names(res)[i]))
    }
}

plot_gene = function(name, d){
  design = colData(d)
  ggplot(plotCounts(d, name, returnData=TRUE, intgroup = "condition",transform = TRUE), aes(x=condition,y=count, color=condition)) + 
    geom_point(size=3) +
    ggtitle(name) + 
    scale_color_brewer(palette = "Set1")
}

reduce = function(genes){
  seen = c()
  idx = sapply(genes, function(x){
    here = as.character(as.vector(unlist(strsplit(x, split = "/"))))
    if (length(here) > 100)
      return(FALSE)
    c = intersect(seen, here)
    seen <<- unique(c(seen, here))
    score = 0.6 * length(here)
    if (length(c) < score)
      return(TRUE)
    FALSE
  })
  idx
}

.isvalid= function(dd){
  if (is.null(dim(dd))){
    return(FALSE)
  }
  if (nrow(dd)==0){
    return(FALSE)
  }
  return(TRUE)
}

```

```{r ddstats}
dds <- DESeq(dds)
res <- handle_deseq2(dds, colData, "condition")
plotDispEsts(dds)
```

```{r dds-ma}
plot_MA(res)
```

```{r dds-show, results='asis'}
de_list = list()
for (cond in names(res)){
  .res = res[[cond]]
  .res$symbol <- convertIDs(row.names(.res), "ENSEMBLTRANS", "SYMBOL", org.Hs.eg.db, "useFirst")
  .res$ensembl <- convertIDs(row.names(.res), "ENSEMBLTRANS", "ENSEMBL", org.Hs.eg.db, "useFirst")

  .res$entrezgene <- convertIDs(row.names(.res), "ENSEMBLTRANS", "ENTREZID", org.Hs.eg.db, "useFirst")
  .res$description <- convertIDs(row.names(.res), "ENSEMBLTRANS", "GENENAME", org.Hs.eg.db, "useFirst")
  .res$xloc <- anno.conv[match(row.names(.res), anno.conv[,2]),1]
  cat("\n\n##", cond, "\n\n")
  summary(.res)
  cat("\n\n")
  cat("\n\n### Top 6 genes\n\n")
  .plots = lapply(rownames(.res[1:6,]), plot_gene, dds)
  do.call(grid.arrange, .plots)  
  cat("\n\n### Top 10 genes:")
  .res = as.data.frame(.res)
  #print(kable(head(.res)))
  cat("\n\n")
  res_tab = cbind(id=rownames(.res), .res)
  print(res_tab$ensembl)
  res_tab$type = get_biomart(res_tab$ensembl, "gene_biotype")
  fn = file.path(res_files,paste0(cond,"_de.tsv"))
  write.table(res_tab,fn,sep="\t",quote=T, row.names=F)
  cat("\n\n") 
  cat("Files with full table is inside attached file with name:", basename(fn))
  cat("\n\n### GO ontology of DE genes:")
  .idx = .res$padj<p.cutoff & abs(.res$log2FoldChange) > 1
  .idx[is.na(.idx)] = FALSE
  .de = .res[.idx,"entrezgene"]
  de_list[[cond]] = unique(.res[.idx,"symbol"])
  ego <- enrichGO(gene = .de[!is.na(.de)], 
                  organism = "human", ont = "BP", pAdjustMethod = "BH",
                  pvalueCutoff = 0.01, qvalueCutoff = 0.05, readable = TRUE)
  if (.isvalid(summary(ego))){
    idx =reduce(summary(ego)$geneID)
    print(kable(summary(ego)[idx, 1:7]))
  }  
    cat("\n\n")
    
}
```

```{r compress, echo=F}
name_res = compress_results(res_files)
```

#R Session Info

(useful if replicating these results)

```{r sessioninfo}
sessionInfo()
```

