---
  html_document:
    toc: true
    highlight: zenburn
    theme: united
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(tidy=TRUE, highlight=TRUE, dev="png",
               cache=FALSE, highlight=TRUE, autodep=TRUE, warning=FALSE, error=FALSE,
               message=FALSE, prompt=TRUE, comment='', fig.cap='')
```

# Overview

```{r qc-setup}
library(ggplot2)
library(reshape)
library(gplots)
library(edgeR)
library(CHBUtils)
library(pheatmap)
project_summary = "/home/lpantano/orch/scratch/kim_sc_lung/lung_sc-merged/final/2015-10-02_lung_sc-merged/project-summary.csv"
counts_file = "/home/lpantano/orch/scratch/kim_sc_lung/lung_sc-merged/final/2015-10-02_lung_sc-merged/combined.counts"
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442",
"#0072B2", "#D55E00", "#CC79A7")
summarydata = data.frame(read.table(project_summary, header=TRUE, sep=","), row.names="Name", check.rows=FALSE)
summarydata$Name = rownames(summarydata)
summarydata = summarydata[order(summarydata$Name),]
counts = read.table(counts_file, header=TRUE, row.names="id", check.names=FALSE)
counts = counts[, order(colnames(counts))]
colnames(counts) = gsub(".counts", "", colnames(counts))
# this is a list of all non user-supplied metadata columns that could appear
known_columns = c("Name", "X.GC", "Sequences.flagged.as.poor.quality",
    "rRNA.rate", "Fragment.Length.Mean", "Intronic.Rate", "Intergenic.Rate",
    "Quality.format", "Duplication.Rate.of.Mapped", "Mapped",
    "rRNA", "Sequence.length", "Transcripts.Detected", "Mean.Per.Base.Cov.",
    "Genes.Detected", "Unique.Starts.Per.Read", "unique_starts_per_read",
    "complexity","descrition")
summarydata$mapped_log10 = log10(summarydata$Mapped)
```

```{r heatmap-function}
get_heatmap_fn = function(summarydata) {
    # return the pheatmap function with or without metadata
    metadata = summarydata[, !colnames(summarydata) %in% known_columns, drop=FALSE]
    if(ncol(metadata) == 0) {
       return(pheatmap)
    }
    else {
    # rownames(metadata) = summarydata$Name
    heatmap_fn = function(data, ...) {
        pheatmap(data, annotation=metadata, clustering_distance_rows = "correlation", clustering_distance_cols = "correlation", clustering_method = "ward.D2",...)
    }
    return(heatmap_fn)
}}
heatmap_fn = get_heatmap_fn(summarydata)
```

# Quality control metrics

## Mapped reads
```{r mapped-plot}
ggplot(summarydata, aes(x=Name, y=Mapped)) +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) +
    geom_bar(stat="identity") +
    ylab("mapped reads") + xlab("")
```

## Genomic mapping rate
```{r mapping-rate-plot}
ggplot(summarydata, aes(x=Name, y=Mapping.Rate)) +
    geom_bar(stat="identity") +
    ylab("mapping rate") + xlab("") +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90))
```

## Number of genes detected
```{r genes-detected-plot}
dd = data.frame(Name=names(counts), Genes.Detected = colSums(counts > 0))
ggplot(dd, aes(x=Name, y=Genes.Detected)) +
    geom_bar(stat="identity") +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) +
    ylab("genes detected") + xlab("")
```

## Gene detection saturation
```{r saturation-plot}
dd = data.frame(Mapped=summarydata$Mapped, Genes.Detected = colSums(counts > 0))
ggplot(dd, aes(x=Mapped, y=Genes.Detected)) +
    geom_point() +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) +
    ylab("genes detected") + xlab("reads mapped")
```

## Exonic mapping rate
```{r exonic-mapping-plot}
ggplot(summarydata, aes(x=Name, y=Exonic.Rate)) +
    geom_bar(stat="identity") +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) +
    ylab("exonic mapping rate") + xlab("")
```

## rRNA mapping rate
```{r rRNA-rate-plot}
ggplot(summarydata, aes(x=Name, y=rRNA.rate)) +
    geom_bar(stat="identity") +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) +
    ylab("rRNA rate") + xlab("")
```

## Estimated fragment length of paired-end reads
```{r fragment-length-plot}
ggplot(summarydata, aes(x=Name, y=Fragment.Length.Mean)) +
    geom_bar(stat="identity") +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) +
    ylab("fragment length") + xlab("")
```

## Boxplot of log10 counts per gene
```{r boxplot-raw}
melted = melt(counts)
colnames(melted) = c("sample", "count")
melted$sample = factor(melted$sample)
melted = melted[order(melted$sample),]
melted$count = log(melted$count)
ggplot(melted, aes(x=sample, y=count)) + geom_boxplot() +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) + xlab("")
```

## Boxplot of log10 TMM-normalized counts per gene
Trimmed mean of M-values (TMM) normalization is described
[here](http://genomebiology.com/2010/11/3/R25)

Robinson, M. D., & Oshlack, A. (2010). A scaling normalization method for differential expression analysis of RNA-seq data. Genome Biology, 11(3). doi:10.1186/gb-2010-11-3-r25

```{r boxplot-normalized}
y = DGEList(counts=counts)
y = calcNormFactors(y)
normalized_counts = cpm(y, normalized.lib.sizes=TRUE)
melted = melt(normalized_counts)
colnames(melted) = c("gene", "sample", "count")
melted$sample = factor(melted$sample)
melted = melted[order(melted$sample),]
melted$count = log(melted$count)
ggplot(melted, aes(x=sample, y=count)) + geom_boxplot() +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) + xlab("")
```

## Density of log10 TMM-normalized counts
```{r density-normalized}
ggplot(melted, aes(x=count, group=sample)) +
    geom_density() +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) + xlab("")
```

## Correlation (Pearson) heatmap of TMM-normalized counts
```{r pearson-heatmap-normalized}
heatmap_fn(cor(normalized_counts, method="pearson"))
```

## Correlation (Spearman) heatmap of TMM-normalized counts
```{r spearman-heatmap-normalized}
heatmap_fn(cor(normalized_counts, method="spearman"))
```

## MDS plot of TMM-normalized counts
```{r mds-normalized}
mds(normalized_counts, k=length(colnames(normalized_counts)) - 1,summarydata$Exonic.Rate)
```


## Heatmap by concordance correlation coefficient
http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1004075

```{r propcor-heatmap}
propcor = function(x, y) {
    x = log(x + 0.1)
    y = log(y + 0.1)
    num = 2 * cov(x, y)
    denom = var(x) + var(y)
return(num/denom)}

do_propcor = function(x) {
     mat = list()
     for(i in seq_len(ncol(x))) {
         for(j in seq_len(ncol(x))) {
        x2 = x[, i]
        y2 = x[, j]
        mat = c(mat, propcor(x2, y2)) } }
    mat = unlist(mat)
    mat = matrix(mat, ncol(x), ncol(x))
    colnames(mat) = colnames(x)
    rownames(mat) = colnames(x)
    return(mat)}

heatmap_fn(do_propcor(normalized_counts))
```

## Heatmap of top 30 most expressed genes
```{r top-count-genes, results='asis'}
library(dplyr)
.ens2name = function(x){
    require(org.Mm.eg.db)
    symbol = AnnotationDbi::select(org.Mm.eg.db, as.character(x), "SYMBOL", keytype="ENSEMBL")
    symbol = symbol %>% distinct(SYMBOL)
    idx = match(x, symbol$ENSEMBL)
    # symbol$ENTREZID[!is.na(symbol$ENTREZID)]
    symbol$SYMBOL[idx]
}
select = order(rowMeans(normalized_counts),decreasing=TRUE)[1:50]
select_names = rownames(normalized_counts[select,])
select_names = .ens2name(select_names)
ma = log2(normalized_counts[select,]+0.5)
rownames(ma) = select_names

heatmap_fn(ma)
```

## Heatmap of top spc and sca most expressed genes

Sca1 was annotated as Ly6a as the official name (need confirmation)

```{r top-count-genes, results='asis'}
select_names = rownames(normalized_counts)
select_names = .ens2name(select_names)
ma = log2(normalized_counts + 0.5)
rownames(ma) = select_names
select = grepl("Sftpc",select_names) | grepl("Atx",select_names) 
select = rowSums(ma[select,]>-1)>3
heatmap_fn(ma[names(select)[select],])
```

## Filtering bad samples

* less than 1000 genes
* less than 0.5 exonic rate

## Heatmap of 30 top express and markers
```{r filter}
idx = colSums(counts>0) > 1000
summaryfilter = summarydata %>% filter(Exonic.Rate > 0.5) %>% filter(Name %in% names(idx)[idx])
counts_filter = counts[,summaryfilter$Name]

y = DGEList(counts=counts_filter)
y = calcNormFactors(y)
normalized_counts = cpm(y, normalized.lib.sizes=TRUE, log = TRUE)

select = order(rowMeans(normalized_counts),decreasing=TRUE)[1:50]
select_names = rownames(normalized_counts[select,])
select_names = .ens2name(select_names)
ma = normalized_counts[select,]
rownames(ma) = select_names

heatmap_fn(ma, main="top 50 genes")

select_names = rownames(normalized_counts)
select_names = .ens2name(select_names)
ma = normalized_counts
rownames(ma) = select_names
select = grepl("Sftpc",select_names) | grepl("Ly6a",select_names) 
select = rowSums(ma[select,]>-1)>3
heatmap_fn(ma[names(select)[select],], main="markers")

```


```{r seurat, echo=F, eval=F}
library(seurat)
nbt.data = data.frame(normalized_counts)
nbt = new("seurat", raw.data = nbt.data)
nbt = setup(nbt, project = "NBT", min.cells = 3, min.genes = 1000, is.expr = 1)

nbt = mean.var.plot(nbt)

length(nbt@var.genes)

nbt = pca(nbt)

nbt = jackStraw(nbt, num.replicate = 5, num.pc = 10)

jackStrawPlot(nbt, PCs=1:2)
```

