---
title: "small RNA-seq analysis."
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: readable
    highlight: zenburn
---

```{r output-setup, echo=FALSE}
path_results = "~/orch/scratch/Project_ManusBiosynthesis/srnaseq-rnaseq-run/final_nosp/2016-02-02_srnaseq-merged/files/clusters"
dir.create(path_results, showWarnings = FALSE)
source("~/repos/myRfunctions/transactions.R")
```

```{r knitrsetup, echo=FALSE}
library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(dev="png",echo=FALSE,
                      highlight=TRUE, autodep=TRUE,
                      comment='', fig.cap='',
                      tidy.opts=list(keep.blank.line=FALSE, width.cutoff=200),
                      cache=TRUE,warning=FALSE,message=FALSE,
                      error=FALSE, fig.width = 10, fig.height = 7)
options(bitmapType = 'cairo')
```

```{r render,eval=FALSE, echo=FALSE}
#render_2_drop("qc-summary-clusters.Rmd", "manusbio_small_rnaseq_clusters_Jan2016")
```

```{r setup}
library(knitr)
library(ggplot2)
library(reshape)
library(DESeq2)
library(genefilter)
library(CHBUtils)
library(gtools)
library(gridExtra)
library(devtools)
library(dplyr)
library(isomiRs)

root_path = "/home/lpantano/orch/scratch/Project_ManusBiosynthesis/srna-k12/work/report"
root_file = file.path(root_path)
novel_mirna_count = "/home/lpantano/orch/scratch/Project_ManusBiosynthesis/srna-k12/final/2016-02-03_srnaseq-merged/counts_novel.tsv"

condition = "condition"
```

```{r load}
# setwd(root_path)
files = read.table(file.path(root_path, "summary_re.csv"), sep=",",header=T,colClasses = "character")

samples = files[,"sample_id"]

names_stats = files[,"size_stats"]
names(names_stats) = samples

groups = files[,"group"]
names(groups) = samples

summarydata = data.frame(row.names=samples,samples=samples,group=groups)
design <- data.frame(row.names=files$sample_id, condition=files$group)

```

This analysis only takes into account sequences < 40nt and appeared in more than 2 samples and with more than 2 counts.

# Exploratory analysis

In this section we will see descriptive figures about quality of the data, 
reads with adapter, reads mapped to miRNAs, reads mapped to other small RNAs. 

## size distribution

After adapter removal, we can plot the size distribution of the small RNAs.

```{r adapter,fig.width=10}
tab = data.frame()
for (sample in samples){
    d = read.table(file.path(root_path,names_stats[sample]), sep=" ")
    tab = rbind(tab, d %>% mutate(sample=sample, group=groups[sample]))
}


reads_adapter = tab %>% group_by(sample, group) %>% summarise(total=sum(V2))
ggplot(reads_adapter, aes(x=sample,y=total,fill=group)) +
    geom_bar(stat="identity", position = "dodge") +
    ggtitle("total number of reads with adapter") +
    ylab("# reads") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


ggplot(tab, aes(x=V1,y=V2,fill=sample)) +
    geom_bar(stat="identity", position = "dodge") +
    facet_wrap(~group, ncol=2)+
    ggtitle("size distribution") +
    ylab("# reads") + xlab("size") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Novel miRNAs

```{r load-novel}
mirna = read.table(novel_mirna_count, header=T,row.names=1)
```

### Total miRNA expression

```{r total-mirna}
ggplot( data.frame(sample=colnames(mirna), total=colSums(mirna))) +
    geom_bar(aes(x=sample,y=total), stat='identity')+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Distribution of miRNAs

```{r dist-mirna}
ggplot(melt(mirna)) +
    geom_boxplot(aes(x=variable,y=value))+
    scale_y_log10()+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## Others small RNA

The data was analyzed with [seqcluster](http://seqcluster.readthedocs.org/)

This tools used all reads, uniquely mapped and multi-mapped reads. The first
step is to cluster sequences in all locations they overlap. The second step is to 
create meta-clusters: is the unit that merge all clusters that share the 
same sequences. This way the output are meta-clusters, common sequences that could
come from different region of the genome.


### classification

Number of reads in the data after each step:

* raw: initial reads
* cluster: after cluster detection
* multimap: after meta-cluster detection

```{r reads-track}
reads_stats <- read.table(file.path(root_path, "..","seqcluster", "cluster", "read_stats.tsv"),sep="\t",check.names = F)
ggplot(reads_stats, aes(x=V2, y=V1, fill=V3)) + 
    geom_bar(stat = 'identity', position = 'dodge') +
    labs(list(x="samples", y="reads")) +
    scale_fill_brewer("steps", palette = 'Set1')+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


```{r cluster}
clus <- read.table(file.path(root_path, "..", "seqcluster", "cluster", "counts.tsv"),header=T,sep="\t",row.names=1, check.names = FALSE)
ann <- clus[,2]
toomany <- clus[,1]
clus_ma <- clus[,3:ncol(clus)]
clus_ma = clus_ma[,row.names(design)]
```


Check complex meta-clusters: This kind of events happen when there are small RNA over the whole genome, and all
repetitive small rnas map to thousands of places and sharing many sequences in many positions.
If any meta-cluster is > 40% of the total data, maybe it is worth to add some filters
like: minimum number of counts `-e` or `--min--shared` in `seqcluster prepare`


```{r complex-meta-clusters}
library(edgeR)
clus_ma_norm = cpm(DGEList(clus_ma),normalized.lib.sizes = TRUE)
head(clus_ma_norm[toomany>0,])
```

### complexity

Number of miRNAs with > 10 counts.

```{r complexity-clus}

kable(as.data.frame(colSums(clus_ma>10)))

```

### Contribution by class

```{r cluster_type}
rRNA <- colSums(clus_ma[grepl("rRNA",ann) & grepl("miRNA",ann)==F,])
miRNA <- colSums(clus_ma[grepl("miRNA",ann),])
tRNA <- colSums(clus_ma[grepl("tRNA",ann) & grepl("rRNA",ann)==F & grepl("ncRNA",ann)==F & grepl("miRNA",ann)==F,])
rmsk <- colSums(clus_ma[grepl("ncRNA",ann) & grepl("rRNA",ann)==F & grepl("miRNA",ann)==F,])
total <- colSums(clus_ma)

dd <- data.frame(samples=names(rRNA),
                 rRNA=rRNA,
                 miRNA=miRNA,
                 tRNA=tRNA,
                 ncRNA=rmsk,
                total=total)
ggplot(melt(dd)) +
    geom_bar(aes(x=samples,y=value,fill=variable),
             stat='identity',
             position="dodge")+
    scale_fill_brewer(palette = "Set1")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

dd_norm = dd
dd_norm[,2:5] = sweep(dd[,2:5],1,dd[,6],"/")
ggplot(melt(dd_norm[,1:5])) +
    geom_bar(aes(x=samples,y=value,fill=variable),
             stat='identity',
             position="dodge")+
    scale_fill_brewer(palette = "Set1")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
    labs(list(title="relative proportion of small RNAs",y="% reads"))
```

```{r clus-size-by-class}
size_clus <- read.table(file.path(root_path, "..", "seqcluster", "cluster", "size_counts.tsv"),sep="\t",check.names = F)
```


# Differential expression

[DESeq2](http://bioconductor.org/packages/release/bioc/html/DESeq2.html) is used for this analysis.

```{r de-setup}
library(DESeq2)
library(DEGreport)
library(vsn)
```

```{r deseq2-handler}

filter_handle <- function(res){
    res_nona <- res[!is.na(res$padj),]
    keep <- res_nona$padj < 0.1 
    res_nona[keep,]
}

handle_deseq2 = function(dds, summarydata, column) {
  all_combs = combn(levels(summarydata[,column]), 2, simplify=FALSE)
  all_results = list()
  contrast_strings = list()
  for(comb in all_combs) {
    contrast_string = paste(comb, collapse="_vs_")
    contrast = c(column, comb)
    res = results(dds, contrast=contrast)
    res = res[order(res$padj),]
    all_results = c(all_results, res)
    contrast_strings = c(contrast_strings, contrast_string)
  }
  names(all_results) = contrast_strings
  return(all_results)
}

plot_MA = function(res){
    for(i in seq(length(res))) {
        DESeq2::plotMA(res[[i]])
        title(paste("MA plot for contrast", names(res)[i]))
    }
}

plot_volcano = function(res){
    for(i in seq(length(res))) {
        stats = as.data.frame(res[[i]][,c(2,6)])
        p = volcano_density_plot(stats, title=names(res)[i], lfc.cutoff=1)
        print(p)
    }
}

do_de = function(raw, summarydata, condition, minc=3){
    dss = DESeqDataSetFromMatrix(countData = raw[rowMeans(raw)>minc,],
                       colData = summarydata,
                       design = ~ condition)
    dss = DESeq(dss)
    plotDispEsts(dss)
    dss
}

do_norm = function(dss, root_path, prefix){
    rlog_ma = assay(rlog(dss))
    count_ma = counts(dss, normalized=TRUE)
    raw = counts(dss, normalized=FALSE)
    fn_log = file.path(path_results, paste0(prefix, "log_matrix.txt"))
    write.table(rlog_ma,fn_log,sep="\t")
    fn_count = file.path(path_results, paste0(prefix, "count_matrix.txt"))
    write.table(count_ma,fn_count,sep="\t")
    fn_raw = file.path(path_results, paste0(prefix, "raw_matrix.txt"))
    write.table(count_ma,fn_raw,sep="\t")
}

```


## Analysis for clusters

```{r de-c}
dds = do_de(clus_ma, design, condition)
```

```{r tables-c}
do_norm(dds, root_path, "clusters_")
```

### MA-plots
  
```{r DESeq-output-c, results='asis'}
all_results = handle_deseq2(dds, design, condition)
plot_MA(all_results)

clust_des = all_results[[1]]
ann_des = clus[rownames(clust_des),]
fn = file.path(path_results, "clusters_de.tsv")
write.table(cbind(clust_des, ann_des),fn,sep="\t")
    
kable(head(cbind(clust_des, ann_des)))

```

## Analysis for novel miRNAs

```{r de-m}
dds = do_de(mirna, design, condition)
```

```{r tables-m}
do_norm(dds, root_path, "mirna_")
```

### MA-plots
  
```{r DESeq-output-m, results='asis'}
all_results = handle_deseq2(dds, design, condition)
plot_MA(all_results)

clust_des = head(all_results[[1]])
fn = file.path(path_results, "mirna_de.tsv")
write.table(clust_des,fn,sep="\t")

kable(clust_des)

```


# Files

Files generated contains raw count, normalized counts, log2 normalized counts and DESeq2 results.

For visualization of clusters download: https://github.com/lpantano/seqclusterViz/archive/master.zip, open reader.html
gile with a browser, and load clusters/seqcluster.db file.

#R Session Info

```{r compress, echo=FALSE,cache=FALSE}
file.copy("/home/lpantano/orch/scratch/Project_ManusBiosynthesis/srna-k12/work/seqcluster/report/seqcluster.db", file.path(path_results, "seqcluster.db"))
# name_res = compress_results(path_results)
```

(useful if replicating these results)

```{r sessioninfo}
sessionInfo()
```

