---
output:
  knitrBootstrap::bootstrap_document:
    theme: readable
    highlight: zenburn
    theme.chooser: TRUE
    highlight.chooser: TRUE
  html_document:
    toc: true
    highlight: zenburn
---

Report last updated `r date()`.


```{r knitrsetup, echo=FALSE}
library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(dev="png",echo=FALSE,
                      highlight=TRUE, autodep=TRUE,
                      comment='', fig.cap='',
                      tidy.opts=list(keep.blank.line=FALSE, width.cutoff=200),
                      cache=TRUE,warning=FALSE,message=FALSE,
                      error=FALSE, fig.width = 12, fig.height = 9)
options(bitmapType = 'cairo')
```

```{r create-report, echo=FALSE, eval=FALSE}
#render_2_drop("srna-summary.Rmd", "vishal_kidney/FA_small_rna")
```

```{r setup}

library(ggplot2)
library(reshape)
library(DESeq2)
library(genefilter)
library(CHBUtils)
library(gtools)
library(gridExtra)
library(devtools)
library(dplyr)
library(tidyr)
library(isomiRs)
library(pheatmap)
library(cluster)
library(myRfunctions)

root_path = "~/orch/scratch/vishal_mirna_kidney/FA_model/mirna_bio/final"
root_file = file.path(root_path, "files")
dir.create(root_file, showWarnings = FALSE)

```

Only considered genes and miRNAs with FDR < 5%. Targets from targetscan
were filtered using total context score value of -0.1 or lower. 
This value correlates with the effect on the gene expression. 
In this case will be at leas 1% of effect. As well aggregate Pct should be 
higher of 0.1 that is the probability to be a real miRNA binding site
for the gene.

```{r load-files}
ma_rda = "~/orch/scratch/vishal_mirna_kidney/FA_model/mirna_bio/final/files/ma.rda"
mirna_results = "~/orch/scratch/vishal_mirna_kidney/FA_model/mirna_bio/final/files/mirna_log_matrix.txt"
mirna_design = "~/orch/scratch/vishal_mirna_kidney/FA_model/mirna_bio/final/2016-02-08_mirna_bio/report/summary_re.csv"
mirna_de = "~/orch/scratch/vishal_mirna_kidney/FA_model/mirna_bio/final/files/mirna_de.tsv"

mrna_results = "~/orch/scratch/vishal_mirna_kidney/FA_model/mrna_bio/final/2016-02-05_mrna_bio/files/rlog_counts.tsv"
mrna_design = "~/orch/scratch/vishal_mirna_kidney/FA_model/mrna_bio/final/2016-02-05_mrna_bio/project-summary.csv"
mrna_de = "~/orch/scratch/vishal_mirna_kidney/FA_model/mrna_bio/final/2016-02-05_mrna_bio/files/fa_model_simple.tsv"

mirna_ma = as.matrix(read.table(mirna_results, header=T, row.names=1))
mrna_ma = as.matrix(read.table(mrna_results, header=T, row.names=1))
mrna_ma=mrna_ma[rowSums(mrna_ma)>0,]

mirna_des = read.csv(mirna_design, row.names="sample_id")
mrna_des = read.csv(mrna_design, row.names="descritiption")

mirna_sign = read.table(mirna_de, header=T, sep="\t", row.names = 1)
mrna_sign = read.table(mrna_de, header = T, sep = "\t", row.names = 1)

mirna_keep = row.names(mirna_sign)[mirna_sign$padj<0.05]
write.table(data.frame(m=mirna_keep), "~/orch/scratch/vishal_mirna_kidney/FA_model/mirna_bio/final/files/mirna_diff.txt",row.names=F, col.names=F, quote=F)
mrna_keep = row.names(mrna_sign)[mrna_sign$padj<0.05 & abs(mrna_sign$log2FoldChange)]
mrna_keep = intersect(mrna_keep, row.names(mrna_ma)) #' fix this

if ( !file.exists(ma_rda) ){
    pairs=read.table("~/orch/scratch/vishal_mirna_kidney/FA_model/mirna_bio/final/files/matrix.tsv") %>% filter(V3< -.1)
    map = data.frame(tx=unique(pairs$V1))
    map_gene = annotate_df(map, "tx", 'mmusculus_gene_ensembl', "ensembl_transcript_id", "ensembl_gene_id")
    idx = match(pairs$V1, map_gene$tx)
    pairs$gene = map_gene[idx,2]
    pairs$value = 1
    ma = pairs %>% dplyr::select(gene,V2,value) %>%
        dplyr::distinct() %>% spread(V2,value, fill=0) %>% filter(!is.na(gene))
    row.names(ma) = ma$gene
    save(ma, file=ma_rda)
}else{
    load(ma_rda)    
}

```


# Interaction between genss and mirna over GO terms

I will do GO enrichment analysis using only the genes that are de-regulated
and are targeted by any miRNA (as well de-regulated) using targetscan database.

The universe for the analysis is all genes expressed in these samples. So,
results from GO enrichment will mean that these genes are over represented over
the expression profile of these samples.

The I will detect common patterns for genes in each term, and plot the gene
and miRNA expression in each case.

```{r annotation-preparation}
library(clusterProfiler)
library(KEGG.db)
library(org.Mm.eg.db)
library(reshape)
library(cluster)

mirna_clean = mirna_ma[,order(mirna_des$group)]
gr = mirna_des$group[order(mirna_des$group)]
mirna_clean = mirna_clean[,gr!="Day28"]
mirna_g = factor( gsub("day0", "control", 
                       sort(
                           as.character(tolower(gr[gr!="Day28"]))
                           ) ),
                 levels=c("control", "day1", "day2", "day3", "day7", "day14"))

mrna_clean = mrna_ma[,order(mrna_des$group)]
gr = mrna_des$group[order(mrna_des$group)]

data = sync_terms_with_mirs(mirna_clean, mrna_clean, as.matrix(ma), 
                     mrna_keep, mirna_keep, mirna_g, org.Mm.eg.db)


data$summary %>% dplyr::select(Description, Count, nmir)


```

```{r annotation-run}
final_df = viz_mirna_gene_enrichment(data, mirna_clean, mrna_clean, mirna_g, org.Mm.eg.db)

# save_file(final_df,"mir_gene_cluster.tsv",root_file)
```

```{r null, echo=F, eval=F}
# kegg annotation
# change to the specie you want
library(goseq) #For getting KEGG pathways easily
library(KEGG.db)

kegg = getgo(row.names(mrna_ma), fetch.cats = "KEGG", genome = "mm10", id = "ensGene")
kegg2ens = Biobase::reverseSplit(kegg)
kname = mget(names(kegg2ens), env = KEGGPATHID2NAME)
names(kegg2ens) = kname

# mirPath
cor = cor(t(mirna_norm[, -5]), t(mrna_norm), method="kendall")

target = lapply(colnames(ma)[2:ncol(ma)], function(x){
  ma$gene[ma[,x]==1]
})
names(target) = colnames(ma)[2:ncol(ma)]

Zmi = mirna_sign$stat ; names(Zmi) = row.names(mirna_sign)
Zg = mrna_sign$stat; names(Zg) = row.names(mrna_sign)

output = mirPath(DEmirs = mirna_keep, Zmi = Zmi, Zg = Zg,
                 cor = cor,
                 targets = target, pathways = kegg2ens,
                 fdr="bonferroni")
rownames(output) = output$path
output$true = 0
output[names(others),"true"] = others
output$score = output$genes/output$weight
resEnrichment = output

# network
library(igraph)
g = graph.data.frame(pairs[,c("V2", "gene")] %>% filter(gene %in% mrna_keep), directed = F)

# plot.igraph(g, vertex.color=V(g)$color, vertex.label=NA)
ebc <- edge.betweenness.community(g, directed=T)
dist = table(ebc$membership)
big = names(dist[dist>3])
del = ebc$names[!(ebc$membership %in% big)]
g = delete.vertices(g, del)
V(g)$color[V(g)$name %in%  mirna_keep]<-"grey80"
V(g)$color[V(g)$name %in%  mrna_keep]<-"orange"
g
# plot.igraph(g, vertex.color=V(g)$color, vertex.label=NA)

# custom
sum(colSums(cor[mirna_keep,mrna_keep]< -.7) > 1)

#' go through all kegg names
enrich = lapply(names(kegg2ens)[3], function(x){
   print(x)
  .genes = kegg2ens[[x]]
  if (length(.genes)<3)
    return(NULL)
  .is_regulator = rowSums(cor[mirna_keep,.genes]< -.7) > 1
  .is_target = colSums(cor[mirna_keep,.genes]< -.7) > 1
  .ts = ma[intersect(.genes, rownames(ma)), ]
  if (nrow(.ts)==0)
    return(NULL)
  .ts_target = names(rowSums(.ts[,intersect(mirna_keep, colnames(.ts))] > 0))
  .total_target = intersect(.genes[.is_target], .ts_target)
  # .total_mirs = rownames(cor)[.is_regulator]
  # .de_reg_mirna = intersect(.total_mir, mirna_keep)
  .de_target_gene = intersect(.total_target, mrna_keep)
  .de_here = intersect(.genes, mrna_keep)
  phyper(length(.de_target_gene),
         length(.total_target),
         length(.genes) - length(.total_target),
         length(.de_here)
  )
})


#
library(STATegRa)

mirna_b1 <- createOmicsExpressionSet(Data=as.matrix(mirna_ma[,row.names(mirna_de)]),
                               pData=mirna_de[,"group",drop=FALSE],
                               pDataDescr=c("classname"))

mrna_b2 <- createOmicsExpressionSet(Data=as.matrix(mrna_ma[,row.names(mrna_de)]),
                                     pData=mrna_de[,"group",drop=FALSE],
                                     pDataDescr=c("classname"))


cc <- selectCommonComps(X=mirna_ma, Y=mrna_ma, Rmax=3)
cc$common
```

