---
title: "RNA-seq analysis."
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: readable
    highlight: zenburn
---

```{r projsetup, echo=FALSE, warning=FALSE, message=FALSE}
project=" FA model - RNAseq "
labPI="Vishal"
lablocation="Longwood"
analystname="Lorena Pantano"
analystemail="lpantano@hsph.harvard.edu"
```

RNA-seq analysis for`r labPI` group at `r lablocation`.  

Contact `r analystname` (`r analystemail`) for additional details.

The most recent update of this html document occurred: `r date()`

The sections below provide code to reproduce the included results and plots. 


```{r knitrsetup, echo=FALSE}
library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(dev="png",echo=FALSE,
                      highlight=TRUE, autodep=TRUE,
                      comment='', fig.cap='',
                      tidy.opts=list(keep.blank.line=FALSE, width.cutoff=200),
                      cache=TRUE,warning=FALSE,message=FALSE,
                      error=FALSE, fig.width = 12, fig.height = 9)
options(bitmapType = 'cairo')
```

```{r project-options, echo=FALSE}
library(myRfunctions)
source("~/repos/DEGreport/R/clustering.R")
path_results = "~/orch/scratch/vishal_mirna_kidney/FA_model/mrna_bio/final/2016-02-05_mrna_bio/files"
dir.create(path_results, showWarnings = FALSE)
```

```{r create-report, echo=FALSE, eval=FALSE}
#render_2_drop("mrna-summary.Rmd", "vishal_kidney/FA_mrna")
```

# Overview

```{r qc-setup}
library(ggplot2)
library(reshape)
library(gplots)
library(edgeR)
library(CHBUtils)
library(pheatmap)
library(DESeq2)
library(tximport)
library(logging)
library(gridExtra)
library(clusterProfiler)
library(org.Mm.eg.db)
library(cluster)
library(dplyr)

# library(DT)

basicConfig()
project_summary = "~/orch/scratch/vishal_mirna_kidney/FA_model/mrna_bio/final/2016-02-05_mrna_bio/project-summary.csv"
counts_file = "~/orch/scratch/vishal_mirna_kidney/FA_model/mrna_bio/final/2016-02-05_mrna_bio/combined.counts"
tx2genes_file = "~/orch/scratch/vishal_mirna_kidney/FA_model/mrna_bio/final/2016-02-05_mrna_bio/tx2gene.csv"

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442",
"#0072B2", "#D55E00", "#CC79A7")

summarydata = data.frame(read.table(project_summary, header=TRUE, sep=","), row.names="Name", check.rows=FALSE)
summarydata$Name = rownames(summarydata)
summarydata = summarydata[order(summarydata$Name),]

if (file.exists(tx2genes_file)) {
  loginfo("Using gene counts calculated from the Sailfish transcript counts.")
  sf_files = file.path("~/orch/scratch/vishal_mirna_kidney/FA_model/mrna_bio/final",
                       rownames(summarydata), "sailfish",
                       rownames(summarydata), "quant.sf")
  names(sf_files) = summarydata$descritiption
  tx2gene = read.table(tx2genes_file, sep=",")
  txi.salmon = tximport(sf_files, type="salmon", tx2gene = tx2gene,
                        reader=readr::read_tsv,
                        countsFromAbundance="lengthScaledTPM")
  counts = txi.salmon$counts
} else {
  loginfo("Using gene counts calculated from featureCounts.")
  counts = read.table(counts_file, header=TRUE, row.names="id", check.names=FALSE)
}

# this is a list of all non user-supplied metadata columns that could appear
known_columns = c("Name", "X.GC", "Exonic.Rate", "Sequences.flagged.as.poor.quality",
    "rRNA_rate", "Fragment.Length.Mean", "Intronic.Rate", "Intergenic.Rate",
    "Mapping.Rate", "Quality.format", "Duplication.Rate.of.Mapped", "Mapped",
    "rRNA", "Sequence.length", "Transcripts.Detected", "Mean.Per.Base.Cov.",
    "Genes.Detected", "Unique.Starts.Per.Read", "unique_starts_per_read",
    "complexity", "X5.3.bias", "Duplicates.pct", "Duplicates", "Mapped.reads",
    "Median.insert.size", "Mapped.reads.pct", "Total.reads",
    "avg_coverage_per_region", "Mapped.Reads", "descritiption")
metadata = summarydata[, !colnames(summarydata) %in% known_columns, drop=FALSE]
rownames(metadata) = summarydata$descritiption
metadata$group = factor(metadata$group, levels = c("control", "day1",
                                                   "day2", "day3",
                                                   "day7", "day14"))
```

```{r heatmap-function}
get_heatmap_fn = function(summarydata) {
    # return the pheatmap function with or without metadata
    if(ncol(metadata) == 0) {
       return(pheatmap)
    }
    else {
    # rownames(metadata) = summarydata$Name
    heatmap_fn = function(data, ...) {
        pheatmap(data, annotation=metadata, clustering_method = "ward.D2", clustering_distance_cols = "correlation", ...)
    }
    return(heatmap_fn)
}}
heatmap_fn = get_heatmap_fn(summarydata)
```

# Quality control metrics {.tabset}

## Mapped reads
```{r mapped-plot, eval="Mapped" %in% colnames(summarydata)}
ggplot(summarydata, aes(x=Name, y=Mapped)) +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) +
    geom_bar(stat="identity") +
    ylab("mapped reads") + xlab("") 
```

## Genomic mapping rate
```{r mapping-rate-plot, eval="Mapping.Rate" %in% colnames(summarydata)}
ggplot(summarydata, aes(x=Name, y=Mapping.Rate)) +
    geom_bar(stat="identity") +
    ylab("mapping rate") + xlab("") +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90))
```

```{r mapping-rate-plot-noqualimap, eval="Mapped.reads.pct" %in% colnames(summarydata)}
ggplot(summarydata, aes(x=Name, y=Mapped.reads.pct)) +
    geom_bar(stat="identity") +
    ylab("mapping rate") + xlab("") +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90))
```

## Number of genes detected
```{r genes-detected-plot, eval="Genes.Detected" %in% colnames(summarydata)}
dd = data.frame(Name=colnames(counts), Genes.Detected = colSums(counts > 0))
ggplot(dd, aes(x=Name, y=Genes.Detected)) +
    geom_bar(stat="identity") +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) +
    ylab("genes detected") + xlab("")
```

## Gene detection saturation
```{r saturation-plot, eval="Genes.Detected" %in% colnames(summarydata)}
dd = data.frame(Mapped=summarydata$Mapped, Genes.Detected = colSums(counts > 0))
ggplot(dd, aes(x=Mapped, y=Genes.Detected)) +
    geom_point() +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) +
    ylab("genes detected") + xlab("reads mapped")
```

## Exonic mapping rate
```{r exonic-mapping-plot, eval="Exonic.Rate" %in% colnames(summarydata)}
ggplot(summarydata, aes(x=Name, y=Exonic.Rate)) +
    geom_bar(stat="identity") +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) +
    ylab("exonic mapping rate") + xlab("")
```

## rRNA|mitRNA mapping rate
```{r check-rRNA-eval}
eval_rRNA = "rRNA_rate" %in% colnames(summarydata) & !sum(is.na(summarydata$rRNA_rate)) == nrow(summarydata)
```
```{r rRNA-rate-plot, eval=eval_rRNA}
ggplot(summarydata, aes(x=Name, y=rRNA_rate)) +
    geom_bar(stat="identity") +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) +
    ylab("rRNA rate") + xlab("")
```

## Estimated fragment length of paired-end reads
```{r fragment-length-plot, eval="Fragment.Length.Mean" %in% colnames(summarydata)}
ggplot(summarydata, aes(x=Name, y=Fragment.Length.Mean)) +
    geom_bar(stat="identity") +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) +
    ylab("fragment length") + xlab("")
```

## Boxplot of log10 counts per gene
```{r boxplot-raw}
melted = melt(as.data.frame(counts))
colnames(melted) = c("sample", "count")
melted$sample = factor(melted$sample)
melted = melted[order(melted$sample),]
melted$count = log(melted$count)
ggplot(melted, aes(x=sample, y=count)) + geom_boxplot() +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) + xlab("")
```

## Boxplot of log10 TMM-normalized counts per gene
Trimmed mean of M-values (TMM) normalization is described
[here](http://genomebiology.com/2010/11/3/R25)

Robinson, M. D., & Oshlack, A. (2010). A scaling normalization method for differential expression analysis of RNA-seq data. Genome Biology, 11(3). doi:10.1186/gb-2010-11-3-r25

```{r boxplot-normalized}
y = DGEList(counts=counts)
y = calcNormFactors(y)
normalized_counts = cpm(y, normalized.lib.sizes=TRUE)
melted = melt(normalized_counts)
colnames(melted) = c("gene", "sample", "count")
melted$sample = factor(melted$sample)
melted = melted[order(melted$sample),]
melted$count = log(melted$count)
ggplot(melted, aes(x=sample, y=count)) + geom_boxplot() +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) + xlab("")
```

## Density of log10 TMM-normalized counts
```{r density-normalized}
ggplot(melted, aes(x=count, group=sample)) +
    geom_density() +
    theme_bw(base_size=10) +
    theme(panel.grid.major = element_line(size = .5, color = "grey"),
          axis.text.x = element_text(angle=90)) + xlab("")
```

## Correlation (Spearman) heatmap of TMM-normalized counts
```{r spearman-heatmap-normalized, eval=ncol(counts) > 1}
heatmap_fn(cor(normalized_counts, method="spearman"))
```

## MDS plot
```{r pca, eval=ncol(counts) > 1}
dds = DESeqDataSetFromTximport(txi.salmon, colData=metadata, design=~1)
vst = varianceStabilizingTransformation(dds)
# plotPCA(vst, intgroup=c("group")) + geom_text(aes(label=summarydata$Name),hjust = 0, nudge_x = 0.3)
mds(assay(vst), condition = metadata$group)
```

## MDS after variation removal

We used RUVSeq package for this with two factors to explain the unwanted variation.
```{r ruv-seq}
library(RUVSeq)
replicates=matrix(1:nrow(metadata), nrow=6, byrow = TRUE)
ruv = RUVs(assay(vst), k=2, scIdx=replicates, isLog=TRUE)
mds(log2(ruv$normalizedCounts), condition = metadata$group)
```


## Heatmap by concordance correlation coefficient
http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1004075

```{r propcor-heatmap, eval=ncol(counts) > 1}
propcor = function(x, y) {
    x = log(x + 0.1)
    y = log(y + 0.1)
    num = 2 * cov(x, y)
    denom = var(x) + var(y)
return(num/denom)}

do_propcor = function(x) {
     mat = list()
     for(i in seq_len(ncol(x))) {
         for(j in seq_len(ncol(x))) {
        x2 = x[, i]
        y2 = x[, j]
        mat = c(mat, propcor(x2, y2)) } }
    mat = unlist(mat)
    mat = matrix(mat, ncol(x), ncol(x))
    colnames(mat) = colnames(x)
    rownames(mat) = colnames(x)
    return(mat)}

heatmap_fn(do_propcor(normalized_counts))
```


```{r de-setup}
library(DESeq2)
library(DEGreport)
library(vsn)
design = ~ group
condition = "group"
```

# Differential expression

```{r deseq2-expression-analysis, results='asis'}
dds = DESeqDataSetFromTximport(txi.salmon,
    colData=metadata, design = ~ group)
dds = DESeq(dds, test="LRT", reduced = ~ 1)
```

## Effect of variance stabilization

```{r deseq-diagnostics, results='asis'}
notAllZero <- (rowSums(counts(dds))>0)
rld <- rlog(dds)
vsd <- varianceStabilizingTransformation(dds)
rlogMat <- assay(rld)
vstMat <- assay(vsd)

save_file(rlogMat, "rlog_counts.tsv", path_results)

p1=meanSdPlot(log2(counts(dds,normalized=TRUE)[notAllZero,] + 1), plot=F)$gg + ggtitle("log2 transformation")
p2=meanSdPlot(rlogMat[notAllZero,], plot=F)$gg + ggtitle("variance stabilization transformation")
p3=meanSdPlot(vstMat[notAllZero,], plot=F)$gg + ggtitle("regularized log transformation")
grid.arrange(p1,p2,p3)
```

## Dispersion estimates

```{r dispersion-estimate}
plotDispEsts(dds)
```

```{r deseq2-handler}
handle_deseq2 = function(dds, summarydata, column, all_combs=NULL) {
  if (is.null(all_combs)){
    all_combs = combn(levels(summarydata[,column]), 2, simplify=FALSE)
  }
  all_results = list()
  contrast_strings = list()
  for(comb in all_combs) {
    contrast_string = paste(comb, collapse="_vs_")
    contrast = c(column, comb)
    res = results(dds, contrast=contrast)
    res = res[order(res$padj),]
    all_results = c(all_results, res)
    contrast_strings = c(contrast_strings, contrast_string)
  }
  names(all_results) = contrast_strings
  return(all_results)
}

plot_genes = function(dds, gene){
  DESeq2::plotCounts(dds, gene, intgroup = "group")
}


cluster_genes = function(mat, sign){
  
}

```


```{r get-groups}
get_groups <- function(d, comp, condition)
{
  g <- unlist(strsplit(comp,"_"))
  g1 <- d$Name[d[, (names(d)==condition)]==g[1]]
  g2 <- d$Name[d[, (names(d)==condition)]==g[3]]
  list(g1,g2)
}
```

```{r DESeq-output, results='asis'}
res =  results(dds)
deg_res = show_deseq2_results(res, dds, rlogMat, "fa_model", 
                    FDR=0.01, FC=3, group=NULL, xs="group",
                    org=org.Mm.eg.db, 
                    path_results = path_results,
                    contrast=list(c("group", "day1", "control"),
                                  c("group", "day2", "control"),
                                  c("group", "day3", "control"),
                                  c("group", "day7","control"), 
                                  c("group", "day14", "control")))
```

## Clustering in common patterns

We used __diana__ function inside __cluster__ R package to separate genes
using the expression correlation with time. Clusters with more than 20
genes are shown. A GO and KEGG enrichment is done for each group.

A summary of diana function is copied here:

```
The diana-algorithm constructs a hierarchy of clusterings, starting with one large cluster containing all n observations. Clusters are divided until each cluster contains only a single observation.
At each stage, the cluster with the largest diameter is selected. (The diameter of a cluster is the largest dissimilarity between any two of its observations.)
To divide the selected cluster, the algorithm first looks for its most disparate observation (i.e., which has the largest average dissimilarity to the other observations of the selected cluster). This observation initiates the "splinter group". In subsequent steps, the algorithm reassigns observations that are closer to the "splinter group" than to the "old party". The result is a division of the selected cluster into two new clusters.
```

Working with `r length(deg_res$sign)` genes.

```{r clustering-res, results='asis', fig.width=12, fig.height=12, cache=FALSE}
clusters = degPatterns(rlogMat[deg_res$sign,], metadata, minc = 20, summarize = "group", time="group", col = NULL)
.df=clusters$df
.df$symbol = convertIDs(rownames(.df), "ENSEMBL", "SYMBOL", org.Mm.eg.db, "useFirst")
save_file(.df, "clusters_genes.tsv", path_results)
```

```{r clustering-fig, results='asis'}
cat("\nWorking with ", nrow(clusters$df), "genes.\n\n")
.void = lapply(clusters$pass, function(x){
  cat("\n### group:", x)
  cat("\n\n")
  .df = clusters$df
  .g = as.character(.df[.df$cluster==x, "genes"])
  .accnum = convertIDs(.g,
                       "ENSEMBL", "ENTREZID", org.Mm.eg.db, "useFirst")
  .symbol = convertIDs(.g,
                       "ENSEMBL", "SYMBOL", org.Mm.eg.db, "useFirst")
  ego <- enrichKEGG(gene = .accnum[!is.na(.accnum)], 
                  organism = "mouse",use_internal_data = TRUE)
  
  if ( "result" %in% slotNames(ego) )
    if ( nrow(ego@result) >0 )  print(kable(ego@result[,1:7]))
  
  cat("\n\n")
  ego <- enrichGO(gene = .accnum[!is.na(.accnum)], 
                  OrgDb = "org.Mm.eg.db", ont = "BP", pAdjustMethod = "BH",
                  pvalueCutoff = 0.01, qvalueCutoff = 0.1, readable = TRUE)
   if ( "result" %in% slotNames(ego) )
    print(print_enrichGO(ego@result, 30))
  cat("\n\n")
  cbind(ensembl=.g, symbol=.symbol, group=x)
})


  
```


```{r maSigPro, eval=FALSE}
## Time series nanalysis

library(maSigPro)

time <- c(rep(c(0,1,2,3,7,14), each = 3))
replicates <- c(rep(c(1:6), each = 3))
group <- rep(1, 18)
edesign <- cbind(time, replicates, group)
rownames(edesign) <- colnames(counts)

t =  maSigPro (rlogMat, edesign, degree = 6, vars = "each", main = "Test", pdf=FALSE)

p = see.genes(t$sig.genes$time3, k=3, newX11 = F)
```


#R Session Info

```{r compress, echo=FALSE, cache=FALSE}
name_res = compress_results(path_results, prefix = "FA_results", zip=TRUE)
```

(useful if replicating these results)

```{r sessioninfo}
sessionInfo()
```
