---
title: "RNA-seq analysis."
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
    theme: readable
    highlight: zenburn
---

```{r projsetup, echo=FALSE, warning=FALSE, message=FALSE}
project="LDR - RNAseq "
labPI="Greenberg"
lablocation="Longwood"
analystname="Lorena Pantano"
analystemail="lpantano@hsph.harvard.edu"
```

RNA-seq analysis for `r labPI` group at `r lablocation`.  

Contact `r analystname` (`r analystemail`) for additional details.

The most recent update of this html document occurred: `r date()`

The sections below provide code to reproduce the included results and plots. 


```{r knitrsetup, echo=FALSE}
library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(dev="png",echo=TRUE,
                      highlight=TRUE, autodep=TRUE,
                      comment='', fig.cap='',
                      tidy.opts=list(keep.blank.line=FALSE, width.cutoff=200),
                      cache=TRUE,warning=FALSE,message=FALSE,
                      error=FALSE, fig.width = 12, fig.height = 9)
options(bitmapType = 'cairo')
```

```{r project-options, echo=FALSE}
library(myRfunctions)
path_results = "~/orch/scratch/greenberg_rnaseq/ldr_model/final/files"
dir.create(path_results, showWarnings = FALSE)
```

```{r create-report, echo=FALSE, eval=FALSE}
render_2_drop("mrna-goenrichment-P4P10P16.Rmd", "greenberg_ldr/ldr-go_enrichment-P4P10P16_time")
```

```{r qc-setup}
library(ggplot2)
library(reshape)
library(gplots)
library(CHBUtils)
library(DESeq2)
library(logging)
library(gridExtra)
library(clusterProfiler)
library(DOSE)
library(org.Mm.eg.db)
library(cluster)
library(dplyr)
library(DEGreport)

# library(DT)

basicConfig()

comp_keep = c("ctl_over_time.tsv")
comp = list()
for (name in comp_keep){
    comp[[name]] = readr::read_tsv(file.path(path_results, name))
}
metadata_full = read.csv(file.path(path_results, "..", "2016-02-29_ldr_model","project-summary.csv"), row.names="Name")
```

Used genes with peaks at P4/P10/P16: 

* P4: groups 9
* P10: groups 8,11
* P16: groups 3,4

To create the groups we used all de-regulated genes over time in Ctl, and then
clustered together to show the patterns. The patterns that made a peak at the
desired times were selected as input for the GO analysis.

```{r create-tpm}

tpm = read.csv(file.path(path_results, "counts_tpm_with_fc.csv"),
                 row.names=1)

keep = comp[[1]]$padj<0.01 & !is.na(comp[[1]]$padj) & abs(comp[[1]]$log2FoldChange) > 2
samples = (grepl("ctl", names(tpm))) & !grepl("log2", names(tpm)) & !grepl("padj", names(tpm))
# samples = !grepl("log2", names(tpm)) & !grepl("padj", names(tpm)) & !(grepl("symbol", names(tpm)))
ma=as.matrix(tpm[as.character(comp[[1]]$id[keep]),samples])
ma[is.na(ma)] = 0
metadata = metadata_full[colnames(ma),]
metadata$type=metadata$group
metadata$group=droplevels(as.factor(paste0(metadata$type, metadata$time)))
pattern = degPatterns(ma, metadata = metadata, col = "type", time = "time", summarize = "group", minc = 15)

genes_keep = pattern$df$genes
comp_run = list()
for (c in unique(pattern$df$cluster)){
    comp_run[[paste0("C",c)]] = as.character(pattern$df$genes[pattern$df$cluster==c])
}
# tpm = tpm[!is.na(tpm[,1]),]
save_file(tpm[genes_keep,],"P4P10P16_counts_tpm_with_fc.csv", path_results)
```


```{r load-go-setup}
asd_terms = read.csv("~/orch/scratch/greenberg_rnaseq/lgn2/final/all-terms.csv", row.names=1)
```

# GO enrichment using DE genes

```{r write-used, echo=FALSE}
write.csv(tpm[genes_keep,], file.path(path_results, "P4P10P16_time-tpm.csv"))
```

We considered DE genes as genes with FDR < 10% and abs(log2FC) > 0.5.

## Standard GO analysis
```{r go-standard, results='asis'}
for (name in names(comp_run)){
    cat("### Go enrichment for cluster:", name, "\n\n")
    keep = as.character(comp_run[[name]])
    cat("Working with ", length(keep), " genes\n\n")
    ego <- enrichGO(keep, org.Mm.eg.db, keytype = "ENSEMBL", ont = "BP", universe = as.character(comp[[1]]$id))
    ego <- simplify(ego)
    cat("\n\n")
    print(kable(ego@result[,1:7]))
    cat("\n\n")
}
```

## Custom GO analysis

```{r clusterprofile, results='asis'}
go_res = data.frame()
for (name in names(comp_run)){
    cat("### Go enrichment for cluster:", name, "\n\n")
    keep = as.character(comp_run[[name]])
    cat("Working with ", length(keep), " genes\n\n")
    go_fake = rbind(asd_terms, data.frame(go="GO_control", genes=comp[[1]]$id))
    
    go_name = data.frame(go=c(as.character(unique(asd_terms$go)),
                              "GO_everything.else"), 
                         name=c(as.character(unique(asd_terms$go)), "control"))
    
    if (length(keep) == 0){
        cat("\n\nNo DE genes in this comparison\n\n");next}
    write.csv(tpm[keep,], file.path(path_results, paste0(name, "used4go-tpm.csv")))
    cp = enricher( keep, universe = as.character(comp[[1]]$id), 
                   pvalueCutoff = 1, qvalueCutoff = 1, 
                   TERM2GENE = go_fake, maxGSSize = 8000,
                   TERM2NAME = go_name, minGSSize = 2)
    cat("\n")
    if (!("result" %in% slotNames(cp))){
        cat("\n\nNo Genes in any terms in this comparison\n\n");next}
        
    print(kable(cp@result[,1:6]))
    # cat("\n\ngenes in GO term\n\n")
    symbol = paste(convertIDs(unique(unlist(strsplit(cp@result$geneID,"/"))),
                          "ENSEMBL", "SYMBOL", 
                          org.Mm.eg.db, "useFirst"), collapse = " ")
    cat("\n\n")
    dt = cp@result
    dt$symbol = symbol
    dt$comparision = name
    go_res = rbind(go_res, dt)
}
write.csv(go_res, file.path(path_results, paste0("P4P10P16", "_time-go_res.csv")))
```

# Difference between CDR and Ctl

```{r cdr-vs-ctl}
keep = comp[[1]]$padj<0.01 & !is.na(comp[[1]]$padj) & abs(comp[[1]]$log2FoldChange) > 2
#samples = (grepl("ctl", names(tpm))) & !grepl("log2", names(tpm)) & !grepl("padj", names(tpm))
samples = !grepl("log2", names(tpm)) & !grepl("padj", names(tpm)) & !(grepl("symbol", names(tpm)))
ma=as.matrix(tpm[as.character(comp[[1]]$id[keep]),samples])
ma[is.na(ma)] = 0
metadata = metadata_full[colnames(ma),]
metadata$type=metadata$group
metadata$group=droplevels(as.factor(paste0(metadata$type, metadata$time)))
pattern = degPatterns(ma, metadata = metadata, col = "type", time = "time", summarize = "group", minc = 15)
```


```{r cdr-vs-ctl-inspection, results='asis'}
library(pheatmap)
chosen = list("1"="P16", "2"="P16", "3"="P16", "4"="P16", "6"="P16", 
              "7"="P16","9"="P16", "10"="P16", "11"="P16", "12"="P16", "13"="P16")
for (c in names(chosen)){
    cat(paste0("## Cluster ", c, " at time point ", chosen[c]),"\n\n")
    t <- chosen[c]
    g <- as.character(pattern$df$genes[pattern$df$cluster==c])
    s <- colnames(ma)[grepl(t, colnames(ma))]
    cat("\n### heatmap of genes only in that point\n\n")
    pheatmap(log2(ma[g,s]+1), clustering_distance_cols = "correlation", clustering_method = "ward.D2", show_rownames = FALSE, annotation = metadata[,c("type", "time")], main = c)
    cat("\n### PCA of genes only in that point\n\n")
    print(mds(ma[g,s], d = "cor", k =2, condition = metadata[s, "type"]) + ggtitle(c))
    cat("\n### GO standard analysis\n\n")
    cat("\nWorking with ", length(g), " genes\n\n")
    ego <- enrichGO(g, org.Mm.eg.db, keytype = "ENSEMBL", ont = "BP", universe = as.character(comp[[1]]$id))
    ego <- simplify(ego)
    cat("\n\n")
    print(kable(ego@result[,1:7]))
    cat("\n\n")

}
```


#R Session Info

```{r compress, echo=FALSE, cache=FALSE}
#name_res = compress_results(path_results,prefix = "ldr_results")
```

(useful if replicating these results)

```{r sessioninfo}
sessionInfo()
```
