---
title: "RNA-seq analysis."
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
    theme: readable
    highlight: zenburn
---

```{r projsetup, echo=FALSE, warning=FALSE, message=FALSE}
project="LDR - RNAseq "
labPI="Greenberg"
lablocation="Longwood"
analystname="Lorena Pantano"
analystemail="lpantano@hsph.harvard.edu"
```

RNA-seq analysis for `r labPI` group at `r lablocation`.  

Contact `r analystname` (`r analystemail`) for additional details.

The most recent update of this html document occurred: `r date()`

The sections below provide code to reproduce the included results and plots. 


```{r knitrsetup, echo=FALSE}
library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(dev="png",echo=TRUE,
                      highlight=TRUE, autodep=TRUE,
                      comment='', fig.cap='',
                      tidy.opts=list(keep.blank.line=FALSE, width.cutoff=200),
                      cache=TRUE,warning=FALSE,message=FALSE,
                      error=FALSE, fig.width = 12, fig.height = 9)
options(bitmapType = 'cairo')
```

```{r project-options, echo=FALSE}
library(myRfunctions)
path_results = "~/orch/scratch/greenberg_rnaseq/ldr_model/final/files"
dir.create(path_results, showWarnings = FALSE)
```

```{r create-report, echo=FALSE, eval=FALSE}
render_2_drop("mrna-goenrichment.Rmd", "greenberg_ldr/ldr-go_enrichment")
```


```{r qc-setup}
library(ggplot2)
library(reshape)
library(gplots)
library(CHBUtils)
library(DESeq2)
library(logging)
library(gridExtra)
library(clusterProfiler)
library(DOSE)
library(org.Mm.eg.db)
library(cluster)
library(dplyr)

# library(DT)

basicConfig()

comp_keep = c("Ctrl_E18_vs_P04.tsv", "Ctrl_P04_vs_P10.tsv", "Ctrl_P10_vs_P16.tsv", "Ctrl_P16_vs_P20.tsv", "Ctrl_P20_vs_P27.tsv", "Ctrl_P27_vs_P32.tsv")
comp = list()
for (name in comp_keep){
    comp[[name]] = readr::read_tsv(file.path(path_results, name))
}

goterms =  readr::read_tsv(file.path(path_results, "..", "immune-terms.txt"), 
                           col_names = FALSE) %>%
    tidyr::separate(X1, into=c("id", "term"), sep="  ")    
goterms2 =  readr::read_tsv(file.path(path_results, "..", "microglia-terms.txt"), 
                            col_names = FALSE) %>%
    tidyr::separate(X1, into=c("id", "term"), sep="  ")    
goterms3 =  rbind(goterms, goterms2)    

# https://www.biostars.org/p/52101/#68158
library(biomaRt)
ensembl = useMart("ensembl",dataset="mmusculus_gene_ensembl") #uses human ensembl annotations
#gets gene symbol, transcript_id and go_id for all genes annotated with GO:0007507
gene.goterms <- getBM(attributes=c('ensembl_gene_id', 'go_id'),
                   filters = 'go_id', values = goterms$id, mart = ensembl)
gene.goterms2 <- getBM(attributes=c('ensembl_gene_id', 'go_id'),
                   filters = 'go_id', values = goterms2$id, mart = ensembl)
gene.goterms3 <- data.frame(ensembl_gene_id=unique(c(gene.goterms$ensembl_gene_id, gene.goterms2$ensembl_gene_id)))

```

## GO enrichment using DE genes

We considered DE genes as genes with FDR < 10%.

Using following GO terms as one:

```{r ,results='asis'}
kable(rbind(goterms, goterms2))
```


```{r clusterprofile, results='asis'}
for (i in 1:length(comp)){
    cat("### Go enrichment for:", names(comp)[i], "\n\n")
    keep = comp[[i]]$padj<0.1 & !is.na(comp[[i]]$padj)

    go_fake = rbind(
              data.frame(go="GO_immune",
                               genes=as.character(unique(gene.goterms$ensembl_gene_id))),
              data.frame(go="GO_microglia",
                         genes=as.character(unique(gene.goterms2$ensembl_gene_id))),
              data.frame(go="GO_both",
                         genes=as.character(unique(gene.goterms3$ensembl_gene_id)))
              )
    
    go_fake = rbind(go_fake, data.frame(go="GO_control", genes=comp[[i]]$id))
    
    # go_fake = data.frame(go="GO_immune", genes=comp[[1]]$id[keep])
    # go_fake = rbind(go_fake, go_fake)
    go_name = data.frame(go=c("GO_immune", "GO_microglia", "GO_both", "GO_everything.else"), 
                         name=c("custom", "custom", "custom", "control"))
    
    if (sum(keep) == 0){
        cat("\n\nNo DE genes in this comparison\n\n");next}
    
    cp = enricher( comp[[i]]$id[keep], universe = comp[[i]]$id, 
                   pvalueCutoff = 1, qvalueCutoff = 1,
                   TERM2GENE = go_fake, maxGSSize = 2000,
                   TERM2NAME = go_name, minGSSize = 2)
    cat("\n")
    if (!("result" %in% slotNames(cp))){
        cat("\n\nNo Genes in any terms in this comparison\n\n");next}
        
    print(kable(cp@result[,1:6]))
    cat("\n\ngenes in GO term\n\n")
    print(paste(convertIDs(unique(unlist(strsplit(cp@result$geneID,"/"))),
                           "ENSEMBL", "SYMBOL", 
                           org.Mm.eg.db, "useFirst"), collapse = " "))
    cat("\n\n")
}
```


#R Session Info

```{r compress, echo=FALSE, cache=FALSE}
#name_res = compress_results(path_results,prefix = "ldr_results")
```

(useful if replicating these results)

```{r sessioninfo}
sessionInfo()
```
