---
title: "RNA-seq analysis."
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
    theme: readable
    highlight: zenburn
---

```{r projsetup, echo=FALSE, warning=FALSE, message=FALSE}
project="LDR - RNAseq "
labPI="Greenberg"
lablocation="Longwood"
analystname="Lorena Pantano"
analystemail="lpantano@hsph.harvard.edu"
```

RNA-seq analysis for `r labPI` group at `r lablocation`.  

Contact `r analystname` (`r analystemail`) for additional details.

The most recent update of this html document occurred: `r date()`

The sections below provide code to reproduce the included results and plots. 


```{r knitrsetup, echo=FALSE}
library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(dev="png",echo=TRUE,
                      highlight=TRUE, autodep=TRUE,
                      comment='', fig.cap='',
                      tidy.opts=list(keep.blank.line=FALSE, width.cutoff=200),
                      cache=TRUE,warning=FALSE,message=FALSE,
                      error=FALSE, fig.width = 12, fig.height = 9)
options(bitmapType = 'cairo')
```

```{r project-options, echo=FALSE}
library(myRfunctions)
path_results = "~/orch/scratch/greenberg_rnaseq/ldr_model/final/files"
dir.create(path_results, showWarnings = FALSE)
```

```{r create-report, echo=FALSE, eval=FALSE}
render_2_drop("mrna-goenrichment-CTL_over_time.Rmd", "greenberg_ldr/ldr-go_enrichment-CTL_over_time")
```

```{r qc-setup}
library(ggplot2)
library(reshape)
library(gplots)
library(CHBUtils)
library(DESeq2)
library(logging)
library(gridExtra)
library(clusterProfiler)
library(DOSE)
library(org.Mm.eg.db)
library(cluster)
library(dplyr)

# library(DT)

basicConfig()

comp_keep = c("ctl_over_time.tsv", "Ctrl_E18_vs_P04.tsv", "Ctrl_P04_vs_P10.tsv", "Ctrl_P10_vs_P16.tsv")
comp = list()
for (name in comp_keep){
    comp[[name]] = readr::read_tsv(file.path(path_results, name))
}
```

```{r create-tpm}
comp_keep = list.files(path_results, "vs")
comp_tpm = list()
for (name in comp_keep){
    if (!grepl("go", name))
        comp_tpm[[name]] = readr::read_delim(file.path(path_results, name), delim = "\t")
}

tpm = read.table(file.path(path_results, "..", "2016-02-29_ldr_model","combined.gene.sf.tpm"),
                 row.names=1, header=T)
tpm$symbol = convertIDs(row.names(tpm), "ENSEMBL", "SYMBOL", org.Mm.eg.db, "useFirst")

for (name in names(comp_tpm)){
    .df = as.data.frame(comp_tpm[[name]])
    rownames(.df) = .df$id
    names(.df) = paste0(names(.df), "_", gsub(".tsv", "", name), sep="")
    
    tpm = cbind(tpm, .df[row.names(tpm), c(3,7)])
}

# tpm = tpm[!is.na(tpm[,1]),]
save_file(tpm,"counts_tpm_with_fc.csv", path_results)

```

```{r load-go-setup}
asd_terms = read.csv("~/orch/scratch/greenberg_rnaseq/lgn2/final/all-terms.csv", row.names=1)
```

# GO enrichment using DE genes

We considered DE genes as genes with FDR < 10% and abs(log2FC) > 0.5.

## Standard GO analysis
```{r go-standard, results='asis'}
for (i in 1:length(comp)){
    cat("### Go enrichment for:", names(comp)[i], "\n\n")
    keep = comp[[i]]$padj<0.1 & !is.na(comp[[i]]$padj) & abs(comp[[i]]$log2FoldChange) > 0.5
    cat("Working with ", sum(keep), " genes\n\n")
    ego <- enrichGO(as.character(comp[[i]]$id[keep]), org.Mm.eg.db, keytype = "ENSEMBL", ont = "BP", universe = as.character(comp[[i]]$id))
    ego <- simplify(ego)
    cat("\n\n")
    print(kable(ego@result[,1:7]))
    cat("\n\n")
}
```

## Custom GO analysis

```{r clusterprofile, results='asis'}
go_res = data.frame()
for (i in 1:length(comp)){
    cat("### Go enrichment for:", names(comp)[i], "\n\n")
    keep = comp[[i]]$padj<0.1 & !is.na(comp[[i]]$padj) & abs(comp[[i]]$log2FoldChange) > 0.5
    cat("Working with ", sum(keep), " genes\n\n")
    go_fake = rbind(asd_terms, data.frame(go="GO_control", genes=comp[[i]]$id))
    
    go_name = data.frame(go=c(as.character(unique(asd_terms$go)),
                              "GO_everything.else"), 
                         name=c(as.character(unique(asd_terms$go)), "control"))
    
    if (sum(keep) == 0){
        cat("\n\nNo DE genes in this comparison\n\n");next}
    
    write.csv(tpm[as.character(comp[[i]]$id[keep]),], file.path(path_results, paste0( names(comp)[i], "used4go-tpm.csv")))
    cp = enricher( as.character(comp[[i]]$id[keep]), universe = as.character(comp[[i]]$id), 
                   pvalueCutoff = 1, qvalueCutoff = 1, 
                   TERM2GENE = go_fake, maxGSSize = 8000,
                   TERM2NAME = go_name, minGSSize = 2)
    cat("\n")
    if (!("result" %in% slotNames(cp))){
        cat("\n\nNo Genes in any terms in this comparison\n\n");next}
        
    print(kable(cp@result[,1:6]))
    # cat("\n\ngenes in GO term\n\n")
    symbol = paste(convertIDs(unique(unlist(strsplit(cp@result$geneID,"/"))),
                          "ENSEMBL", "SYMBOL", 
                          org.Mm.eg.db, "useFirst"), collapse = " ")
    cat("\n\n")
    dt = cp@result
    dt$symbol = symbol
    dt$comparision = names(comp)[i]
    go_res = rbind(go_res, dt)
}

write.csv(go_res, "~/orch/scratch/greenberg_rnaseq/ldr_model/final/files/ctl_over_time-go_res.csv")
```

#R Session Info

```{r compress, echo=FALSE, cache=FALSE}
#name_res = compress_results(path_results,prefix = "ldr_results")
```

(useful if replicating these results)

```{r sessioninfo}
sessionInfo()
```
