---
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
    toc_float: true
    theme: readable
    highlight: zenburn
---

Report last updated `r date()`.


```{r knitrsetup, echo=FALSE}
library(myRfunctions)
library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(dev="png",
                      highlight=TRUE, autodep=TRUE,
                      comment='', fig.cap='',
                      tidy.opts=list(keep.blank.line=FALSE, width.cutoff=200),
                      cache=TRUE,warning=FALSE,message=FALSE,
                      error=FALSE, fig.width = 12, fig.height = 9)
options(bitmapType = 'cairo')
```
This report analyzes FA model from Vishal data for Jsha group.

```{r libraries}
library(ggplot2)
library(reshape)
library(limma)
library(CHBUtils)
library(biomaRt)
library(dplyr)
library(cluster)
library(gridExtra)
library(logging)
library(org.Mm.eg.db)
library(DEGreport)
library(dplyr)
order_group=c("control", "day1", "day2", "day7", "day14")
root_path = "~/orch/scratch/jshah_fibrosis_rnaseq/protein"
result_files = file.path(root_path, "FA_vishal_files")
dir.create(result_files, showWarnings = F, recursive = T)
basicConfig(level='INFO')
```

# Protein analysis {.tabset}

FC > 2 and FDR < 1%

```{r protein-de, results='asis'}
counts = read.csv(file.path(root_path, "FA_vishal_phosquant_summedSN.csv")) %>% tidyr::separate(Protein.ID, c("sp", "Uniprot.ID", "type"),sep = "[::|::]", extra = "merge") %>%
             mutate(site_position=gsub("NA", "", paste(site1, site2, site3, site4, sep=";"))) %>%
             mutate(localization=gsub("NA", "", paste(loc1, loc2, loc3, loc4, sep=";")))
             
    
row.names(counts) = paste0(counts$Uniprot.ID, ":", counts$site_position)

clean_counts = counts[,19:28]
names(clean_counts)[1:10] = c("control-1", "control-2", "day1-1", "day1-2",
                       "day2-1", "day2-2", "day7-1", "day7-2", "day14-1", "day14-2")

cat("## Expression density of 'raw' data\n\n")
ggplot(melt(clean_counts), aes(x=log2(value), colour=variable)) + geom_density()

dge = edgeR::DGEList(clean_counts)
dge = edgeR::calcNormFactors(dge, method="TMM")

coldata = data.frame(row.names=colnames(clean_counts), samples=colnames(clean_counts)) %>% tidyr::separate(samples,into = c("group"), extra = "drop")
coldata$group = sapply(coldata$group, tolower)
norm_counts = edgeR::cpm(dge, log=T)

MA <- normalizeBetweenArrays(as.matrix(norm_counts), method="none")

cat("## Expression density of normalize data\n\n")
ggplot(melt(as.data.frame(MA)), aes(x=value, colour=variable)) + geom_density()

cat("## MDS plot\n\n")
mds(MA, condition = coldata$group)
save_file(MA, "fa_model_log2_counts.tsv", result_files)
design <- model.matrix(~ 0 + coldata$group)
colnames(design) = c("normal", "day1", "day14", "day2", "day7")
fit <- lmFit(MA, design)
contrast.matrix <- makeContrasts(day1-normal, day2-day1, day7-day2, day14-day7, 
                                 day2-normal, day7-normal, day14-normal,
                                 day7-day1, day7-day2, 
                                 day14-day1, day14-day2,
                                 levels=design)
fit2 <- contrasts.fit(fit, contrast.matrix)
fit2 <- eBayes(fit2)
res = topTable(fit2, adjust="BH", n="Inf")
save_file(res, "fa_model.tsv", result_files)
absMaxFC = rowMax(as.matrix(abs(res[,1:11])))
sign = row.names(res[absMaxFC>1 & res$adj.P.Val<0.1,])

cat("## MDS plot of DE proteins\n\n")
mds(MA[sign,], condition = coldata$group)

coldata$group = factor(coldata$group, levels=order_group)
cat("## Clustering analysis\n\n")
clusters = degPatterns(MA[sign,], coldata, minc = 15, summarize = "group", time="group", col = NULL)
.df=clusters$df
save_file(.df, "clusters_genes_cluster.tsv", result_files)

```

# R Session Info

(useful if replicating these results)

```{r sessioninfo}
sessionInfo()
```


