---
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
    toc_float: true
    theme: readable
    highlight: zenburn
---

Report last updated `r date()`.


```{r knitrsetup, echo=FALSE}
library(myRfunctions)
library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(dev="png",
                      highlight=TRUE, autodep=TRUE,
                      comment='', fig.cap='',
                      tidy.opts=list(keep.blank.line=FALSE, width.cutoff=200),
                      cache=TRUE,warning=FALSE,message=FALSE,
                      error=FALSE, fig.width = 12, fig.height = 9)
options(bitmapType = 'cairo')
```

```{r create-report, echo=FALSE, eval=FALSE}
# library(rmarkdown);library(knitr);render("meta_proteomic.Rmd")
# render_2_drop("meta_proteomic.Rmd", "shah_fibrosis/proteomics_original")
```




```{r libraries}
library(ggplot2)
library(reshape)
library(limma)
library(CHBUtils)
library(biomaRt)
library(dplyr)
library(cluster)
library(gridExtra)
library(logging)
library(org.Mm.eg.db)
library(DEGreport)
library(UpSetR)
library(dplyr)

result_files = "~/orch/scratch/jshah_fibrosis_rnaseq/protein/files_original"
root_path = "~/orch/scratch/jshah_fibrosis_rnaseq/protein"
dir.create(result_files, showWarnings = F, recursive = T)
basicConfig(level='INFO')
```

```{r fn}

intersect_fa <- function(.df, fa_de, universe){
    all_genes=convertIDs(unique(c(fa_de)), "ENSEMBL", "UNIPROT", org.Mm.eg.db)
    all_genes = all_genes[!is.na(all_genes)]
    .t=lapply(unique(.df$cluster), function(nc){
        cat("\n\ncluster:", nc, "\n\n")
        .this=as.character(.df$genes[.df$cluster==nc])
        common_genes = unique(c(all_genes,.this))
        all_set = data.frame(row.names=common_genes, id=common_genes, fa=0,cluster=0)
        fa_de_c=intersect(fa_de, common_genes)
        all_set[fa_de_c,"fa"] = 1;all_set[.this,"cluster"] = 1;
        set_names=c("fa", "cluster")
        if (length(.this)<10)
            return(all_set)
        if (sum(colSums(all_set[,2:3])>0)==1){
            cat("\n**No shared genes**\n")
            return(all_set)
        }
        upset(all_set, sets=set_names[colSums(all_set[,2:3])>0])
        .this_only = setdiff(.this, fa_de)
        cat("\n**Number of unique genes", length(.this_only), "**\n")
        # res =  runGO(.this_only, org.Mm.eg.db, from="ENSEMBL", universe = universe)
        # cat("\n\nGo enrichment:\n\n")
        # if(!is.null(res$table))
        #     print(res$table)
        # cat("\n\n")
        # if(!is.null(res$table))
        #     print(res$plot)
        return(all_set[all_set$cluster==1,])
    })
    
    do.call(rbind, .t)
}


do_de <- function(counts, coldata, design, contrast, prefix, path){
    # mapping = data.frame(id=rownames(counts))
    # mapping = annotate_df(mapping, "id", 'mmusculus_gene_ensembl', "uniprot_genename", "ensembl_gene_id")  %>% distinct(ensembl_gene_id)
    
    # mapping2 = data.frame(id=unique(counts$Gene.symbol)[!is.na(unique(counts$Gene.symbol))])
    # mapping2 = annotate_df(mapping2, "id", 'mmusculus_gene_ensembl', "wikigene_name", "ensembl_gene_id")  %>% distinct(ensembl_gene_id)
    
    # counts[as.character(mapping$id), "ensembl"] = mapping$ensembl_gene_id 
    # idx = match(as.character(mapping2$id),as.character(counts$Gene.symbol))
    # counts[idx, "ensembl"] = mapping2$ensembl_gene_id
    # counts$symbol = convertIDs(as.character(counts$ensembl), "ENSEMBL", "SYMBOL", org.Mm.eg.db)
    counts$ensembl = convertIDs(as.character(counts$Uniprot.Id), "UNIPROT", "ENSEMBL", org.Mm.eg.db)
    save_file(counts, paste0(prefix, "_counts_annotated.csv"), path)
    
    # counts = counts %>% distinct(ensembl) %>% filter(!is.na(ensembl))
    clean_counts = counts[,row.names(coldata)]
    # rownames(clean_counts) = counts$ensembl
    names(clean_counts) = sapply(names(clean_counts), tolower)
    
    cat("## Expression density of 'raw' data\n\n")
    print(ggplot(melt(clean_counts), aes(x=log2(value), colour=variable)) + geom_density())
    
    ref = which.max(colSums(clean_counts))
    ratio = colSums(clean_counts) / ref
    norm_counts = sweep(clean_counts,2,ratio,"/")
    norm_counts = sweep(norm_counts,1,rowSums(norm_counts),"/") * 100
    norm_counts[is.na(norm_counts)] = 0.5
    norm_counts[norm_counts==0] = 0.5
    
    cat("\n\n## Mean vs SD")
    plot(rowMeans(norm_counts),apply(norm_counts, 1, sd), 
         pch=20, ylab="SD counts", xlab="average counts")
    
    MA <- normalizeBetweenArrays(as.matrix(norm_counts), method="none")
    
    cat("\n\n## Expression density of normalize data\n\n")
    print(ggplot(melt(as.data.frame(MA)), aes(x=value, colour=variable)) + geom_density())
    
    cat("\n\n## MDS plot\n\n")
    print(mds(MA, condition = coldata$condition))
    save_file(MA, paste0(prefix, "_model_log2_counts.tsv"), path)
    # fit <- lmFit(MA, design)
    # fit2 <- contrasts.fit(fit, contrast.matrix)
    # fit2 <- eBayes(fit2)
    # res = topTable(fit2, adjust="BH", n="Inf")
    # res
    MA
}
```

```{r load-extra}
root_path_fa = "~/orch/scratch/vishal_mirna_kidney/publish/FA-model/protein/files_publish"
fa_matrix = read.csv(file.path(root_path_fa, "fa_model.tsv"), row.names = 1)
fa_de = rownames(fa_matrix[fa_matrix$adj.P.Val<0.01 & rowMax(abs(as.matrix(fa_matrix[,1:11]))) > 2, ])
```

In this study there are no replicates. I use log2(ratio) to filter genes
for clustering step. The ratio is between JCK vs KO, and 
there is one for each time point. I am using a cutoff of 1 and refering to them
like **DE proteins**.


I use TMM normalization from edgeR package assuming the majority of genes
are not de-regulated, so the ratio between the genes for each sample should be 
the factor size to use to bring all library to the same scale.

# Proteomic

## Early

```{r protein-early, results='asis'}
counts = read.csv(file.path(root_path, "early_prot_P5-P26ProtQuant 2016.csv"),skip = 1) %>% tidyr::separate(Uniprot.protein.Id, c("sp", "Uniprot.Id", "type"),sep = "[::|::]", extra = "merge")
row.names(counts) = counts$Uniprot.Id

names(counts)[6:15] = c("JCK_P5","JCK_P10","JCK_P15","WT_P26","JCK_P26" ,"WT_P5","WT_P10","WT_P15","WT_P20","JCK_P20")

coldata = data.frame(row.names=colnames(counts)[6:15], group=colnames(counts)[6:15]) %>% tidyr::separate(group,into = c("condition","time"), extra = "drop",remove = F)
coldata$group = sapply(coldata$group, tolower)

design <- model.matrix(~ 0 + coldata$group)
colnames(design) = gsub("coldata[$]group", "",colnames(design))
contrast.matrix <- makeContrasts(jck_p5-wt_p5,jck_p10-wt_p10,jck_p15-wt_p15,jck_p20-wt_p20, levels=design)

res = do_de(counts, coldata, design, contrast, "prot_early", result_files)
comp = list(c("jck_p5", "wt_p5"), c("jck_p10", "wt_p10"),
            c("jck_p15", "wt_p15"), c("jck_p20", "wt_p20"),
            c("jck_p26", "wt_p26"))
fc = lapply(comp, function(c){
    log2(res[,c[1]] / (res[,c[2]]))
})
fc_tab = do.call(cbind, fc)
colnames(fc_tab) = c("P5", "P10", "P15", "P20", "P26_1")

#save_file(res, "prot_early_model.tsv", result_files)
absMaxFC = rowMax(abs(fc_tab))
sign = row.names(res)[absMaxFC>1]

cat("\n\n## MDS plot of DE proteins\n\n")
mds(res[sign,], condition = coldata$condition)

order_group = c("P5", "P10", "P15", "P20", "P26_1")
coldata_fc = data.frame(row.names=colnames(fc_tab), group = factor(colnames(fc_tab), levels=order_group))
early = as.data.frame(fc_tab)
early_col = coldata_fc
early_res = res
```

## Clustering early

```{r protein-early-clus}
absMaxFC = rowMax(abs(as.matrix(early)))
sign = row.names(early)[absMaxFC>1]
prot_early_clusters = degPatterns(as.matrix(early[sign,]), early_col, minc = 15, summarize = "group", time="group", col = NULL, reduce=TRUE, scale=FALSE, fixy=c(-2,2), cutoff = 0.1)
.df=prot_early_clusters$df[prot_early_clusters$df$cluster %in% prot_early_clusters$pass,]
save_file(.df, "prot_early_clusters_genes_cluster.tsv", result_files)
```

## Late

```{r protein-late, results='asis'}
counts = read.csv(file.path(root_path, "late_prot_150715_ProQuant_tc-d161.csv"),skip = 1) %>% tidyr::separate(Protein.Id, c("sp", "Uniprot.Id", "type"),sep = "[::|::]", extra = "merge")
row.names(counts) = counts$Uniprot.Id
names(counts)[4] = "Gene.symbol"
names(counts)[6:15] = c("JCK_P26","JCK_P35","JCK_P44","JCK_P50", "JCK_P64",
                        "WT_P26","WT_P35","WT_P44","WT_P50","WT_P64")

coldata = data.frame(row.names=colnames(counts)[6:15], group=colnames(counts)[6:15]) %>% tidyr::separate(group,into = c("condition","time"), extra = "drop",remove = F)
coldata$group = sapply(coldata$group, tolower)

design <- model.matrix(~ 0 + coldata$group)
colnames(design) = gsub("coldata[$]group", "",colnames(design))

res = do_de(counts, coldata, design, contrast, "prot_late", result_files)
comp = list(c("jck_p26", "wt_p26"), c("jck_p35", "wt_p35"),
            c("jck_p44", "wt_p44"), c("jck_p50", "wt_p50"),
            c("jck_p64", "wt_p64"))
fc = lapply(comp, function(c){
    log2(res[,c[1]] / res[,c[2]])
})
fc_tab = do.call(cbind, fc)
colnames(fc_tab) = c("P26_2", "P35", "P44", "P50", "P64")

#save_file(res, "prot_late_model.tsv", result_files)
absMaxFC = rowMax(abs(fc_tab))
sign = row.names(res)[absMaxFC>1]

cat("\n\n## MDS plot of DE proteins\n\n")
mds(res[sign,], condition = coldata$condition)

order_group = c("P26_2", "P35", "P44", "P50", "P64")
coldata_fc = data.frame(row.names=colnames(fc_tab), group = factor(colnames(fc_tab), levels=order_group))
late = as.data.frame(fc_tab)
late_col = coldata_fc
late_res = res
```

## Clustering late

```{r protein-late-clus}
absMaxFC = rowMax(abs(as.matrix(late)))
sign = row.names(late)[absMaxFC>1]
prot_late_clusters = degPatterns(as.matrix(late[sign,]), late_col, minc = 15, summarize = "group", time="group", col = NULL, reduce=TRUE, scale=FALSE, fixy=c(-2,2), cutoff = 0.1)
.df=prot_late_clusters$df[prot_late_clusters$df$cluster %in% prot_late_clusters$pass,]
save_file(.df, "prot_late_clusters_genes_cluster.tsv", result_files)
```

## P26 replicate QC

```{r p26-comp}
keep = intersect(rownames(late), rownames(early))
plot(early_res[keep,"jck_p26"], late_res[keep, "jck_p26"], pch=20, main="early vs late JCK26")
plot(early_res[keep,"wt_p26"], late_res[keep, "wt_p26"], pch=20,  main="early vs late WT26")
plot(early_res[,"jck_p26"], early_res[, "wt_p26"], pch=20,  main="early JCK26 vs WTP16")
plot(late_res[,"jck_p26"], late_res[, "wt_p26"], pch=20, 
     main="late JCK26 vs WTP16")
plot(early[keep,"P26_1"], late[keep, "P26_2"], pch=20, main="FC correlation")
abline(h=1, col="red")
abline(h=-1, col="red")
abline(v=1, col="red")
abline(v=-1, col="red")

```

## MDS batch effect
```{r prot-batch}
col = grepl("wt", c(colnames(early_res), colnames(late_res)))
mds(cbind(early_res[keep,], late_res[keep,]), condition = col)
```


## Clustering analysis

```{r cluster-merge, results='asis'}
keep = intersect(rownames(late), rownames(early))
prot_meta = as.data.frame(cbind(early[keep,], late[keep,]))
prot_meta_col = rbind(early_col, late_col)
absMaxFC = rowMax(abs(as.matrix(prot_meta)))
names(absMaxFC) = rownames(prot_meta)
sign = row.names(prot_meta)[absMaxFC>1]
prot_clusters = degPatterns(as.matrix(prot_meta[sign,]), prot_meta_col, minc = 15, summarize = "group", time="group", col = NULL, reduce=TRUE, scale=FALSE, fixy=c(-2,2))
.df=prot_clusters$df[prot_clusters$df$cluster %in% prot_clusters$pass,]
save_file(.df, "prot_clusters_genes_cluster.tsv", result_files)

prot_meta_fa = intersect_fa(.df, fa_de, rownames(prot_meta))
prot_meta$absMaxFC = absMaxFC
prot_meta[as.character(.df$genes), "cluster"] <- .df$cluster 
prot_meta[as.character(prot_meta_fa$id), "fa"] <- prot_meta_fa$fa
prot_meta$symbol = convertIDs(rownames(prot_meta), "UNIPROT", "SYMBOL", org.Mm.eg.db)
prot_meta$uniprot = convertIDs(rownames(prot_meta), "UNIPROT", "ENSEMBL", org.Mm.eg.db)
save_file(prot_meta, "prot_meta.tsv", result_files)
```


## Clustering analysis with batch normalization

```{r cluster-merge-2, results='asis'}
keep = intersect(rownames(late), rownames(early))
absMaxFC = rowMax(as.matrix(data.frame(early=rowMax(abs(as.matrix(early[keep,]))), late=rowMax(abs(as.matrix(late[keep,]))))))
fs = early[keep, "P26_1"] / late[keep, "P26_2"]
prot_meta = as.data.frame(cbind(early[keep,], late[keep,] * fs))
prot_meta_col = rbind(early_col, late_col)
names(absMaxFC) = rownames(prot_meta)
sign = row.names(prot_meta)[absMaxFC>1]
prot_clusters = degPatterns(as.matrix(prot_meta[sign,]), prot_meta_col, minc = 15, summarize = "group", time="group", col = NULL, reduce=TRUE, scale=FALSE, fixy=c(-2,2))
.df=prot_clusters$df[prot_clusters$df$cluster %in% prot_clusters$pass,]
save_file(.df, "prot_norm_clusters_genes_cluster.tsv", result_files)

prot_meta_fa = intersect_fa(.df, fa_de, rownames(prot_meta))
prot_meta$absMaxFC = absMaxFC
prot_meta[as.character(.df$genes), "cluster"] <- .df$cluster 
prot_meta[as.character(prot_meta_fa$id), "fa"] <- prot_meta_fa$fa
prot_meta$symbol = convertIDs(rownames(prot_meta), "UNIPROT", "SYMBOL", org.Mm.eg.db)
prot_meta$uniprot = convertIDs(rownames(prot_meta), "UNIPROT", "ENSEMBL", org.Mm.eg.db)
save_file(prot_meta, "prot_norm_meta.tsv", result_files)
```


# Phospho-proteomic

##  Early

```{r phospho-protein-early, results='asis'}
counts = read.csv(file.path(root_path, "early_phospo_P5-P26PhosphoSiteQuant2016.csv"),skip = 1) %>% tidyr::separate(Protein.Id, c("sp", "Uniprot.Id", "type"),sep = "[::|::]", extra = "merge")
row.names(counts) = paste0(counts$Uniprot.Id, ":", counts$Site.Position)
names(counts)[4] = "Gene.symbol"

names(counts)[9:18] = c("JCK_P5","JCK_P10","JCK_P15","WT_P26","JCK_P26" ,"WT_P5","WT_P10","WT_P15","WT_P20","JCK_P20")

coldata = data.frame(row.names=colnames(counts)[9:18], group=colnames(counts)[9:18]) %>% tidyr::separate(group,into = c("condition","time"), extra = "drop",remove = F)
coldata$group = sapply(coldata$group, tolower)

design <- model.matrix(~ 0 + coldata$group)
colnames(design) = gsub("coldata[$]group", "",colnames(design))
contrast.matrix <- makeContrasts(jck_p5-wt_p5,jck_p10-wt_p10,jck_p15-wt_p15,jck_p20-wt_p20, levels=design)

res = do_de(counts, coldata, design, contrast, "phospho-prot_early", result_files)
comp = list(c("jck_p5", "wt_p5"), c("jck_p10", "wt_p10"),
            c("jck_p15", "wt_p15"), c("jck_p20", "wt_p20"),
            c("jck_p26", "wt_p26"))
fc = lapply(comp, function(c){
    log2(res[,c[1]] / res[,c[2]])
})
fc_tab = do.call(cbind, fc)
colnames(fc_tab) = c("P5", "P10", "P15", "P20", "P26_1")

#save_file(res, "early_phos-prot_model.tsv", result_files)
absMaxFC = rowMax(abs(fc_tab))
sign = row.names(res)[absMaxFC>1]

cat("\n\n## MDS plot of DE proteins\n\n")
mds(res[sign,], condition = coldata$condition)

order_group = c("P5", "P10", "P15", "P20", "P26_1")
coldata_fc = data.frame(row.names=colnames(fc_tab), group = factor(colnames(fc_tab), levels=order_group))
early = as.data.frame(fc_tab)
early_col = coldata_fc
early_res = res
```

## Clustering early

```{r pprotein-early-clus}
absMaxFC = rowMax(abs(as.matrix(early)))
sign = row.names(early)[absMaxFC>1.5]
pprot_early_clusters = degPatterns(as.matrix(early[sign,]), early_col, minc = 15, summarize = "group", time="group", col = NULL, reduce=TRUE, scale=FALSE, fixy=c(-2,2), cutoff = 0.1)
.df=pprot_early_clusters$df[pprot_early_clusters$df$cluster %in% pprot_early_clusters$pass,]
save_file(.df, "phospho-prot_early_clusters_genes_cluster.tsv", result_files)
```

## Late 

```{r phospho-protein-late, results='asis'}
counts = read.csv(file.path(root_path, "late_phospo_150715_PhosSiteQuant_tc-d161p_Ti-TMT.csv"),skip = 1) %>% tidyr::separate(Protein.Id, c("sp", "Uniprot.Id", "type"),sep = "[::|::]", extra = "merge")
row.names(counts) = paste0(counts$Uniprot.Id, ":", counts$Site.Position)
names(counts)[4] = "Gene.symbol"
names(counts)[9:18] = c("JCK_P26","JCK_P35","JCK_P44","JCK_P50", "JCK_P64",
                        "WT_P26","WT_P35","WT_P44","WT_P50","WT_P64")

coldata = data.frame(row.names=colnames(counts)[9:18], group=colnames(counts)[9:18]) %>% tidyr::separate(group,into = c("condition","time"), extra = "drop",remove = F)
coldata$group = sapply(coldata$group, tolower)

design <- model.matrix(~ 0 + coldata$group)
colnames(design) = gsub("coldata[$]group", "",colnames(design))

res = do_de(counts, coldata, design, contrast, "phospho-prot_late", result_files)
comp = list(c("jck_p26", "wt_p26"), c("jck_p35", "wt_p35"),
            c("jck_p44", "wt_p44"), c("jck_p50", "wt_p50"),
            c("jck_p64", "wt_p64"))
fc = lapply(comp, function(c){
    log2(res[,c[1]] / res[,c[2]])
})
fc_tab = do.call(cbind, fc)
colnames(fc_tab) = c("P26_2", "P35", "P44", "P50", "P64")

#save_file(res, "late_phos-prot_model.tsv", result_files)
absMaxFC = rowMax(abs(fc_tab))
sign = row.names(res)[absMaxFC>1]

cat("\n\n## MDS plot of DE proteins\n\n")
mds(res[sign,], condition = coldata$condition)

order_group = c("P26_2", "P35", "P44", "P50", "P64")
coldata_fc = data.frame(row.names=colnames(fc_tab), group = factor(colnames(fc_tab), levels=order_group))
late = as.data.frame(fc_tab)
late_col = coldata_fc
late_res = res
```

## Clustering late

```{r pprotein-late-clus}
absMaxFC = rowMax(abs(as.matrix(late)))
sign = row.names(late)[absMaxFC>1]
pprot_late_clusters = degPatterns(as.matrix(late[sign,]), late_col, minc = 15, summarize = "group", time="group", col = NULL, reduce=TRUE, scale=FALSE, fixy=c(-2,2), cutoff = 0.1)
.df=pprot_late_clusters$df[pprot_late_clusters$df$cluster %in% pprot_late_clusters$pass,]
save_file(.df, "phospho-prot_late_clusters_genes_cluster.tsv", result_files)
```

## P26 replicate QC

```{r p26-pp-comp}
keep = intersect(rownames(late), rownames(early))
plot(early_res[keep,"jck_p26"], late_res[keep, "jck_p26"], pch=20)
plot(early_res[keep,"wt_p26"], late_res[keep, "wt_p26"], pch=20)
plot(early[keep,"P26_1"], late[keep, "P26_2"], pch=20, main="FC correlation")
# low=sapply(c(early_res[keep,"wt_p26"],late_res[keep, "wt_p26"]),mean)<4
# points((early[keep,"P26"])[low], (late[keep, "P26"])[low], pch=20, col="red")
abline(h=1, col="red")
abline(h=-1, col="red")
abline(v=1, col="red")
abline(v=-1, col="red")

```

## MDS batch effect
```{r pprot-batch}
col = grepl("wt", c(colnames(early_res), colnames(late_res)))
mds(cbind(early_res[keep,], late_res[keep,]), condition = col)
```

## Clustering analysis

```{r phospho-cluster-merge, results='asis'}
keep = intersect(rownames(late), rownames(early))
pprot_meta = as.data.frame(cbind(early[keep,], late[keep,]))
pprot_meta_col = rbind(early_col, late_col)
absMaxFC = rowMax(abs(as.matrix(pprot_meta)))
names(absMaxFC) = rownames(pprot_meta)
sign = row.names(pprot_meta)[absMaxFC>1.5]
pprot_clusters = degPatterns(as.matrix(pprot_meta[sign,]), prot_meta_col, minc = 15, summarize = "group", time="group", col = NULL, reduce=TRUE, scale=FALSE, fixy=c(-2,2))
.df=pprot_clusters$df[pprot_clusters$df$cluster %in% pprot_clusters$pass,]
save_file(.df, "phospho-prot_clusters_genes_cluster.tsv", result_files)

# pprot_meta_fa = intersect_fa(.df, fa_de, rownames(pprot_meta))
pprot_meta$absMaxFC = absMaxFC
pprot_meta[as.character(.df$genes), "cluster"] <- .df$cluster 
# pprot_meta[as.character(pprot_meta_fa$id), "fa"] <- pprot_meta_fa$fa
pprot_meta$symbol = convertIDs(gsub(":.*", "", rownames(pprot_meta)), "UNIPROT", "SYMBOL", org.Mm.eg.db)
pprot_meta$uniprot = convertIDs(gsub(":.*", "", rownames(pprot_meta)), "UNIPROT", "ENSEMBL", org.Mm.eg.db)
save_file(pprot_meta, "phospho-prot_meta.tsv", result_files)
```

## Clustering analysis with batch normalization

```{r phospho-cluster-merge-2, results='asis'}
keep = intersect(rownames(late), rownames(early))
absMaxFC = rowMax(as.matrix(data.frame(early=rowMax(abs(as.matrix(early[keep,]))), late=rowMax(abs(as.matrix(late[keep,]))))))
fs = early[keep, "P26_1"] / late[keep, "P26_2"]
pprot_meta = as.data.frame(cbind(early[keep,], late[keep,] * fs))
pprot_meta_col = rbind(early_col, late_col)
names(absMaxFC) = rownames(pprot_meta)
sign = row.names(pprot_meta)[absMaxFC>1.5]
pprot_clusters = degPatterns(as.matrix(pprot_meta[sign,]), prot_meta_col, minc = 15, summarize = "group", time="group", col = NULL, reduce=TRUE, scale=FALSE, fixy=c(-2,2))
.df=pprot_clusters$df[pprot_clusters$df$cluster %in% pprot_clusters$pass,]
save_file(.df, "phospho-prot_norm_clusters_genes_cluster.tsv", result_files)

# pprot_meta_fa = intersect_fa(.df, fa_de, rownames(pprot_meta))
pprot_meta$absMaxFC = absMaxFC
pprot_meta[as.character(.df$genes), "cluster"] <- .df$cluster 
# pprot_meta[as.character(pprot_meta_fa$id), "fa"] <- pprot_meta_fa$fa
pprot_meta$symbol = convertIDs(gsub(":.*", "", rownames(pprot_meta)), "UNIPROT", "SYMBOL", org.Mm.eg.db)
pprot_meta$uniprot = convertIDs(gsub(":.*", "", rownames(pprot_meta)), "UNIPROT", "ENSEMBL", org.Mm.eg.db)
save_file(pprot_meta, "phospho-prot_norm_meta.tsv", result_files)
```


# Join both data

I take the clusters in both data set and calculate the correlation value using the 
median values of the genes in each cluster in each data. I transform the correlation 
in distance because it is easy to plot heatmaps. The distance is **1-cor**

Then, I take only the cluster pairs with distance < 0.3 (cor > 0.7). For those pairs, I show
the common number of genes and the GO enrichment using all the genes in both 
clusters. This may help to focus on some of them first. 

```{r merge, results='asis'}
prot_clus = unique(prot_meta$cluster)[!is.na(unique(prot_meta$cluster))]
pprot_clus = unique(pprot_meta$cluster)[!is.na(unique(pprot_meta$cluster))]
cor = lapply(prot_clus, function(nc1){
    sapply(pprot_clus, function(nc2){
        g1 = apply(prot_meta[prot_meta$cluster==nc1 & !is.na(prot_meta$cluster),1:9], 2, median)
        g2 = apply(pprot_meta[pprot_meta$cluster==nc2 & !is.na(pprot_meta$cluster),1:9], 2, median)
        1-cor.test(g1, g2)$estimate
    })
})
cor <- do.call(rbind, cor)
colnames(cor) <- paste0("pProt-", pprot_clus)
rownames(cor) <- paste0("Prot-", prot_clus)
pheatmap::pheatmap(cor, breaks = c(0,0.1,0.2,0.3,2), color = c("blue4","steelblue","lightblue","white"), display_numbers = T, main="correlation")

universe = unique(c(rownames(prot_meta), 
                    gsub(":.*", "", rownames(pprot_meta))))
.void = apply(melt(cor<0.3) %>% filter(value==TRUE), 1, function(x){
    cat("\n\n### Clusters", paste(x[1], x[2]), "\n\n")
    nc1 = gsub("Prot-", "", x[1])
    nc2 = gsub("pProt-", "", x[2])
    g1 = rownames(prot_meta[prot_meta$cluster==nc1 & !is.na(prot_meta$cluster),])
    g1c = apply(prot_meta[g1, 1:9], 2, median)
    g2 = rownames(pprot_meta[pprot_meta$cluster==nc2 & !is.na(pprot_meta$cluster),])
    g2c = apply(pprot_meta[g2, 1:9], 2, median)
    g2 = gsub(":.*$", "", g2)
    common = length(intersect(g1,g2))
    cat("\n\ncommon genes:", common, "\n\n")
    df = rbind(
        data.frame(log2FC=g1c, x=names(g1c), molecule="protein"), 
        data.frame(log2FC=g2c, x=names(g2c), molecule="phospho-protein"))
    df$x = factor(df$x, levels=names(g1c))
    p <- ggplot(df,aes(x=x, y=log2FC, color=molecule, group=molecule)) +
        geom_point() + geom_line()
    print(p)
    res =  runGO(unique(c(g1,g2)), org.Mm.eg.db, from="UNIPROT", universe = universe)
    # .this = unique(c(g1,g2))
    # .this_prot = convertIDs(.this, "ENSEMBL", "UNIPROT", org.Mm.eg.db)
    # res_kegg = enrichKEGG(.this_prot, organism="mmu", keyType = "uniprot")
    # print(res_kegg)
    cat("\n\nGo enrichment:\n\n")
    if(!is.null(res$table))
        print(res$table)
    cat("\n\n")
    if(!is.null(res$table))
        print(res$plot)
    NULL
})
```

```{r-test, eval=FALSE, echo=FALSE}
intersect(rownames(pprot_meta[!is.na(pprot_meta$cluster),]),
rownames(prot_meta[!is.na(prot_meta$cluster),]))

g = unique(c((pprot_meta[!is.na(pprot_meta$cluster),"uniprot"]),
(prot_meta[!is.na(prot_meta$cluster),"uniprot"])))
g = g[!is.na(g)]

kegg = enrichKEGG(g,organism = "mmu", keyType = "uniprot", qvalueCutoff  = 1,pvalueCutoff = 1)

p = "04150"
library(pathview)
g_c = c(pprot_meta[match(g, pprot_meta$uniprot), "cluster"],
        prot_meta[match(g, prot_meta$uniprot), "cluster"])
names(g_c) = c(g,g)
g_c = g_c[!is.na(g_c)]
g_c_s = g_c[unlist(strsplit(split="/",kegg@result$geneID[3]))]
pathview(g_c_s, pathway.id = p, species = "mmu", kegg.native = T,gene.idtype="UNIPROT",discrete=list(gene=TRUE, cpd=TRUE),limit=list(gene=8, cpd=2),bins=list(gene=8,cpd=2),both.dirs=list(gene=FALSE, cpd=FALSE))
```

# R Session Info

```{r compress, echo=FALSE, cache=FALSE}
#name_res = compress_results(result_files, prefix = "model_omic_ournorm", zip = TRUE)
```

(useful if replicating these results)

```{r sessioninfo}
sessionInfo()
```


