---
output:
    knitrBootstrap::bootstrap_document:
        theme: readable
        highlight: zenburn
        theme.chooser: TRUE
        highlight.chooser: TRUE
    html_document:
        toc: true
        highlight: zenburn

---
    
# Report
    
## Overview

There are 4 samples, `s1` and `s2` get long fragments and `s3` and `s4` get short fragments.
Among the pairs, there are some difference in protocol, like amplification cycles 
and different ligation method (i am guessing from the metadata)

## Method

* Map read 1
* QC for read 1
* Detect polyA in read 2: remove first 5 nt TAG at position ~17
* Overlap read 1 with genes and retrieve read-gene information
* Summarize read 1 and read 2 to get information for each gene (counts, polyA, modifications)
As well an extra tuner step is applied to get more precisely polyA fragment.
* Remove false positive polyA. I mapped read 2
to the genome and remove reads that map entirely since we are trying to detect 
post-modifications. Make sense? There were a lot of reads with short polyA
that ended up with long modifications, and mapped perfectly to the genome. 

### Setup: loading packages and data
    
```{r setup,cache=FALSE}
library(knitr)
library(rmarkdown)
library(knitrBootstrap)
library(ggplot2)
library(gplots)
library(reshape)
library(DESeq2)
library(edgeR)
library(genefilter)
library(CHBUtils)
library(gtools)
library(gridExtra)
library(devtools)
library(dplyr)
library(rjson)
library(GGally)
source("~/repos/myRfunctions/transactions.R")
dn <- "daley_tailseq"
root_path<-"~/ody/projects/daley_mRNA_mod/tailseq-112014"
path_files = "~/repos/pipelines/daley_tailseq/reports/"
```

```{r render,eval=FALSE}
knitr::opts_chunk$set(tidy=TRUE, highlight=TRUE, dev="png",
                      cache=TRUE, highlight=TRUE, autodep=TRUE, warning=FALSE, error=FALSE,
                      cache.path="~/.cache/tailseq-report/html_cached/", eval=TRUE,
                      message=FALSE, prompt=TRUE, comment='', fig.cap='', bootstrap.show.code=FALSE)
render_2_drop("~/repos/pipelines/daley_tailseq/reports/tailseq.rmd",dn)
```


```{r load}
setwd(root_path)
```

```{r read-summary}
read_summary = function(dd){
    dd = dd %>% filter(V3 > 1 & !grepl("N", V2) & !is.na(V2))
    dd$type = as.character(dd$V2)
    dd$type[ dd$type!="counts" & dd$type!="polyA" ] = "Mod"
    dd_melt = dd %>% group_by(type) %>% summarise(genes=length(unique(V1)))
    p = ggplot(dd_melt, aes(y=genes, x=type)) + 
        geom_bar( stat = 'identity') +
        ggtitle("Number of genes")
    print(p)
    dd_melt = dd %>% group_by(V1,type) %>% summarise(total=sum(V3))
    p <- ggplot(as.data.frame(dd_melt) , aes(x=type, y=total, fill=type))+
        geom_boxplot() + 
        scale_fill_brewer( palette = "Set1" ) + 
        scale_y_log10() +
        ggtitle("Average expression")
    print(p)
    dd_melt = dd %>% filter(V2=="AA" | V2=="A" | V2=="C") %>% 
              dplyr::select(V1,type=V2) %>% group_by(type) %>% 
              summarise(genes=length(unique(V1)))
    p = ggplot(dd_melt, aes(y=genes, x=type)) + 
        geom_bar( stat = 'identity', positions="dodge" ) +
        ggtitle("Number of genes")
    print(p)
}

get_set_genes = function(dd){
    poly = unique((dd %>% filter(V2=="polyA"))[,1])
    aa = unique((dd %>% filter(V2=="AA"))[,1])
    a = unique((dd %>% filter(V2=="A"))[,1])
    c = unique((dd %>% filter(V2=="C"))[,1])
    mod = unique((dd %>% filter(V2!="counts" & V2!="polyA"))[,1])
    n10 = length( c(aa,a,c) ) * 0.1
    n5 = length( c(aa,a,c) ) * 0.05
    top10 = unique(( dd %>% filter( V2=="A" |  V2=="AA" | V2=="C"  ) %>%
                         arrange( desc(V3) ))[1:n10,1])
    top5 = unique(( dd %>% filter( V2=="A" |  V2=="AA" | V2=="C"  ) %>%
                         arrange( desc(V3) ))[1:n5,1])
    return( list("polyA"=poly, "mod"=mod, "a"=a, "aa"=aa, "c"=c, "top10"=top10, "top5"=top5) )
}

filter_out = function(dd, cuta=30, cutmod=2){
    dd = dd %>% filter(V3 > cutmod & !grepl("N", V2) & !is.na(V2))
    genes = (dd %>% filter( V2=="polyA" & V3>=cuta ))[,1]
    dd[ dd$V1 %in%  genes, ]
}

poly_read_summary = function(dd){
    dd = dd %>% filter(V3 > 1 & !grepl("N", V2) & !is.na(V2))
    dd$type = as.character(dd$V2)
    dd$type[ dd$type!="counts" & dd$type!="polyA" ] = "Mod"
    dd_melt = dd %>% group_by(type) %>% summarise( total=sum(V3) )
    p = ggplot(dd_melt, aes(y=total, x=type)) + 
        geom_bar( stat = 'identity') +
        ggtitle("Number of reads")
    print(p)
    dd_melt = dd %>% filter(V2=="AA" | V2=="A" | V2=="C") %>% 
              dplyr::select(mod=V2, V3) %>% group_by(mod) %>%
              summarise( total=sum(V3) )
    p = ggplot(dd_melt, aes(y=total, x=mod)) + 
        geom_bar(stat="identity", position = "dodge") +
        ggtitle("Number of reads")
    print(p)
}


rm_na = function(x){
    x[!is.na(x)]
}

```


```{r go,eval=FALSE}
library(org.Hs.eg.db)
#set <- universeGO(org.Hs.egGO,"Homo sapiens")

run_one_go = function(g, name){
    g = g[!is.na(g)]
    mf <- runGO(g,set,"MF",as.list(GOMFCHILDREN))
    #plotGO(mf, name)
    write.table(mf[[1]], paste0(name, ".txt") )
    #print(head(mf[[1]]))
    return(mf[[1]][1:20,])
}

run_go_enrichment = function(g, sample){
    table = vector("list")
    table[[1]] = run_one_go(g[[1]], paste0(sample,"_mf_polyA"))
    table[[2]] = run_one_go(g[[2]], paste0(sample,'_mf_mod'))
    table[[3]] = run_one_go(g[[3]], paste0(sample,'_mf_a'))
    table[[4]] = run_one_go(g[[4]], paste0(sample,'_mf_aa'))
    table[[5]] = run_one_go(g[[5]], paste0(sample,'_mf_c'))
    table[[6]] = run_one_go(g[[6]], paste0(sample,'_mf_top10'))
    table[[7]] = run_one_go(g[[7]], paste0(sample,'_mf_top5'))
    table
}

```

```{r venn-function}
get_entrez = function(g){
  require(biomaRt)
  ensembl = useMart('ensembl', dataset = "hsapiens_gene_ensembl")
  df = getBM(attributes=c("ensembl_gene_id", "entrezgene"), filters="ensembl_gene_id", values=g,mart=ensembl)
  return(df[,2])
}


do_df = function(sample, groups){
    data.frame(gene=groups[[sample]],sample=sample,is=1)
}

do_venn = function(groups)
{
    n = length(groups)
    all_genes = do.call(rbind, lapply(names(groups), do_df, groups) )
    
    total_genes = all_genes %>% group_by(sample) %>% summarise(total = sum(is)) %>% ungroup()
    all_genes = reshape(all_genes, timevar = "sample",idvar = "gene", direction = "wide")
    all_genes[is.na(all_genes)] = 0
    
    all_genes_sum = all_genes %>% group_by_(.dots = (names(all_genes[,1:n+1]))) %>% summarise(total = n()) %>% ungroup()
    per = apply( all_genes_sum, 1, function(x){
        unlist((x[n+1]/as.vector(total_genes[,2])) * x[1:n] * 100 )
        })
    per[per==0] = NA
    rownames(per) = names(groups)
 
    num = apply( all_genes_sum, 1, function(x){
        unlist(x[n+1] * x[1:n]  )
        })
    num[num==0] = NA
    rownames(num) = names(groups)
 
    
    aheatmap( t(as.matrix(t(per))), annCol = all_genes_sum[,1:n], Colv = NA, Rowv = NA,cellheight = 45, cellwidth = 25, fontsize=9, txt = num )
    
}

```


## polyA in read-2

purple (reads with polyA), blue (reads without polyA) and orange (reads with polyA < 6 nt) are the total amount of reads. red (modifications) are reads with modifications after polyA. And green are the ones without the 5 nt TAG along the read.

These are the stats of read_2 before taking into account the ones where the paired
read_1 maps onto a gene. 

And we applied here another filter to reduce the number of false positive of polyA.
Normally short polyA with long modifications are removed here. We just don't count when read2 maps perfectly to the genome and is at the expected distance from the read1.

```{r ploya}
dd = data.frame()
for (file in list.files(paste0(root_path,"/work"), "stat$") ){
    data = read.table(paste0(root_path, "/work/", file) )
    dd = rbind(dd, data)
}
dd$sample = c(rep("s1",5), rep("s2",5), rep("s3",5) , rep("s4",5) )
ggplot( dd, aes( x=sample, y=V2, fill=V1 ) ) +
    geom_bar( stat = 'identity', position = 'dodge') + 
    scale_fill_brewer( palette = "Set1" ) +
    labs( list( y="reads" ) )

```

## After filtering

After that filter, we reduce the total number of reads, but still have a lot.
Similar number than public database.
Showing sample 1 only. 

```{r after-filtering}
poly_read_summary(s1)
```

### polyA > 200 and size >= 8 nt

```{r polya-filter-200}
s1_raw = read.table(paste0(root_path,"/work/s1_summary.dat"))
s1 = filter_out(s1_raw,200)
s2_raw = read.table(paste0(root_path,"/work/s2_summary.dat"))
s2 = filter_out(s2_raw,200)
s3_raw = read.table(paste0(root_path,"/work/s3_summary.dat"))
s3 = filter_out(s3_raw,200)
s4_raw = read.table(paste0(root_path,"/work/s4_summary.dat"))
s4 = filter_out(s4_raw,200)
genes_s1 = get_set_genes(s1)
genes_s2 = get_set_genes(s2)
genes_s3 = get_set_genes(s3)
genes_s4 = get_set_genes(s4)


```


```{r public-data}
sp<-read.table("~/ody/projects/daley_mRNA_mod/tailseq_GSE51299/SPtable.csv",header=T,sep=",")
library(biomaRt)
mart = useMart("ensembl", dataset="hsapiens_gene_ensembl")
g <- getBM( attributes=c("ensembl_gene_id", "hgnc_symbol","gene_biotype") , filters=
"hgnc_symbol"    , values =as.character(sp$Gene.Name) ,mart=mart)
idx<-match(sp$Gene.Name,g$hgnc_symbol)
sp$ens<-g[idx,1]
sp.f<-sp[,c(24,2,3,11,12,13,14,15,16,17,18,19),]

public = vector("list")

sp.f = sp.f[!is.na(sp.f$ens),]
public[["polyA"]] = data.frame(V1=sp.f$ens, V2="polyA", V3=sp.f$Total.poly.A...tags)

p = sp.f[sp.f[,5]>1,1]
public[["a"]] = data.frame(V1=p, V2="A", V3=sp.f[sp.f$ens %in% p ,5])

p = sp.f[sp.f[,8]>1,1]
public[["aa"]] = data.frame(V1=p, V2="AA", V3=sp.f[sp.f$ens %in% p ,8])

p = sp.f[sp.f[,6]>1,1]
public[["c"]] = data.frame(V1=p, V2="C", V3=sp.f[sp.f$ens %in% p ,6])

public_raw = do.call(rbind, public)

sp_reworked<-read.table("~/ody/projects/daley_mRNA_mod/tailseq_GSE51299/summary.complete.8As",header=F,sep="\t")
rw_raw = sp_reworked[,c(1,3,4)]
names(rw_raw) = c("V1","V2","V3")
rw_raw$V2 = as.character(rw_raw$V2)
rw_raw$V2[rw_raw$V2=="None"] = "polyA"

public = filter_out(public_raw,200)
genes_public = get_set_genes(public)
rw = filter_out(rw_raw,200)
genes_rw = get_set_genes(rw)

```


```{r biotype, eval=FALSE}
s1_g <- getBM( attributes=c("ensembl_gene_id", "hgnc_symbol","gene_biotype") , filters=
"ensembl_gene_id"    , values =as.character(s1[[1]]) ,mart=mart)
s2_g <- getBM( attributes=c("ensembl_gene_id", "hgnc_symbol","gene_biotype") , filters=
"ensembl_gene_id"    , values =as.character(s2[[1]]) ,mart=mart)
s3_g <- getBM( attributes=c("ensembl_gene_id", "hgnc_symbol","gene_biotype") , filters=
"ensembl_gene_id"    , values =as.character(s3[[1]]) ,mart=mart)
s4_g <- getBM( attributes=c("ensembl_gene_id", "hgnc_symbol","gene_biotype") , filters=
"ensembl_gene_id"    , values =as.character(s4[[1]]) ,mart=mart)
```


#### Common genes among 4 samples

I changed venn diagrams by heatmaps. Each top bar is a sample, when two bars
overlaps, the cells below are the number of genes in common. The color mean the
percentage of common genes for that sample. 

For instance, the first right column shows the common genes in the four samples: 1788 in common. Being the 60% for the first two samples, and the 90% for the two last samples.

```{r common}

type = c("polyA","Modification","add-U","add-UU","add-G","top 10%","top 5%")
type_name = c("polyA","mod","a","aa","c","top10","top5")
names(type) = type_name
 void = lapply(type_name, function(x){
     do_venn(list(s1=genes_s1[[x]],
               s2=genes_s2[[x]],
               s3=genes_s3[[x]],
               s4=genes_s4[[x]]))
     title(type[x])
 })

```


#### s1/s2 with public data

I added here public data using the supplementary table (as PUBLIC in figures), and the re-analyzed data we did with the public data (as RW in figures - last row)

```{r common-public}
 void = lapply(type_name[c(1,3,4,5)], function(x){
     do_venn(list(s1=genes_s1[[x]],
               s2=genes_s2[[x]],
               public=genes_public[[x]],
               rw=genes_rw[[x]]))
     title(type[x])
 })
```

### polyA > 30 and size >= 8 nt

```{r poya-filter-30}
s1 = filter_out(s1_raw,30)
s2 = filter_out(s2_raw,30)
s3 = filter_out(s3_raw,30)
s4 = filter_out(s4_raw,30)
genes_s1 = get_set_genes(s1)
genes_s2 = get_set_genes(s2)
genes_s3 = get_set_genes(s3)
genes_s4 = get_set_genes(s4)
rw = filter_out(rw_raw,30)
genes_rw = get_set_genes(rw)
public = filter_out(public_raw,30)
genes_public = get_set_genes(public)

```

#### Common genes among 4 samples

```{r common-30}

type = c("polyA","Modification","add-U","add-UU","add-G","top 10%","top 5%")
type_name = c("polyA","mod","a","aa","c","top10","top5")
 void = lapply(type_name, function(x){
     do_venn(list(s1=genes_s1[[x]],
               s2=genes_s2[[x]],
               s3=genes_s3[[x]],
               s4=genes_s4[[x]]))
     title(x)
 })

```

#### s1/s2 with public data

```{r common-public-30}
 void = lapply(type_name[c(1,3,4,5)], function(x){
     do_venn(list(s1=genes_s1[[x]],
               s2=genes_s2[[x]],
               public=genes_public[[x]],
               rw=genes_rw[[x]]))
     title(type[x])
 })
```

### genes in s1/s2 but not in public data

```{r filter-30-not_in_public}
genes_notin_public =   setdiff( intersect(as.character(genes_s1[["polyA"]]),as.character(genes_s2[["polyA"]])) , genes_public[["polyA"]])
genes_in_public =   intersect( intersect(as.character(genes_s1[["polyA"]]),as.character(genes_s2[["polyA"]])) , genes_public[["polyA"]])

#genes_notin_public_type = getBM( attributes=c("ensembl_gene_id", "refseq_mrna","gene_biotype") , filters=
#"ensembl_gene_id"    , values =genes_notin_public ,mart=mart)
write.table(genes_in_public,paste0(root_path,"/work/genes_in_public"),row.names=F,col.names=F,quote=F)
intersect(genes_notin_public,public_raw$V1)
```

They are not in the table, but if I look at least at one of this genes, is full of polyA sequences, 
so, it should be something relating to the method, or maybe they filter something else??

This is what raw data is for this gene

found HISEQ:272:HAVVLADXX:1:1103:13345:22631    TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTT      
found HISEQ:272:HAVVLADXX:1:1103:16584:26706    TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTGT                  
found HISEQ:272:HAVVLADXX:1:1103:2751:31339     TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTAATTTTTTTTTTTTTTTTGTCCACCA 
found HISEQ:272:HAVVLADXX:1:1103:17487:65762 AA-TTTTTTTTATTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTAATTTTTTTTTT 


I have taken all genes like this, and plot the T percentage average and 
the base-quality average per gene at the first 50 positions.

I plot the % of T at the first 50 nucleotides after PRIMER+TAG and MODIFICATIONS removal.
Each row is a gene not seen in the public data. 

```{r T-mean}
unseen = read.table( paste0(root_path,"/work/genes_notin_public.dat"),sep=" ")
seen = read.table( paste0(root_path,"/work/genes_in_public.dat"),sep=" ")
pheatmap( (unseen %>% filter(V1=="seq"))[,3:50],  cluster_rows = T, cluster_cols = F, main = "Not in commont")
pheatmap( (seen %>% filter(V1=="seq"))[,3:50],  cluster_rows = T, cluster_cols = F,main = "In common")
```

Also I plot the base-quality of these same bases, just to make sure they are good

```{r qual-mean}
pheatmap( (unseen %>% filter(V1=="qual"))[,3:50],  cluster_rows = F, cluster_cols = F,main="Not in common" )
pheatmap( (seen %>% filter(V1=="qual"))[,3:50],  cluster_rows = F, cluster_cols = F,main="In common" )
```

If we plot the number of time the same position is tagged by a read (something
proportional to PCR artifacts), we don't see a difference in the distribution
among the common or non-common genes.

```{r pcr}
unseen_pcr = read.table( paste0(root_path,"/work/s1_quality_check.dat.pcr"),sep=" ")
unseen_pcr$seen = 1
unseen_pcr$seen[unseen_pcr$V1 %in% genes_notin_public] = 0
table(unseen_pcr$seen)
ggplot(unseen_pcr, aes(x=factor(seen),y=V3) ) +
    geom_boxplot(alpha=0.3)+
    scale_y_log10() +
    labs(list(y="log10(counts)",x="common"))
```


### genes in public but not in s1/s2

All polyA counts for the genes in the public set but not in the new data are low. 

```{r filter-30-not_in_s1s2}
genes_notin_s1s2 =   setdiff( as.character(genes_public[["polyA"]]), unlist(unique(c(as.character(genes_s1[["polyA"]]),as.character(genes_s2[["polyA"]])))) )
s1_removed = s1_raw[ s1_raw$V1 %in% genes_notin_s1s2, ]
ggplot( (s1_removed %>% filter(V2=="polyA" | V2=="counts")) , aes(x=V2,y=V3) ) +
    geom_jitter() +
    labs(list(x="",y='counts'))
```



## Summary of samples with polyA>30

```{r polya-summary}
print("s1")
poly_read_summary(s1)
print("s2")
poly_read_summary(s2)
print("s3")
poly_read_summary(s3)
print("s4")
poly_read_summary(s4)
```


## Correlation


```{r cor-funct}
get_counts = function(counts){
    y = DGEList(counts=counts)
    y = calcNormFactors(y)
    cpm(y, normalized.lib.sizes=TRUE, log = TRUE)
}
```


### Total gene expression

```{r cor}
all = rbind( s1 %>% mutate( sample="s1"),
             s2 %>% mutate( sample="s2"),
             s3 %>% mutate( sample="s3"),
             s4 %>% mutate( sample="s4") )
all_melt = reshape(all %>% filter( V2=="counts" ), direction = "wide", idvar = c("V1","V2"),  timevar = "sample")
all_melt[is.na(all_melt)] = 0
counts = get_counts(all_melt[,3:6])
ggpairs(counts, alpha=0.4,
        diag = list(continuous = "density"))
ggheatmap.show(ggheatmap(cor(counts)))

```

### Genes expression of polyA-transcripts

```{r cor-polyA}
all_melt = reshape(all %>% filter( V2=="polyA" ), direction = "wide", idvar = c("V1","V2"),  timevar = "sample")
all_melt[is.na(all_melt)] = 0
ggpairs(log2(all_melt[,3:6]+0.1), alpha=0.4,
        diag = list(continuous = "density"))
ggheatmap.show(ggheatmap(cor(counts)))

```


### Genes expression of polyA + modification

```{r cor-mod}
all_melt = reshape(as.data.frame(all %>% filter( V2!="counts" & V3!="polya" ) %>% 
                                group_by(V1,sample) %>% summarise( total=sum(V3)))
                   , direction = "wide", idvar = c("V1"),  timevar = "sample")
all_melt[is.na(all_melt)] = 0
counts = get_counts(all_melt[,2:5])
ggpairs(counts, alpha=0.4,
        diag = list(continuous = "density"))
ggheatmap.show(ggheatmap(cor(counts)))

```


###Genes expression of polyA + U modification

```{r cor-a}
all_melt = reshape(all %>% filter( V2=="A" ), direction = "wide", idvar = c("V1","V2"),  timevar = "sample")
all_melt[is.na(all_melt)] = 0
counts = get_counts(all_melt[,3:6])
ggpairs(counts, alpha=0.4,
        diag = list(continuous = "density"))
ggheatmap.show(ggheatmap(cor(counts)))
```


### Genes expression of polyA + UU modification

```{r cor-aa}
all_melt = reshape(all %>% filter( V2=="AA" ), direction = "wide", idvar = c("V1","V2"),  timevar = "sample")
all_melt[is.na(all_melt)] = 0
counts = get_counts(all_melt[,3:6])
ggpairs(counts, alpha=0.4,
        diag = list(continuous = "density"))
ggheatmap.show(ggheatmap(cor(counts)))
```


### Genes expression of polyA + G modification

```{r cor-c}
all_melt = reshape(all %>% filter( V2=="C" ), direction = "wide", idvar = c("V1","V2"),  timevar = "sample")
all_melt[is.na(all_melt)] = 0
counts = get_counts(all_melt[,3:6])
ggpairs(counts, alpha=0.4,
        diag = list(continuous = "density"))
ggheatmap.show(ggheatmap(cor(counts)))

```

## By sample analysis

It included stats for each sample and GO enrichment analysis

```{r make-table-function}
make_table = function(data, prefix){
    raw = paste0(path_files,prefix,"_raw.txt")
    wide = paste0(path_files,prefix,"_wide.txt")
    write.table(data, raw,row.names=F, col.names=F)
    len = sapply(as.character(data$V2), nchar)
    data$type = "other"
    data$type[ data$V2=="counts" & data$type=="other"] = "counts"
    data$type[ data$V2=="polyA" & data$type=="other"] = "polyA"
    data$type[ data$V2=="A" & data$type=="other"] = "U"
    data$type[ data$V2=="AA" & data$type=="other"] = "UU"
    data$type[ data$V2=="C" & data$type=="other"] = "G"
    data$type[ data$V2=="CC" & data$type=="other"] = "GG"
    data$type[ len=1 & data$type=="other"] = "N"
    data$type[ len==2 & data$type=="other"] = "NN"
    data$type[ len==3 & data$type=="other"] = "NNN"
    data$type[ len==4 & data$type=="other"] = "NNNN"
    data$type[ len>4 & data$type=="other"] = "N>4"
    data = data %>% group_by(V1, type) %>% summarise(reads=sum(V3)) %>% ungroup()
    data$type = factor(data$type, levels=c("counts","polyA","U","UU","G","GG","N","NN","NNN","NNNN","N>4"))
    data = reshape(as.data.frame(data),  timevar = "type", idvar = "V1", direction = "wide")
    names(data) = sub("reads.","",names(data))
    data[is.na(data)] = 0
    write.table(data, wide,row.names=F, col.names=T)
    c(raw,wide)
}

```


### Sample 1

```{r do_go-s1, results='asis',cache=TRUE}
files = make_table(s1, "s1")
files = get_report_links(files)
read_summary(s1)
g_ezid = lapply(genes_s1, get_entrez)
#res_go = run_go_enrichment(g_ezid, "s1")
```
```{r table-s1, results='asis',eval=FALSE}
lapply(1:7, function(x){
    knitr::kable(res_go[[x]], caption=type[x])
})
```

[raw table](`r files[1]`) and [formatted table](`r files[2]`)

### Sample 2

```{r do_go-s2, results='asis',cache=TRUE}
read_summary(s2)
files = make_table(s1, "s2")
files = get_report_links(files)
g_ezid = lapply(genes_s2, get_entrez)
#res_go = run_go_enrichment(g_ezid, "s2")
```
```{r table-s2, results='asis',eval=FALSE}
lapply(1:7, function(x){
    knitr::kable(res_go[[x]], caption=type[x])
})
```

[raw table](`r files[1]`) and [formatted table](`r files[2]`)


### Sample 3

```{r do_go-s3, results='asis',cache=TRUE}
read_summary(s3)
files = make_table(s1, "s3")
files = get_report_links(files)
g_ezid = lapply(genes_s3, get_entrez)
#res_go = run_go_enrichment(g_ezid, "s3")
```
```{r table-s3, results='asis',eval=FALSE}
lapply(1:7, function(x){
    knitr::kable(res_go[[x]], caption=type[x])
})
```

[raw table](`r files[1]`) and [formatted table](`r files[2]`)


### Sample 4

```{r do_go-s4, results='asis',cache=TRUE}
read_summary(s4)
files = make_table(s1, "s4")
files = get_report_links(files)
g_ezid = lapply(genes_s4, get_entrez)
#res_go = run_go_enrichment(g_ezid, "s4")
```
```{r table-s4, results='asis',eval=FALSE}
lapply(1:7, function(x){
    knitr::kable(res_go[[x]], caption=type[x])
})
```

[raw table](`r files[1]`) and [formatted table](`r files[2]`)


### Files

```{r files_go, results='asis',eval=FALSE}
go_files = lapply(c("s1","s2","s3","s4"), function(s){
    files = lapply(1:7, function(x){
        data.frame(plot=paste0("<a href=",s,"_mf_",type_name[x],".png>","See plot for Molecular function for ",type[x]," in ", s,"<a/>"),
                   table=paste0("<a href=",s,"_mf_",type_name[x],".txt>"," Get table for Molecular function for ",type[x]," in ", s,"<a/>"))
        
    })
    do.call(rbind, files)
})
f = copy_batch_2_drop("txt",dn)
f = copy_batch_2_drop("png",dn)
knitr::kable(do.call(rbind, go_files))
```
