---
output:
  knitrBootstrap::bootstrap_document:
    theme: readable
    highlight: zenburn
    theme.chooser: TRUE
    highlight.chooser: TRUE
  html_document:
    toc: true
    highlight: zenburn
---

Report done by Lorena Pantano, last update  `r date()`

# enhancer activity detection

```{r setup}
library(knitr)

library(ggplot2)
library(reshape)
library(CHBUtils)
library(gtools)
library(gridExtra)
library(devtools)
library(dplyr)
library(Biostrings)
source("~/repos/myRfunctions/transactions.R")
knitr::opts_chunk$set(tidy=TRUE, highlight=TRUE, dev="png", fig.width=6,fig.heigh=6,
               cache=FALSE, highlight=TRUE, autodep=TRUE, warning=FALSE, error=FALSE,
               message=FALSE, prompt=TRUE, comment='', fig.cap='', bootstrap.show.code=FALSE)
```


```{r render, eval=FALSE, echo=FALSE}
library(rmarkdown)
library(knitrBootstrap)
render_2_drop("qc_pilot2.rmd",dn = "sui_enhancer/pilot2")
```


```{r samples-pilot2}
sw = list()
sw[['l1']] = read.table("~/orch/scratch/sui_enhancer/pilot2/res/L1_S3.R1_counts.tsv")
sw[['l2']] = read.table("~/orch/scratch/sui_enhancer/pilot2/res/L2_S4.R1_counts.tsv")
sw[['l3']] = read.table("~/orch/scratch/sui_enhancer/pilot2/res/L3_S5.R1_counts.tsv")
sw[['l4']] = read.table("~/orch/scratch/sui_enhancer/pilot2/res/L4_S6.R1_counts.tsv")
```


## General stats of novel vs known barcodes

```{r known_vs_novel}
detection = lapply(names(sw), function(x){
    sw[[x]] %>% group_by(V2) %>% dplyr::summarise(counts=n(),abundance=sum(V3)) %>% mutate(experiment=x)
})
dd = do.call(rbind, detection)

ggplot(dd, aes(x=experiment, y=counts, fill=V2)) +
    geom_bar(stat = 'identity', position = 'dodge') +
    scale_y_log10() +
    scale_fill_brewer("", palette = "Set1")

ggplot(dd, aes(x=experiment, y=abundance, fill=V2)) +
    geom_bar(stat = 'identity', position = 'dodge') +
    scale_y_log10() +
    scale_fill_brewer("", palette = "Set1")

```

## top different barcodes

```{r norm}
library(edgeR)
library(plyr)
cleaned = lapply(names(sw), function(x){
    sw[[x]] %>% filter(V3>1) %>% dplyr::select(V1, V3) %>% mutate(experiment=x)
})

dd = do.call(rbind, cleaned)
mat = reshape(dd, direction = 'wide', idvar = 'V1', timevar = 'experiment')
mat[is.na(mat)] = 0 
dge = edgeR::DGEList(mat[,2:5], genes = mat$V1)
norm = cpm(dge, normalized.lib.sizes = TRUE, log = TRUE)
rownames(norm) = dge$genes$genes
# pheatmap::pheatmap(norm)
```

```{r select}
mm = apply(norm, 1, min)
mx = apply(norm, 1, max)

keep = mx - mm > 10
pheatmap::pheatmap(norm[keep,], show_rownames=FALSE)

write.table(norm[keep,], "pilot2_difference.tsv", quote=F, sep="\t")
```


Download [file](`r get_report_links("pilot2_difference.tsv")`)

# Both pilots

```{r samples-pilot1}
# sw = list()
sw[['sw1']] = read.table("~/orch/scratch/sui_enhancer/pilot1/res9/052715_SW1_ATCACG_L001_R1_001_counts.tsv")
sw[['sw2']] = read.table("~/orch/scratch/sui_enhancer/pilot1/res25/052715_SW2_CGATGT_L001_R1_001_counts.tsv")
sw[['sw3']] = read.table("~/orch/scratch/sui_enhancer/pilot1/res25/052715_SW3_TTAGGC_L001_R1_001_counts.tsv")
sw[['sw4']] = read.table("~/orch/scratch/sui_enhancer/pilot1/res25/052715_SW4_TGACCA_L001_R1_001_counts.tsv")
```

## table with all samples

```{r edger.all}
cleaned = lapply(names(sw), function(x){
    sw[[x]] %>% filter(V3>1) %>% dplyr::select(V1, V3) %>% mutate(experiment=x)
})

dd = do.call(rbind, cleaned)
mat = reshape(dd, direction = 'wide', idvar = 'V1', timevar = 'experiment')
mat[is.na(mat)] = 0 
dge = edgeR::DGEList(mat[,2:9], genes = mat$V1)
norm = cpm(dge, normalized.lib.sizes = TRUE)
rownames(norm) = dge$genes$genes
sort.avg = sort(rowMeans(norm),decreasing = T, index.return=TRUE)
write.table( cbind(barcode=as.character(dge$genes$genes[sort.avg$ix]), round(norm[sort.avg$ix, ])), "matrix_all_norm.tsv", row.names=F, sep="\t", quote=F )

raw = dge$counts
sort.avg = sort(rowMeans(raw),decreasing = T, index.return=TRUE)
write.table( cbind(barcode=as.character(dge$genes$genes[sort.avg$ix]), raw[sort.avg$ix, ]), "matrix_all_raw.tsv", row.names=F, sep="\t", quote=F )

```


Download all raw data [file](`r get_report_links("matrix_all_raw.tsv")`)

Download all normalized data [file](`r get_report_links("matrix_all_norm.tsv")`)

