---
title: "chip-seq overlap and annotation"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    theme: readable
    highlight: zenburn
---
   
```{r projsetup, echo=FALSE, warning=FALSE, message=FALSE}
project="atac-seq"
labPI="Daley’s"
lablocation="Longwood"
analystname="Lorena Pantano"
analystemail="lpantano@hsph.harvard.edu"

```


```{r knitrsetup, echo=FALSE, cache=FALSE}
library(rmarkdown)
library(myRfunctions)
library(knitr)
knitr::opts_chunk$set(tidy=TRUE, highlight=TRUE, dev="png",
                      cache=TRUE, highlight=TRUE, autodep=TRUE, 
                      warning=FALSE, error=FALSE,
                      eval=TRUE, fig.width= 11, fig.height = 11,
                      message=FALSE, prompt=TRUE, comment='', fig.cap='', bootstrap.show.code=FALSE)
options(bitmapType = 'cairo')
```


```{r render,echo=FALSE,eval=FALSE}
# render_2_drop("analysis.Rmd", "daley/atacseq")
```

---

# Chip-seq analysis

Analysis for`r labPI` group.  

Contact `r analystname` (`r analystemail`) for additional details.

The most recent update of this html document occurred: `r date()`

The sections below provide code to reproduce the included results and plots. 

---

```{r functions}
## loading packages
require(ChIPseeker)
require(clusterProfiler)
library(dplyr)
library(ggplot2)
library(UpSetR)
library(GenomicRanges)
library(CHBUtils)
library(org.Hs.eg.db)
path_results = "/Users/lpantano/orch/scratch/daley_atac-seq/atac_1-merged/final/files"
dir.create(path_results,showWarnings = F, recursive = T)

```

```{r fn}
.summary <- function(obj){
    require(ggplot2)
    r <- slot(obj, "anno")
    df <- data.frame(FE=mcols(r)$score, A=as.character(mcols(r)$annotation),
                     stringsAsFactors = F)
    df$A[grepl("Intron", df$A)] <- "Intron"
    df$A[grepl("Exon", df$A)] <- "Exon"
    ggplot(df, aes(x=FE, fill=A)) + geom_density() +
        xlab("log10(Fold Enrichment)") +
        scale_x_log10() + facet_wrap(~A)
}

.go <- function(obj){
    r <- slot(obj, "anno")
    tb <- mcols(r)[grepl("Promoter", mcols(r)$annotation),]
    tb <- tb[order(tb$score, decreasing = TRUE),]
    # print(head(tb))
    g <- unique(tb$ENSEMBL)[1:1000]
    cat("Number of genes", length(g))
    # ego <- runGO(g, org.Hs.eg.db, from = "ENSEMBL", ont="BP")    
    ego <- enrichGO(gene = g, keytype = "ENSEMBL",
        OrgDb = org.Hs.eg.db, ont = "BP", pAdjustMethod = "BH", pvalueCutoff = 0.01, 
        qvalueCutoff = 0.05, readable = TRUE)
    ego = simplify(ego)
}
```


## Peak detection

We used macs2 with these parameters:

`/opt/bcbio/centos/bin/macs2 callpeak -t /n/data1/cores/bcbio/daley_atac-seq/atac_1-merged/work/align/EHT3d/EHT3d-sort.unique_filter.bam --nomodel  -f BAMPE -g 2.7e9 -n EHT3d -B  --nolambda --keep-dup all`

```{r load}
baseDir = "/Users/lpantano/orch/scratch/daley_atac-seq/atac_1-merged/final/idr_analysis/random1"

files <- list(EHT3d = c(file.path(baseDir, "EHT3d_0.2-idr")))


EHT3d = read.table(files$EHT3d, header=F, stringsAsFactors = F, comment.char = "#") %>%
    filter(abs(V5)>540)  %>%
    mutate(score=round(V15 * 10)) %>%
    dplyr::select(chr=V1,start=V2,end=V3,score)

EHT3d_chip = makeGRangesFromDataFrame(EHT3d, keep.extra.columns = TRUE, 
                                    ignore.strand = TRUE, end.field = "end",
                                    start.field = "start", seqnames.field = "chr")

```


```{r anno}
require(TxDb.Hsapiens.UCSC.hg19.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene

peakAnnoList <- lapply(list(EHT3d_chip), annotatePeak,
                       TxDb = txdb, tssRegion = c(-5000,5000), 
                       annoDb = "org.Hs.eg.db", verbose = FALSE)

names(peakAnnoList) = c("EHT3d")
```

## Annotation

```{r print-info, results='asis'}
sprintf("We have %d peaks for EHT3d", 
        length(EHT3d_chip))
```

```{r write-ann}
save_file(as.data.frame(peakAnnoList[["EHT3d"]]@anno), "EHT3d_macs_alone_idr.csv" ,path_results  )
```


### Genome annotation

Promoter defined as the region at +/- 5Kb from TSS

```{r plot-annotation, results='asis'}
ann = data.frame()
for (name in names(peakAnnoList)){
    .dt = peakAnnoList[[name]]@annoStat
    .dt$cells = name
    ann = rbind(ann, .dt)
}
cat("\n\n")
kable(ann)
cat("\n\n")
ggplot(ann, aes(x=Feature, y=Frequency, fill=cells)) +
    geom_bar(stat = 'identity', position = 'dodge') +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Go enrichment of top 1000 genes DE peaks

I took the top 100 genes with higest fold enrichment values for a GO enrichment analysis.

```{r go-diff, results='asis'}
for (name in names(peakAnnoList)){
    cat("\n\n", name, "\n\n")
    e = .go(peakAnnoList[[name]])
    print(dotplot(e))
    cat("\n\n")
    knitr::kable(e@result[,1:7])
    cat("\n\n")
}
```


#R Session Info

(useful if replicating these results)

```{r compress, echo=FALSE, cache=FALSE, eval=FALSE}
name_res = compress_results(path_results,prefix = "atacseq_results")
```

```{r sessioninfo}
sessionInfo()
```


