---
title: "chip-seq overlap and annotation"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    theme: readable
    highlight: zenburn
---
   
```{r projsetup, echo=FALSE, warning=FALSE, message=FALSE}
project="atac-seq"
labPI="Daley’s"
lablocation="Longwood"
analystname="Lorena Pantano"
analystemail="lpantano@hsph.harvard.edu"

```


```{r knitrsetup, echo=FALSE, cache=FALSE}
library(rmarkdown)
library(myRfunctions)
library(knitr)
knitr::opts_chunk$set(tidy=TRUE, highlight=TRUE, dev="png",
                      cache=TRUE, highlight=TRUE, autodep=TRUE, 
                      warning=FALSE, error=FALSE,
                      eval=TRUE, fig.width= 11, fig.height = 11,
                      message=FALSE, prompt=TRUE, comment='', fig.cap='', bootstrap.show.code=FALSE)
options(bitmapType = 'cairo')
```


```{r render,echo=FALSE,eval=FALSE}
# render_2_drop("analysis.Rmd", "daley/atacseq")
```

---

# Chip-seq analysis

Analysis for`r labPI` group.  

Contact `r analystname` (`r analystemail`) for additional details.

The most recent update of this html document occurred: `r date()`

The sections below provide code to reproduce the included results and plots. 

---

```{r functions}
## loading packages
require(ChIPseeker)
require(clusterProfiler)
library(dplyr)
library(ggplot2)
library(UpSetR)
library(GenomicRanges)
library(CHBUtils)
library(org.Hs.eg.db)
require(TxDb.Hsapiens.UCSC.hg19.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene
path_results = "/Users/lpantano/orch/scratch/daley_atac-seq/atac_1-merged/final/files"
dir.create(path_results,showWarnings = F, recursive = T)

```

```{r fn}
.summary <- function(obj){
    require(ggplot2)
    r <- slot(obj, "anno")
    df <- data.frame(FE=mcols(r)$fold_enrichment, A=as.character(mcols(r)$annotation),
                     stringsAsFactors = F)
    df$A[grepl("Intron", df$A)] <- "Intron"
    df$A[grepl("Exon", df$A)] <- "Exon"
    ggplot(df, aes(x=FE, fill=A)) + geom_density() +
        xlab("log10(Fold Enrichment)") +
        scale_x_log10() + facet_wrap(~A)
}

.go <- function(obj){
    r <- slot(obj, "anno")
    tb <- mcols(r)[grepl("Promoter", mcols(r)$annotation),]
    tb <- tb[order(tb$fold_enrichment, decreasing = TRUE),]
    # print(head(tb))
    g <- unique(tb$ENSEMBL)[1:1000]
    cat("Number of genes", length(g))
    # ego <- runGO(g, org.Hs.eg.db, from = "ENSEMBL", ont="BP")    
    ego <- enrichGO(gene = g, keytype = "ENSEMBL",
        OrgDb = org.Hs.eg.db, ont = "BP", pAdjustMethod = "BH", pvalueCutoff = 0.01, 
        qvalueCutoff = 0.05, readable = TRUE)
    ego = simplify(ego)
}
```


## Peak detection

We used macs2 with these parameters:

`/opt/bcbio/centos/bin/macs2 callpeak -t /n/data1/cores/bcbio/daley_atac-seq/atac_1-merged/work/align/EHT3d/EHT3d-sort.unique_filter.bam --nomodel  -f BAMPE -g 2.7e9 -n EHT3d -B  --nolambda --keep-dup all`

```{r load}
baseDir = "/Users/lpantano/orch/scratch/daley_atac-seq/atac_1-merged/final"
files <- list(EHT3d = c(file.path(baseDir, "EHT3d", "EHT3d-macs2.xls")), 
              EHT7d = c(file.path(baseDir, "EHT7d", "EHT7d-macs2.xls")))


EHT7d = read.table(files$EHT7d, header=T, stringsAsFactors = F, comment.char = "#") %>%
    filter(abs(fold_enrichment)>4) %>%
    mutate(score=round(X.log10.qvalue. * 10))
EHT3d = read.table(files$EHT3d, header=T, stringsAsFactors = F, comment.char = "#") %>%
    filter(abs(fold_enrichment)>4)  %>%
    mutate(score=round(X.log10.qvalue. * 10))

EHT7d_chip = makeGRangesFromDataFrame(EHT7d, keep.extra.columns = TRUE, 
                                    ignore.strand = TRUE, end.field = "end",
                                    start.field = "start", seqnames.field = "chr")
EHT3d_chip = makeGRangesFromDataFrame(EHT3d, keep.extra.columns = TRUE, 
                                    ignore.strand = TRUE, end.field = "end",
                                    start.field = "start", seqnames.field = "chr")

```


```{r anno}
peak_rda = file.path(baseDir, "peaks.rda")
if (!file.exists(peak_rda)){
    peakAnnoList <- lapply(list(EHT3d_chip, EHT7d_chip), annotatePeak,
                           TxDb = txdb, tssRegion = c(-5000,5000), 
                           annoDb = "org.Hs.eg.db", verbose = FALSE)
    
    save(peakAnnoList, file=peak_rda)
}else{
    load(peak_rda)
}
names(peakAnnoList) = c("EHT3d", "EHT7d")
```

## Annotation

```{r print-info, results='asis'}
sprintf("We have %d peaks for EHT3d and %d peaks for EHT7d.", 
        length(EHT3d_chip),
        length(EHT7d_chip))
```

```{r write-ann}
save_file(as.data.frame(peakAnnoList[["EHT3d"]]@anno), "EHT3d_macs_alone.csv" ,path_results  )
save_file(as.data.frame(peakAnnoList[["EHT7d"]]@anno), "EHT7d_macs_alone.csv" ,path_results  )
```


### Genome annotation

Promoter defined as the region at +/- 5Kb from TSS

```{r plot-annotation, results='asis'}
ann = data.frame()
for (name in names(peakAnnoList)){
    .dt = peakAnnoList[[name]]@annoStat
    .dt$cells = name
    ann = rbind(ann, .dt)
}
cat("\n\n")
kable(ann)
cat("\n\n")
ggplot(ann, aes(x=Feature, y=Frequency, fill=cells)) +
    geom_bar(stat = 'identity', position = 'dodge') +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Fold Enrichment distribution

We show here how the fold enrichment distrution is for each type of peaks according
genomic annotation.

```{r dist, results='asis'}
for (name in names(peakAnnoList)){
    cat("\n\n", name, "\n\n")
    print(.summary(peakAnnoList[[name]]))
}
```

# Comparison

```{r macs_diff}
files <- list(EHT3d = c(file.path(baseDir, "EHT3d_diff", "EHT3d_diff_peaks.xls")), 
              EHT7d = c(file.path(baseDir, "EHT7d_diff", "EHT7d_diff_peaks.xls")))


EHT7d_de = read.table(files$EHT7d, header=T, stringsAsFactors = F, comment.char = "#") %>%
    mutate(score=round(X.log10.qvalue. * 10))
EHT3d_de = read.table(files$EHT3d, header=T, stringsAsFactors = F, comment.char = "#") %>%
    mutate(score=round(X.log10.qvalue. * 10))

EHT7d_de_chip = makeGRangesFromDataFrame(EHT7d_de, keep.extra.columns = TRUE, 
                                    ignore.strand = TRUE, end.field = "end",
                                    start.field = "start", seqnames.field = "chr")
EHT3d_de_chip = makeGRangesFromDataFrame(EHT3d_de, keep.extra.columns = TRUE, 
                                    ignore.strand = TRUE, end.field = "end",
                                    start.field = "start", seqnames.field = "chr")

```

We used macs2 to get the peaks that were enriched in one or another sample.
We use the other sample as control to do that, to get some kind of statistical 
analysis to remove bias due to coverage.

* EHT7d_diff are the peaks enriched in EHT7d (control was EHT3d)
* EHT3d_diff are the peaks enriched in EHT3d (control was EHT7d)

```{r ann-diff}
peak_diff_rda = file.path(baseDir, "peaks_diff.rda")
if (!file.exists(peak_diff_rda)){
    peakAnnoDeList <- lapply(list(EHT3d_de_chip, EHT7d_de_chip), annotatePeak,
                           TxDb = txdb, tssRegion = c(-5000,5000), 
                           annoDb = "org.Hs.eg.db", verbose = FALSE)
    
    save(peakAnnoDeList, file=peak_diff_rda)
}else{
    load(peak_diff_rda)
}
names(peakAnnoDeList) = c("EHT3d", "EHT7d")

```

## Annotation

```{r print-info-diff, results='asis'}
sprintf("We have %d peaks for EHT3d and %d peaks for EHT7d.", 
        length(EHT3d_de_chip),
        length(EHT7d_de_chip))
```


```{r write-ann-diff}
save_file(as.data.frame(peakAnnoDeList[["EHT3d"]]@anno), "EHT3d_macs_diff.csv" ,path_results  )
save_file(as.data.frame(peakAnnoDeList[["EHT7d"]]@anno), "EHT7d_macs_diff.csv" ,path_results  )
```

### Genome annotation
```{r plot-annotation-diff, results='asis'}
ann = data.frame()
for (name in names(peakAnnoDeList)){
    .dt = peakAnnoDeList[[name]]@annoStat
    .dt$cells = name
    ann = rbind(ann, .dt)
}
cat("\n\n")
kable(ann)
cat("\n\n")
ggplot(ann, aes(x=Feature, y=Frequency, fill=cells)) +
    geom_bar(stat = 'identity', position = 'dodge') +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

### Fold Enrichment distribution
```{r dist-diff, results='asis'}
for (name in names(peakAnnoDeList)){
    cat("\n\n", name, "\n\n")
    print(.summary(peakAnnoDeList[[name]]))
}
```

### Go enrichment of top 1000 genes DE peaks

I took the top 100 genes with higest fold enrichment values for a GO enrichment analysis.

```{r go-diff, results='asis'}
for (name in names(peakAnnoDeList)){
    cat("\n\n", name, "\n\n")
    e = .go(peakAnnoDeList[[name]])
    print(dotplot(e))
    cat("\n\n")
    knitr::kable(e@result[,1:7])
    cat("\n\n")
}
```


#R Session Info

(useful if replicating these results)

```{r compress, echo=FALSE, cache=FALSE, eval=FALSE}
name_res = compress_results(path_results,prefix = "atacseq_results")
```

```{r sessioninfo}
sessionInfo()
```


