---
title: "RNA-seq analysis."
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
    theme: readable
    highlight: zenburn
---

```{r projsetup, echo=FALSE, warning=FALSE, message=FALSE}
project="LGN - RNAseq "
labPI="Greenberg"
lablocation="Longwood"
analystname="Lorena Pantano"
analystemail="lpantano@hsph.harvard.edu"
```

RNA-seq analysis for `r labPI` group at `r lablocation`.  

Contact `r analystname` (`r analystemail`) for additional details.

The most recent update of this html document occurred: `r date()`

The sections below provide code to reproduce the included results and plots. 


```{r knitrsetup, echo=FALSE}
library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(dev="png",echo=TRUE,
                      highlight=TRUE, autodep=TRUE,
                      comment='', fig.cap='',
                      tidy.opts=list(keep.blank.line=FALSE, width.cutoff=200),
                      cache=TRUE,warning=FALSE,message=FALSE,
                      error=FALSE, fig.width = 12, fig.height = 9)
options(bitmapType = 'cairo')
```

```{r project-options, echo=FALSE}
library(myRfunctions)
path_results = "~/orch/scratch/greenberg_rnaseq/lgn2/final/2016-05-06_lgn2/files"
dir.create(path_results, showWarnings = FALSE)
```

```{r create-report, echo=FALSE, eval=FALSE}
render_2_drop("mrna-goenrichment.Rmd", "greenberg/lgn-go_enrichment")
```


```{r set-up}
library(ggplot2)
library(reshape)
library(gplots)
library(CHBUtils)
library(DESeq2)
library(logging)
library(gridExtra)
library(clusterProfiler)
library(DOSE)
library(org.Mm.eg.db)
library(cluster)
library(dplyr)
basicConfig()

comp_keep = c("0_vs_1.tsv", "0_vs_3.tsv", "0_vs_8.tsv", 
              "1_vs_3.tsv", "1_vs_8.tsv", "3_vs_8.tsv")
comp = list()
for (name in comp_keep){
    comp[[name]] = readr::read_csv(file.path(path_results, name))
}


# counts = read.csv(file.path(path_results, "log2_counts.csv"), row.names=1)
# tpm = read.table(file.path(path_results, "..","combined.gene.sf.tpm"), row.names=1, header=T)
# 
# sig = comp[["0_vs_8.tsv"]][comp[["0_vs_8.tsv"]]$padj<0.05 & abs(comp[["0_vs_8.tsv"]]$log2FoldChange)>0.3,"id"]
# h0 = rowMeans(tpm[,1:3])
# h8 = rowMeans(tpm[,10:12])
# col = rep("grey",length(h0))
# col[match(sig$id,rownames(tpm))] = "red"
#plot(log2(h0+1),log2(h8+1),pch=20, col = "grey")
#points(log2(h0[match(sig$id,rownames(tpm))]+1),log2(h8[match(sig$id,rownames(tpm))]+1),pch=20, col = "red")
```


```{r load-go-setup}
# https://www.biostars.org/p/52101/#68158
library(biomaRt)
ensembl = useMart("ensembl",dataset="mmusculus_gene_ensembl") #uses 

asd_files = list.files("~/orch/scratch/greenberg_rnaseq/lgn2/final/", pattern = "txt", full.names = T)

asd_terms = do.call(rbind, lapply(asd_files, function(fn){
    
    if (!grepl("term", fn)){
        goterms =  readr::read_tsv(fn, col_names = F) %>% mutate(gene=X1)
        gene.goterms <- getBM(attributes=c('external_gene_name', 'ensembl_gene_id'),
                              filters = 'external_gene_name', 
                              values = goterms$gene, 
                              mart = ensembl)
        return(data.frame(go=basename(fn), genes=as.character(gene.goterms$ensembl_gene_id)))
    }else{
        goterms =  readr::read_tsv(fn, 
                           col_names = FALSE) %>%
        tidyr::separate(X1, into=c("id", "term"), sep="  ")    
        gene.goterms <- getBM(attributes=c('ensembl_gene_id', 'go_id'),
                   filters = 'go_id', values = goterms$id, mart = ensembl)

        return(data.frame(go=basename(fn), genes=unique(as.character(gene.goterms$ensembl_gene_id))))
    }
}))


trasnmembrane_domain = getBM(attributes=c("ensembl_gene_id", "mgi_symbol",
                                "transmembrane_domain"), mart=ensembl)

asd_terms = rbind(asd_terms, data.frame(go="transmembrane",
                                        genes=unique(as.character(trasnmembrane_domain$ensembl_gene_id[trasnmembrane_domain$transmembrane_domain!=""]))))

write.csv(asd_terms, "~/orch/scratch/greenberg_rnaseq/lgn2/final/all-terms.csv")
```

## GO enrichment using DE genes

We considered DE genes as genes with FDR < 10%.


```{r clusterprofile, results='asis'}
go_res = data.frame()
for (i in 1:length(comp)){
    cat("### Go enrichment for:", names(comp)[i], "\n\n")
    keep = comp[[i]]$padj<0.1 & !is.na(comp[[i]]$padj)
    cat("Working with ", sum(keep), " genes\n\n")
    go_fake = rbind(asd_terms, data.frame(go="GO_control", genes=comp[[i]]$id))
    
    go_name = data.frame(go=c(as.character(unique(asd_terms$go)),
                              "GO_everything.else"), 
                         name=c(as.character(unique(asd_terms$go)), "control"))
    
    if (sum(keep) == 0){
        cat("\n\nNo DE genes in this comparison\n\n");next}
    
    cp = enricher( as.character(comp[[i]]$id[keep]), universe = as.character(comp[[i]]$id), 
                   pvalueCutoff = 1, qvalueCutoff = 1, 
                   TERM2GENE = go_fake, maxGSSize = 8000,
                   TERM2NAME = go_name, minGSSize = 2)
    cat("\n")
    if (!("result" %in% slotNames(cp))){
        cat("\n\nNo Genes in any terms in this comparison\n\n");next}
        
    print(kable(cp@result[,1:6]))
    # cat("\n\ngenes in GO term\n\n")
    # print(paste(
    symbol = paste(convertIDs(unique(unlist(strsplit(cp@result$geneID,"/"))),
                          "ENSEMBL", "SYMBOL", 
                          org.Mm.eg.db, "useFirst"), collapse = " ")
    cat("\n\n")
    dt = cp@result
    dt$symbol = symbol
    dt$comparision = names(comp)[i]
    go_res = rbind(go_res, dt)
}
write.csv(go_res, "~/orch/scratch/greenberg_rnaseq/lgn2/final/go-res.csv")
```


#R Session Info

```{r compress, echo=FALSE, cache=FALSE}
#name_res = compress_results(path_results,prefix = "ldr_results")
```

(useful if replicating these results)

```{r sessioninfo}
sessionInfo()
```
